---
title: 'Lathyrus ms2: Calculation number of years FFD, flowering, fitness, flowering with fitness'
output:
  word_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    toc: yes
    toc_depth: '4'
  pdf_document:
    toc: yes
    toc_depth: 4
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
```

Read definitive data file for the old period

```{r}
datadef_old<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/datadef_old.csv")%>%
  mutate(period="old")
head(datadef_old)
```

Extract data for the new period from data used for ms1

```{r}
datadef_new<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms1/data/clean/alldata.csv",
  header=T,sep="\t",dec=".")[c(1:3,10,12,14,16,18:19)]%>%
  filter(period=="new")%>% # Keep only data from new period
  rename(totseed=n_seeds,intactseed=n_intact_seeds)%>% # Column names similar to datadef_old
  mutate(id_nr=as.integer(gsub("new_","",id)))
head(datadef_new)
```

Read info about number of shoots and shoot volume (to see when plants were not flowering but were vegetative)

```{r}
nshoots_vol<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/info_nshoots_vol_0617.csv",
  header=T,sep=",",dec=".")
nshoots_vol$id_nr<-nshoots_vol$id
nshoots_vol$id<-NULL
head(nshoots_vol)
```

```{r}
datadef_new<-datadef_new%>%
  full_join(nshoots_vol)%>%
  mutate(fcode=as.integer(ifelse(!is.na(FFD)|n_fl>0|n_fr>0|totseed>0|intactseed>0,1,NA)))%>%
  mutate(fcode=as.integer(ifelse(is.na(fcode)&info_nshoots_vol==1,0,fcode)))%>%
  # 1 = flowering, 0 = not flowering, NA = not known
  select(year,id_nr,id,fcode,FFD,n_fl,n_fr,totseed,intactseed,shoot_vol,period)%>%
  # Column order similar to datadef_old
  arrange(id_nr,year)%>%
  mutate(n_fl=ifelse(fcode==0,0,n_fl))%>% # When fcode=0, set n_fl=0
  mutate(n_fr=ifelse(fcode==0,0,n_fr))%>% # When fcode=0, set n_fr=0
  mutate(totseed=ifelse(fcode==0,0,totseed))%>% # When fcode=0, set totseed=0
  mutate(intactseed=ifelse(fcode==0,0,intactseed)) # When fcode=0, intactseed=0
head(datadef_new)
```

Merge the definitive datasets for old and new period

```{r}
datadef<-rbind(datadef_old,datadef_new)
head(datadef)
```

In the table below, for each year, 1 means that there is a value of the variable for that year, and 0 means that there is no value (NA).

```{r}
(variables_def<-pivot_longer(datadef, names_to="variable", values_to="data", fcode:shoot_vol)%>%
   pivot_wider(.,names_from=year,values_from=data)%>%
   rename(y1987="1987",y1988="1988",y1989="1989",y1990="1990",y1991="1991",y1992="1992",
          y1993="1993",y1994="1994",y1995="1995",y1996="1996",
          y2006="2006",y2007="2007",y2008="2008",y2009="2009",y2010="2010",y2011="2011",
          y2012="2012",y2013="2013",y2014="2014",y2015="2015",y2016="2016",y2017="2017")%>%
   mutate(y1987=ifelse(is.na(y1987),0,1),y1988=ifelse(is.na(y1988),0,1),
          y1989=ifelse(is.na(y1989),0,1),y1990=ifelse(is.na(y1990),0,1),
          y1991=ifelse(is.na(y1991),0,1),y1992=ifelse(is.na(y1992),0,1),
          y1993=ifelse(is.na(y1993),0,1),y1994=ifelse(is.na(y1994),0,1),
          y1995=ifelse(is.na(y1995),0,1),y1996=ifelse(is.na(y1996),0,1),
          y2006=ifelse(is.na(y2006),0,1),y2007=ifelse(is.na(y2007),0,1),
          y2008=ifelse(is.na(y2008),0,1),y2009=ifelse(is.na(y2009),0,1),
          y2010=ifelse(is.na(y2010),0,1),y2011=ifelse(is.na(y2011),0,1),
          y2012=ifelse(is.na(y2012),0,1),y2013=ifelse(is.na(y2013),0,1),
          y2014=ifelse(is.na(y2014),0,1),y2015=ifelse(is.na(y2015),0,1),
          y2016=ifelse(is.na(y2016),0,1),y2017=ifelse(is.na(y2017),0,1))%>%
   select(sort(colnames(.))))
```
In the table below, 1 means that the plant was flowering, 0 that it was not flowering (vegetative), and NA that no data is available for that year

```{r}
(flowering_def<-datadef%>%
    select(year,id,fcode)%>%
    pivot_wider(names_from=year,values_from=fcode)%>%
   select("id",sort(colnames(.))))
```
Calculate number of years with information on FFD (n\_years\_FFD), number of years of flowering (n\_years\_flowering), number of years with information on fitness, and number of flowering years with information on fitness for each plant

```{r}
n_years_def<-full_join(
  (subset(variables_def,variable=="FFD")%>%
     mutate(n_years_FFD = rowSums(.[5:26],na.rm=T))%>% # n_years_FFD
     select(id,id_nr,period,n_years_FFD)),
  flowering_def%>%
    mutate(n_years_flowering = rowSums(.[2:23],na.rm=T))%>% # n_years_flowering
    select(id,n_years_flowering))
n_years_def<-full_join(n_years_def,
                   (subset(variables_def,variable=="intactseed")%>%
     mutate(n_years_fitness = rowSums(.[5:26],na.rm=T))%>%
     select(id,id_nr,period,n_years_fitness))) # n_years_fitness
n_years_def<-full_join(n_years_def,
                   subset((pivot_longer(datadef, names_to="variable", values_to="data", fcode:shoot_vol)),
                          variable=="intactseed"|variable=="fcode")%>%
                     pivot_wider(names_from="variable",values_from="data")%>%
                     filter(fcode==1)%>% # Keep only flowering years
                     pivot_wider(names_from="year",values_from="intactseed")%>%
  rename(y1987="1987",y1988="1988",y1989="1989",y1990="1990",y1991="1991",y1992="1992",
          y1993="1993",y1994="1994",y1995="1995",y1996="1996",
          y2006="2006",y2007="2007",y2008="2008",y2009="2009",y2010="2010",y2011="2011",
          y2012="2012",y2013="2013",y2014="2014",y2015="2015",y2016="2016",y2017="2017")%>%
  mutate(y1987=ifelse(is.na(y1987),0,1),y1988=ifelse(is.na(y1988),0,1),
          y1989=ifelse(is.na(y1989),0,1),y1990=ifelse(is.na(y1990),0,1),
          y1991=ifelse(is.na(y1991),0,1),y1992=ifelse(is.na(y1992),0,1),
          y1993=ifelse(is.na(y1993),0,1),y1994=ifelse(is.na(y1994),0,1),
          y1995=ifelse(is.na(y1995),0,1),y1996=ifelse(is.na(y1996),0,1),
          y2006=ifelse(is.na(y2006),0,1),y2007=ifelse(is.na(y2007),0,1),
          y2008=ifelse(is.na(y2008),0,1),y2009=ifelse(is.na(y2009),0,1),
          y2010=ifelse(is.na(y2010),0,1),y2011=ifelse(is.na(y2011),0,1),
          y2012=ifelse(is.na(y2012),0,1),y2013=ifelse(is.na(y2013),0,1),
          y2014=ifelse(is.na(y2014),0,1),y2015=ifelse(is.na(y2015),0,1),
          y2016=ifelse(is.na(y2016),0,1),y2017=ifelse(is.na(y2017),0,1))%>%
  mutate(n_years_fl_fitness = rowSums(.[5:26],na.rm=T))%>%
  select(id,id_nr,period,n_years_fl_fitness) # n_years_fl_fitness
)
head(n_years_def)
```

Remove ids with n_years_FFD==0

```{r}
datadef<-anti_join(datadef,
                   data.frame(id=unique(subset(n_years_def,n_years_FFD==0)$id)))
n_years_def<-n_years_def%>%
  filter(n_years_FFD>0)
```


Read data on duration of the study:

```{r}
duration_study<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/duration_study.csv")
head(duration_study)
```

Add info on duration of the study (n_years_study) to n_years_def:

```{r}
n_years_def<-full_join(n_years_def,duration_study)
head(n_years_def)
```

Some plots:

```{r}
ggplot(n_years_def,aes(x=n_years_study,fill=period))+
  geom_histogram(binwidth=1,position="dodge")
ggplot(n_years_def,aes(x=n_years_fl_fitness,fill=period))+
  geom_histogram(binwidth=1,position="dodge")
```

Add info on n_years_fl_fitness and n_years_study to datadef

```{r}
datadef<-full_join(datadef,n_years_def[c(1:3,7:8)])
head(datadef)
```

Add temperature data (mean April temp) to datadef

```{r}
data_sel<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms1/data/clean/data_sel.csv",
                       header=T,sep="\t",dec=",") 
datadef<-full_join(datadef,unique(data_sel[c(1,171)]))
datadef$cmean_4<-scale(datadef$mean_4,center=T,scale=F) # Center temperature variable
head(datadef)
```

datadef is ready for analyses!

```{r}
write.csv(datadef,
          file="C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/datadef.csv",
          row.names=F)
```


