---
title: "Spring temperature drives phenotypic selection on plasticity of flowering time"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.width=4, fig.height=3) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
library(brms)
library(ggthemes)
library(knitr)
library(bayesplot)
library(tidybayes)
library(parallel)
library(future)
library(ggeffects)
library(arm)
library(cowplot)
library(broom)
library(viridis)
library(MCMCglmm)
```

```{r Define ggplot themes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r load models, include=FALSE}
# Load previously run models and stuff
load("output/large_objects/large_objects_3.RData")
```

# Read data and check ns

```{r}
datadef<-read.csv("data/datadef.csv") 
head(datadef)
```

Number of individuals in each period:

```{r}
length(with(subset(datadef,period=="old"),unique(id)))
length(with(subset(datadef,period=="new"),unique(id)))
```

Number of observations in each period:

```{r}
nrow(subset(datadef,period=="old"))
nrow(subset(datadef,period=="new"))
```

Number of cases with FFD in each period:

```{r}
nrow(subset(datadef,period=="old"&!is.na(FFD)))
nrow(subset(datadef,period=="new"&!is.na(FFD)))
```

# Data preparation

```{r data prep MCMCglmm, message=FALSE, warning=FALSE}
datadef_total<-datadef %>%
  group_by(id)%>%
  # Calculate mean fitness per year of study 
  # and mean fitness per flowering event
  summarise(mean_fitness_study=sum(intactseed,na.rm=T)/mean(n_years_study),
            mean_fitness_fl=sum(intactseed,na.rm=T)/mean(n_years_fl_fitness))%>%
   arrange(.,id) # Order by id

with(datadef_total,cor(mean_fitness_study,mean_fitness_fl))  # Highly corr (0.87)

# Calculate mean shoot volume for each id using values of shoot volume 
# for all ids/years (including flowering and non-flowering years)

shoot_vol_all_means<-datadef[c(1,3,10)]%>%
  group_by(id)%>%
  summarise(shoot_vol_mean=mean(shoot_vol,na.rm=T)) 
# Mean of all available values 

# Join shoot volume data
datadef_total<-datadef_total%>%left_join(shoot_vol_all_means)%>%
  left_join(unique(datadef[c(2,3,11)]))
head(datadef_total)
nrow(subset(datadef_total,is.na(shoot_vol_mean))) 
# 46 ids with no info on shoot volume

# Using sqrt of mean shoot volume over all years when available, centred
datadef_total<-datadef_total%>%
  mutate(shoot_vol_mean_sqrt=sqrt(shoot_vol_mean),
         cn_shoot_vol_mean_sqrt=as.vector(scale(shoot_vol_mean_sqrt,center=T,scale=F)))
```

Filter data to keep individuals that flowered in at least 3 years.

```{r}
datadef3<-datadef%>%filter(n_years_fl_fitness>2)
nrow(filter(datadef3,!is.na(FFD))) 
length(unique((datadef%>%filter(n_years_fl_fitness>2))$id))
```

1977 flowering events and 424 individuals.

# Hypothesis 1: Univariate models

## Models in Table 1

### Table 1A

FFD with random effect of year only. This model assumes no among-individual variation around the average population-level reaction norm.

```{r}
my.cores <- detectCores()
```

FFD with random effect of year only. This model assumes no among-individual variation around the average population-level reaction norm.

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm3<-brm(formula=FFD~cmean_4+(1|year),data=datadef3,
                warmup = 1000,iter = 4000,thin=2,chains=4,
                # 4 chains, each with 4000 iterations
                inits = "random",seed = 12345,cores = my.cores)
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_yearonly.all.brm3)
```

### Table 1B

FFD with random effects of year and individual-intercept. This model assumes among-individual variation in reaction norm elevation.

```{r eval=FALSE, include=FALSE}
univar.FFD.all.brm3<-brm(formula=FFD~cmean_4+(1|year)+(1|id),data=datadef3,
                warmup = 1000,iter = 4000,thin=2,chains=4,
                # 4 chains, each with 4000 iterations
                inits = "random",seed = 12345,cores = my.cores)
# Total of 6000 post-warmup samples
```

```{r }
summary(univar.FFD.all.brm3)
```

### Table 1C

Random regression for FFD, including random effects of individual slopes and covariance between intercept and slope. This model assumes among-individual variation in reaction norm elevation and slope.

```{r eval=FALSE, include=FALSE}
univar.FFD_RR.all.brm3<-brm(formula=FFD~cmean_4+(1|year)+(cmean_4|id),
                           data=datadef3,
                           warmup = 1000,iter = 4000,thin=2,chains=4,
                           inits = "random",seed = 12345,cores = my.cores,
                           sample_prior="yes")
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_RR.all.brm3) 
```

#### Compare models

This allows  us to test if adding among-individual variation in elevation and slope of the RN improves model fit. 

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm3 <- add_criterion(univar.FFD_yearonly.all.brm3,"loo")
univar.FFD.all.brm3 <- add_criterion(univar.FFD.all.brm3,"loo")
univar.FFD_RR.all.brm3 <- add_criterion(univar.FFD_RR.all.brm3,"loo")
```

```{r}
loo_comp3<-loo_compare(univar.FFD_yearonly.all.brm3,univar.FFD.all.brm3,
                      univar.FFD_RR.all.brm3, criterion="loo")
```

```{r}
loo_comp3
```

The results indicate that adding a random intercept (i.e. among-individual variation in RN elevation) improves the fit, and that adding a random slope (i.e. among-individual variation in RN slope) improves the fit even more. Thus, the random regression model is the best model.

#### Bayes_R2

Measure of fit: Bayesian R2, which looks at the model expected variance / (expected variance + residual variance).

```{r}
bayes_R2(univar.FFD_yearonly.all.brm3)
bayes_R2(univar.FFD.all.brm3)
bayes_R2(univar.FFD_RR.all.brm3)
```

#### Extract BLUPs from random regression model

```{r}
BLUPs_MCMC.all.brms3  <- cbind(as.factor(c(1:424)),
                              as.data.frame(ranef(univar.FFD_RR.all.brm3)$id)
                              [c(1:2,5:6)])
colnames(BLUPs_MCMC.all.brms3) <- c("id", "intercept", "intercept_sd",
                                   "slope","slope_sd")
with(BLUPs_MCMC.all.brms3,cor(intercept,slope)) # highly correlated!
```

Add BLUPs to data set

```{r}
datadef3<-datadef3%>%left_join(BLUPs_MCMC.all.brms3%>%
                               dplyr::select(intercept,intercept_sd,
                                             slope,slope_sd)%>%
                               rownames_to_column(var="id"))
```

Save data with BLUPs

```{r}
datadef_total3<-datadef3 %>%
  group_by(id)%>%
  # Calculate mean fitness per year of study 
  # and mean fitness per flowering event
  summarise(mean_fitness_study=sum(intactseed,na.rm=T)/mean(n_years_study),
            mean_fitness_fl=sum(intactseed,na.rm=T)/mean(n_years_fl_fitness))%>%
   arrange(.,id) # Order by id

with(datadef_total3,cor(mean_fitness_study,mean_fitness_fl))  # Highly corr (0.91)

# Join shoot volume data
datadef_total3<-datadef_total3%>%left_join(shoot_vol_all_means)%>%
  left_join(unique(datadef3[c(2,3,11)]))
head(datadef_total3)
nrow(subset(datadef_total3,is.na(shoot_vol_mean))) 
# 0 ids with no info on shoot volume

# Using sqrt of mean shoot volume over all years when available, centred
datadef_total3<-datadef_total3%>%
  mutate(shoot_vol_mean_sqrt=sqrt(shoot_vol_mean),
         cn_shoot_vol_mean_sqrt=as.vector(scale(shoot_vol_mean_sqrt,center=T,
                                                scale=F)))

datadef3<-left_join(datadef3,datadef_total3[c(1:3,8)]) 
```

```{r}
datadef3<-datadef3%>%left_join(datadef_total3%>%
                                 dplyr::select(id,cn_shoot_vol_mean_sqrt))
write_csv(datadef3,file = "data/datadef3_BLUPs.csv")
```

## Hypothesis 2: Bivariate models

## Models in Table 2

### Table 2A NB

Mean fitness per flowering event, with shoot volume.

Using the ID-syntax to specify fitness to be correlated with the intercept and slope of FFD on temperature.

Negative binomial distribution for fitness.

```{r}
bf_FFD <- bf(FFD ~ cmean_4 + (cmean_4|ID1|id) + (1|year)) # Set up model formula
bf_fitness_shoot <- bf(round(mean_fitness_fl) ~  cn_shoot_vol_mean_sqrt +
                         (1|ID1|id)) # Set up model formula
# Specifying group-level effects of the same grouping factor (id here) 
# to be correlated across formulas
# Expand the | operator into |<ID>|, where <ID> can be any value (ID1 here)
# Group-level terms with the same ID1 will be modeled as correlated 
# if they share same grouping factor(s)
```

```{r, eval=FALSE, include=TRUE}
bivar2.all.brm.nb3<-brm(bf_FFD+bf_fitness_shoot, 
                       family = c(gaussian,negbinomial), 
                       data = datadef3,warmup = 1000,iter = 6000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99))
```

```{r}
print(bivar2.all.brm.nb3,digits=3)
```

Measure of fit: Bayesian R2

```{r}
bayes_R2(bivar2.all.brm.nb3)
```

#### Extract selection coefficients

```{r}
sample.gmat2 <- function(data, replicates = 6000) {

  ##Initialize the results list (list of lists)
  foo <- list(gmat = matrix(rep(0,3*3), ncol = 3))
  results.list <- list()
  for(j in 1:replicates) { results.list[[j]] <- foo }

  for(i in 1:replicates) {
    diag(results.list[[i]]$gmat) <- data[i,5:7]^2 #Get the diagonal

    #Upper diagonal
    results.list[[i]]$gmat[1,2] <- data[i,5]*data[i,6]*data[i,9]
    results.list[[i]]$gmat[1,3] <- data[i,5]*data[i,7]*data[i,10]
    results.list[[i]]$gmat[2,3] <- data[i,6]*data[i,7]*data[i,11]

    #Lower diagonal
    results.list[[i]]$gmat[2,1] <- results.list[[i]]$gmat[1,2]
    results.list[[i]]$gmat[3,1] <- results.list[[i]]$gmat[1,3]
    results.list[[i]]$gmat[3,2] <- results.list[[i]]$gmat[2,3]

  }

  return(results.list)
}
```

```{r}
# Extract posterior samples
bivar2.all.brm.nb3_post <- posterior_samples(bivar2.all.brm.nb3)
bivar2.all.brm.nb3_post <- as.mcmc(bivar2.all.brm.nb3_post)

sampled.gmat9 <- sample.gmat2(bivar2.all.brm.nb3_post, replicates = 6000) 

sgmat9 <- lapply(sampled.gmat9, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat9 <- unname(sapply(sgmat9, '[[', 1)) #Change to matrix

sgmat9 <- t(sgmat9)

P.modelBV_RR9 <- sgmat9
P.modelBV_RR9.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR9.mode[k] <- posterior.mode(mcmc(sgmat9[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR9 <- sgmat9[,c(3,6)]
colnames(S.modelBV_RR9) <- c("S_intercepts", "S_slopes")
S.modelBV_RR9.mode <- P.modelBV_RR9.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR9))
HPDinterval(mcmc(S.modelBV_RR9))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR9 <- matrix(NA, nrow(S.modelBV_RR9) ,2)

for (i in 1:nrow(S.modelBV_RR9)) {
  P3_9 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_9[k] <- P.modelBV_RR9[i, k] }  
  P2_9 <- P3_9[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S9 <- P3_9[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR9[i,] <- solve(P2_9) %*% S9   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR9) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR9))
HPDinterval(mcmc(beta_post_RR9))
```

### Table 2A ZINB

```{r, eval=FALSE, include=TRUE}
bivar2.all.brm.nb3_zi<-brm(bf_FFD+bf_fitness_shoot, 
                       family = c(gaussian,zero_inflated_negbinomial), 
                       data = datadef3,warmup = 1000,iter = 6000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99))
```

```{r}
print(bivar2.all.brm.nb3_zi,digits=3)
```

#### Extract selection coefficients

```{r}
# Extract posterior samples
bivar2.all.brm.nb3_zi_post <- posterior_samples(bivar2.all.brm.nb3_zi)
bivar2.all.brm.nb3_zi_post <- as.mcmc(bivar2.all.brm.nb3_zi_post)

sampled.gmat9_zi <- sample.gmat2(bivar2.all.brm.nb3_zi_post, replicates = 6000) 

sgmat9_zi <- lapply(sampled.gmat9_zi, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat9_zi <- unname(sapply(sgmat9_zi, '[[', 1)) #Change to matrix

sgmat9_zi <- t(sgmat9_zi)

P.modelBV_RR9_zi <- sgmat9_zi
P.modelBV_RR9_zi.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR9_zi.mode[k] <- posterior.mode(mcmc(sgmat9_zi[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR9_zi <- sgmat9_zi[,c(3,6)]
colnames(S.modelBV_RR9_zi) <- c("S_intercepts", "S_slopes")
S.modelBV_RR9_zi.mode <- P.modelBV_RR9_zi.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR9_zi))
HPDinterval(mcmc(S.modelBV_RR9_zi))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR9_zi <- matrix(NA, nrow(S.modelBV_RR9_zi) ,2)

for (i in 1:nrow(S.modelBV_RR9_zi)) {
  P3_9_zi <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_9_zi[k] <- P.modelBV_RR9_zi[i, k] }  
  P2_9_zi <- P3_9_zi[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S9_zi <- P3_9_zi[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR9_zi[i,] <- solve(P2_9_zi) %*% S9_zi   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR9_zi) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR9_zi))
HPDinterval(mcmc(beta_post_RR9_zi))
```

### Table 2B

Mean fitness per year of study, with shoot volume.

Negative binomial distribution for fitness.

```{r message=FALSE, warning=FALSE}
bf_fitness_study_shoot <- bf(round(mean_fitness_study) ~  
                               cn_shoot_vol_mean_sqrt +
                         (1|ID1|id)) # Set up model formula
```

```{r, eval=FALSE, include=TRUE}
bivar4.all.brm.nb3<-brm(bf_FFD+bf_fitness_study_shoot, 
                       family = c(gaussian,negbinomial), 
                       data = datadef3,warmup = 1000,iter = 6000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores)
```

```{r}
print(bivar4.all.brm.nb3,digits=3) 
```

Measure of fit: Bayesian R2

```{r}
bayes_R2(bivar4.all.brm.nb3)
```

#### Extract selection coefficients

```{r}
# Extract posterior samples
bivar4.all.brm.nb3_post <- posterior_samples(bivar4.all.brm.nb3)
bivar4.all.brm.nb3_post <- as.mcmc(bivar4.all.brm.nb3_post)

sampled.gmat10 <- sample.gmat2(bivar4.all.brm.nb3_post, replicates = 6000) 

sgmat10 <- lapply(sampled.gmat10, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat10 <- unname(sapply(sgmat10, '[[', 1)) #Change to matrix

sgmat10 <- t(sgmat10)

P.modelBV_RR10 <- sgmat10
P.modelBV_RR10.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR10.mode[k] <- posterior.mode(mcmc(sgmat10[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR10 <- sgmat10[,c(3,6)]
colnames(S.modelBV_RR10) <- c("S_intercepts", "S_slopes")
S.modelBV_RR10.mode <- P.modelBV_RR10.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR10))
HPDinterval(mcmc(S.modelBV_RR10))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR10 <- matrix(NA, nrow(S.modelBV_RR10) ,2)

for (i in 1:nrow(S.modelBV_RR10)) {
  P3_10 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_10[k] <- P.modelBV_RR10[i, k] }  
  P2_10 <- P3_10[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S10 <- P3_10[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR10[i,] <- solve(P2_10) %*% S10   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR10) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR10))
HPDinterval(mcmc(beta_post_RR10))
```

### Figure: Selection differentials and gradients

```{r}
selcoefs<-rbind(
  # mean_fitness_fl, with shoot volume
  rbind(cbind(data.frame(coef=posterior.mode(mcmc(S.modelBV_RR9))),
              data.frame(HPDinterval(mcmc(S.modelBV_RR9)))),
        cbind(data.frame(coef=posterior.mode(mcmc(beta_post_RR9))),
              data.frame(HPDinterval(mcmc(beta_post_RR9)))))%>%
    mutate(fitness_meas="mean_fitness_fl",condition="yes",
           type=c("Differentials","Differentials",
                  "Gradients","Gradients"),
           param=c("Intercept","Slope","Intercept","Slope")),
  # mean_fitness_study, with shoot volume
  rbind(cbind(data.frame(coef=posterior.mode(mcmc(S.modelBV_RR10))),
              data.frame(HPDinterval(mcmc(S.modelBV_RR10)))),
        cbind(data.frame(coef=posterior.mode(mcmc(beta_post_RR10))),
              data.frame(HPDinterval(mcmc(beta_post_RR10)))))%>%
    mutate(fitness_meas="mean_fitness_study",condition="yes",
           type=c("Differentials","Differentials",
                  "Gradients","Gradients"),
           param=c("Intercept","Slope","Intercept","Slope"))
  )
```

```{r}
ggplot(selcoefs,
       aes(x=param,y=coef,color=fitness_meas))+
  geom_errorbar(aes(ymin=lower, ymax=upper),
                width=.2,position=position_dodge(.3))+
  geom_point(size=2,position=position_dodge(.3))+
  facet_wrap(~type)+
  geom_hline(yintercept=0,lty=3)+
  scale_color_manual(values=c("black","darkgrey"))+
  scale_y_continuous(breaks=c(-2,-1,0,1,2,3,4))+
  my_theme()+xlab(NULL)+ylab("Coefficient value")
# black: mean_fitness_fl, grey: mean_fitness_study
```

## Hypothesis 3

```{r}
datadef3_BLUPs<-read.csv("data/datadef3_BLUPs.csv") 
head(datadef3_BLUPs)
```

## Model in Table 3

Calculate residuals of linear model of slope of the RN on the intercept of the RN, and use these residuals as explanatory variable.

```{r}
model_residuals3<-lm(slope~intercept,datadef3_BLUPs)
summary(model_residuals3)
datadef3_BLUPs$slope_residuals<-residuals(model_residuals3)
hist(datadef3_BLUPs$slope_residuals)
```

```{r eval=FALSE, include=FALSE}
# Models with zero-inflation
modelBLUP_8_3_brms3<-brm(bf(round(intactseed)~slope_residuals*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                           zi~slope_residuals*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef3_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.85))
```

```{r}
print(modelBLUP_8_3_brms3,digits=3)
```

```{r}
conditional_effects(modelBLUP_8_3_brms3,effects="slope_residuals:cmean_4",
                    dpar="zi")
```


# Session information

```{r include=FALSE}
sessionInfo()
```
