---
title: "Lathyrus ms2: Selection on reaction norms - multivariate modeling for phenotypic selection on plasticity 4 (Arnold et al. 2019 Phil. Trans. R. Soc. B)"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load data created in previous notebooks, include=FALSE}
load("C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_4yrs5.RData")
load("C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_4yrs_total2.RData")
load("C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_5yrs5.RData")
load("C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_5yrs_total2.RData")
```

```{r load packages, include=FALSE}
library(knitr)
library(pander)
library(tidyverse)
library(ggthemes)
library(broom)
library(gridExtra)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ggrepel)
library(car)
library(effects)
library(RColorBrewer)
library(MCMCglmm)
library(blme)
library(beepr)
```

```{r Define ggplot themes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r pander options, include=FALSE}
panderOptions('table.continues', '')
```

Repeating some models using within-subject centering (van de Pol & Wright 2009).

Within-subject centering of mean April temperature: We substract the id's mean value from each observation value, deriving a new predictor variable (cmean_4_ws) to use as a fixed effect that expresses only the within-subject (within-id) variation component. We then need to derive a second new fixed predictor variable (cmean_4_bs) to express only the between-subject (between-id) variation component, and this is simply the ids’ means.

```{r ws centering}
data_4yrs<-data_4yrs%>%
  group_by(id)%>%
  mutate(cmean_4_ws=mean_4-mean(mean_4),cmean_4_bs=mean(mean_4))
data_5yrs<-data_5yrs%>%
  group_by(id)%>%
  mutate(cmean_4_ws=mean_4-mean(mean_4),cmean_4_bs=mean(mean_4))
```

# Ids with 4 years of data, mean April temperature

```{r MCMCglmm models 1, warning=FALSE, message=FALSE}
# Scaling factor for MCMCglmm iterations
sc <- 100#0 # Increase this parameter for longer runs

priorUV2_RR_ws <- list(G = list(G1 = list(V = diag(1), nu = 1),  # other random effect (YEAR)
                             G2 = list(V = diag(2), nu = 1),
                             G3 = list(V = diag(2), nu = 1)),  
                        # ^ 2x2 variance-covariance matrix for var in slopes + intercepts
                    R = list(R1 = list(V = diag(1), nu = 2)))  

univar.FFD_RR_ws4 <- MCMCglmm(FFD ~ cmean_4_ws + cmean_4_bs,
                        random = ~year + us(1 + cmean_4_ws):id + us(1 + cmean_4_bs):id,
                        rcov = ~units,
                        data = data_4yrs,
                        prior = priorUV2_RR_ws, 
                        family = "gaussian",
                        nitt = 1100 * sc, thin = sc, burnin = 100 * sc, verbose = F)
summary(univar.FFD_RR_ws4)
```

# Ids with 5 years of data, mean April temperature

```{r MCMCglmm models 2, warning=FALSE, message=FALSE}
univar.FFD_RR_ws5 <- MCMCglmm(FFD ~ cmean_4_ws + cmean_4_bs,
                        random = ~year + us(1 + cmean_4_ws):id + us(1 + cmean_4_bs):id,
                        rcov = ~units,
                        data = data_5yrs,
                        prior = priorUV2_RR_ws, 
                        family = "gaussian",
                        nitt = 1100 * sc, thin = sc, burnin = 100 * sc, verbose = F)
summary(univar.FFD_RR_ws5)
```

What I understand from van de Pol & Wright 2009: We tested whether either the within-subject effect (cmean_4_ws) or the between-subject effect (cmean_4_bs) is itself significant. As the parameter estimates of these two effects do not seem to differ, we can say that the within- and between-subject effects are effectively the same.
