---
title: 'Lathyrus ms2: Compilation data 88-95'
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
```

# OLD PERIOD

Read data 88-95

Some problems fixed (see Excel file LV88-96Alicia\_AV.xlsx).

```{r}
data8895<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/data88-95.csv",
  header=T,sep=",",dec=".")  
```

```{r}
head(data8895)
str(data8895)
```

| Variable   | Years available |
|------------|-----------------|
| vol        | 88--95          |
| fcode      | 88-95           |
| fenol      | 88-95           |
| nfl        | 88-95           |
| fruit      | 88-95           |
| totseed    | 88-94           |
| intactseed | 88-94           |

Convert data from wide to long format

```{r}
data8895 <- gather(data8895, key=variable, value=data, vol88:fruit95,
                   factor_key=F)
head(data8895)
str(data8895)
```

Divide variable into "variable\_name" and "year"

```{r}
data8895<-separate(data8895, variable, 
                   into = c("name", "value"), "(?<=[a-z])(?=[0-9])")
names(data8895)<-c("id_nr","ruta","genet","variable_name","year","data")
head(data8895)
str(data8895)
```

In the table below, 1 means that the plant was flowering, 0 that it was not flowering (vegetative), and NA that no data is available for that year

```{r}
(flowering<-spread(data8895,key=variable_name,value=data)%>%
    select(id_nr:fcode)%>%
    spread(key=year,value=fcode)%>%
    arrange(ruta,genet))
```

In the table below, for each year, 1 means that there is a value of the variable for that year, and 0 means that there is no value (NA).

I assume that 0 values are true zeros! (so if the value of the variable is 0, the code will be 1 in the binary table).

```{r}
(variables<-spread(data8895,key=year,value=data)%>%
    rename(y88="88",y89="89",y90="90",y91="91",y92="92",
           y93="93",y94="94",y95="95")%>%
    mutate(y88=ifelse(is.na(y88),0,1),
           y89=ifelse(is.na(y89),0,1),
           y90=ifelse(is.na(y90),0,1),
           y91=ifelse(is.na(y91),0,1),
           y92=ifelse(is.na(y92),0,1),
           y93=ifelse(is.na(y93),0,1),
           y94=ifelse(is.na(y94),0,1),
           y95=ifelse(is.na(y95),0,1)))
```

Read data 87 and 96

In this data, there were no problems in the Excel file (cleandata87and96.xlsx).

```{r}
data87<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/cleandata87.csv",
  header=T,sep=",",dec=".") 
data96<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/cleandata96.csv",
  header=T,sep=",",dec=".") 
```

```{r}
head(data87)
head(data96)
```

Add data 87 and 96 to flowering

```{r}
flowering87<-data87[c(1:3,9)]
flowering96<-data96[c(1:3,10)]
names(flowering87)<-c("ruta87","genet87","id_nr","87")
names(flowering96)<-c("ruta","genet","id_nr","96")
(flowering<-flowering%>%
  full_join(flowering87)%>%
  full_join(flowering96)%>%
    relocate("87", .after = genet)%>%
    relocate("ruta87", .afer = genet)%>%
    relocate("genet87", .after = ruta87))
```

Add data 87 and 96 to variables

```{r}
variables<-variables%>%
  full_join(
    full_join(gather(data87, key=variable_name, value=y87, nfl:fcode,
                     factor_key=F),
              gather(data96, key=variable_name, value=y96, nfl:fcode,
                     factor_key=F))%>%
      mutate(y87=ifelse(is.na(y87),0,1),y96=ifelse(is.na(y96),0,1))%>%
      arrange(id_nr))
full_join(gather(data87, key=variable_name, value=y87, nfl:fcode,factor_key=F),
          gather(data96, key=variable_name, value=y96, nfl:fcode,factor_key=F))%>%
  mutate(y87=ifelse(is.na(y87),0,1),y96=ifelse(is.na(y96),0,1))%>%
  arrange(id_nr)
variables<-variables%>%
  relocate("ruta87", .after = genet)%>%
  relocate("genet87", .after = ruta87)%>%
  relocate("y87", .after = variable_name)
```

Create binary table for fitness

```{r}
fitness_bin_J<-subset(variables,variable_name=="intactseed")%>%
  select(id_nr,y87:y96)%>%
  arrange(id_nr)
fitness_bin_j<-write.csv(fitness_bin_J,file="fitness_bin_J.csv")
```


Calculate number of years with information on FFD (n\_years\_FFD), number of years of flowering (n\_years\_flowering), number of years with information on fitness, and number of flowering years with information on fitness for each plant

```{r}
n_years<-full_join(
  (subset(variables,variable_name=="fenol")%>%
     mutate(n_years_FFD = rowSums(.[7:16],na.rm=T))%>%
     select(id_nr,ruta,genet,ruta87,genet87,n_years_FFD)),
  flowering%>%
    mutate(n_years_flowering = rowSums(.[6:15],na.rm=T))%>%
    select(id_nr,ruta,genet,n_years_flowering))
n_years<-full_join(n_years,
                   (subset(variables,variable_name=="intactseed")%>%
     mutate(n_years_fitness = rowSums(.[7:16],na.rm=T))%>%
     select(id_nr,ruta,genet,ruta87,genet87,n_years_fitness)))
n_years<-full_join(n_years,
                   subset(data8796,
                          variable_name=="intactseed"|variable_name=="fcode")%>%
  pivot_wider(names_from="variable_name",values_from="data")%>%
  filter(fcode==1)%>% # Keep only flowering years
  pivot_wider(names_from="year",values_from="intactseed")%>%
  rename("y87"='87',"y88"='88',"y89"='89',"y90"='90',"y91"='91',
         "y92"='92',"y93"='93',"y94"='94',"y95"='95',"y96"='96')%>%
  mutate(y87=ifelse(is.na(y87),0,1),y88=ifelse(is.na(y88),0,1),
         y89=ifelse(is.na(y89),0,1),y90=ifelse(is.na(y90),0,1),
         y91=ifelse(is.na(y91),0,1),y92=ifelse(is.na(y92),0,1),
         y93=ifelse(is.na(y93),0,1),y94=ifelse(is.na(y94),0,1),
         y95=ifelse(is.na(y95),0,1),y96=ifelse(is.na(y96),0,1))%>%
  mutate(n_years_fl_fitness = rowSums(.[3:12],na.rm=T))%>%
  select(id_nr,n_years_fl_fitness)
)
```

All the data

```{r}
data8796<-rbind(
  (gather(data87, key=variable_name, value=data, nfl:fcode,
                   factor_key=F)%>%
  mutate(year=87)%>%
  select(id_nr,variable_name,data,year)),
  (data8895%>%
  select(id_nr,variable_name,data,year)),
  (gather(data96, key=variable_name, value=data, nfl:fcode,
                   factor_key=F)%>%
  mutate(year=96)%>%
  select(id_nr,variable_name,data,year))
)
(data8796<-data8796%>%arrange(id_nr,variable_name,year))
```

Number of years when each plant was recorded --\> calculate in Excel from flowering table

--\> See what to do with fitness data lacking in 1995. OK with what I calculated before?

# NEW PERIOD

Read data (got from mydata)

```{r}
data0617<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/data_new_period.csv",
  header=T,sep=",",dec=".")  
data0617$X<-NULL
head(data0617)
```

Read info about number of shoots and shoot volume (to see when plants were not flowering but were vegetative)

```{r}
nshoots_vol<-read.csv(
  "C:/Users/User/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/data/info_nshoots_vol_0617.csv",
  header=T,sep=",",dec=".")
head(nshoots_vol)
nshoots_vol$id_nr<-nshoots_vol$id
nshoots_vol$id<-NULL
```

```{r}
data0617<-data0617%>%
  full_join(subset(nshoots_vol,info_nshoots_vol==1))
data0617$status<-with(data0617,
                      ifelse(!is.na(FFD)|!is.na(n_fl)|!is.na(n_intact_seeds),1, 0))
```

In the table below, 1 means that the plant was flowering, 0 that it was not flowering (vegetative), and NA that no data is available for that year

```{r}
flowering_new<-data0617[c(1,10,12)]%>%
  pivot_wider(names_from="year",values_from="status")
```

```{r}
flowering_new<-flowering_new%>%
  mutate(n_years_flowering = rowSums(.[2:13],na.rm=T))
head(flowering_new)
```

```{r}
(variables_new<-subset(mydata,period=="new")%>%
   select(year,FFD,n_fl,shoot_vol,n_intact_seeds,id_nr)%>%
   pivot_longer(names_to="variable_name",values_to="data",FFD:n_intact_seeds)%>%
   pivot_wider(names_from="year",values_from="data")%>%
   rename(y06="2006",y07="2007",y08="2008",y09="2009",y10="2010",y11="2011",
          y12="2012",y13="2013",y14="2014",y15="2015",y16="2016",y17="2017")%>%
   mutate(y06=ifelse(is.na(y06),0,1),
          y07=ifelse(is.na(y07),0,1),
          y08=ifelse(is.na(y08),0,1),
          y09=ifelse(is.na(y09),0,1),
          y10=ifelse(is.na(y10),0,1),
          y11=ifelse(is.na(y11),0,1),
          y12=ifelse(is.na(y12),0,1),
          y13=ifelse(is.na(y13),0,1),
          y14=ifelse(is.na(y14),0,1),
          y15=ifelse(is.na(y15),0,1),
          y16=ifelse(is.na(y16),0,1),
          y17=ifelse(is.na(y17),0,1)))
```

```{r}
n_years_new<-full_join(
  subset(variables_new,variable_name=="FFD")%>%
     mutate(n_years_FFD = rowSums(.[3:14],na.rm=T))%>%
     select(id_nr,n_years_FFD),
  flowering_new[c(1,14)])
n_years_new<-full_join(n_years_new,
                       (subset(variables_new,variable_name=="n_intact_seeds")%>%
                          mutate(n_years_fitness = rowSums(.[3:14],na.rm=T))%>%
                          select(id_nr,n_years_fitness)))
n_years_new<-full_join(n_years_new,
                       data0617%>%
                       select(year,n_intact_seeds,id_nr,status)%>%
                         filter(status==1)%>% # Keep only flowering years
                         pivot_wider(names_from="year",values_from="n_intact_seeds")%>%
                         rename("y06"='2006',"y07"='2007',"y08"='2008',"y09"='2009',
                                "y10"='2010',"y11"='2011',"y12"='2012',"y13"='2013',
                                "y14"='2014',"y15"='2015',"y16"='2016',"y17"='2017')%>%
                         mutate(y06=ifelse(is.na(y06),0,1),y07=ifelse(is.na(y07),0,1),
                                y08=ifelse(is.na(y08),0,1),y09=ifelse(is.na(y09),0,1),
                                y10=ifelse(is.na(y10),0,1),y11=ifelse(is.na(y11),0,1),
                                y12=ifelse(is.na(y12),0,1),y13=ifelse(is.na(y13),0,1),
                                y14=ifelse(is.na(y14),0,1),y15=ifelse(is.na(y15),0,1),
                                y16=ifelse(is.na(y16),0,1),y17=ifelse(is.na(y17),0,1),)%>%
                         mutate(n_years_fl_fitness=rowSums(.[3:14],na.rm=T))%>%
                         select(id_nr,n_years_fl_fitness)
)
head(n_years_new)
```




