---
title: "Lathyrus ms2: selection on reaction norms for flowering time"
subtitle: "Models with all data performed with brms"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.width=4, fig.height=3) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
library(MCMCglmm)
library(brms)
library(beepr)
library(ggthemes)
library(knitr)
library(bayesplot)
library(tidybayes)
library(parallel)
library(future)
library(ggeffects)
library(arm)
library(cowplot)
library(broom)
```

```{r Define ggplot themes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r load models, include=FALSE}
# Load previously run models and stuff
load("output/large_objects/large_objects.RData")
```

# Read data and check ns

```{r}
datadef<-read.csv("data/datadef.csv") 
head(datadef)
```

Number of individuals in each period:

```{r}
length(with(subset(datadef,period=="old"),unique(id)))
length(with(subset(datadef,period=="new"),unique(id)))
```

Number of observations in each period:

```{r}
nrow(subset(datadef,period=="old"))
nrow(subset(datadef,period=="new"))
```

Number of cases with FFD in each period:

```{r}
nrow(subset(datadef,period=="old"&!is.na(FFD)))
nrow(subset(datadef,period=="new"&!is.na(FFD)))
```

# Data preparation

```{r data prep MCMCglmm, message=FALSE, warning=FALSE}
datadef_total<-datadef %>%
  group_by(id)%>%
  # Calculate mean fitness per year of study 
  # and mean fitness per flowering event
  summarise(mean_fitness_study=sum(intactseed,na.rm=T)/mean(n_years_study),
            mean_fitness_fl=sum(intactseed,na.rm=T)/mean(n_years_fl_fitness))%>%
   arrange(.,id) # Order by id

with(datadef_total,cor(mean_fitness_study,mean_fitness_fl))  # Highly corr (0.87)

# Calculate mean shoot volume for each id using values of shoot volume 
# for all ids/years (including flowering and non-flowering years)

shoot_vol_all_means<-datadef[c(1,3,10)]%>%
  group_by(id)%>%
  summarise(shoot_vol_mean=mean(shoot_vol,na.rm=T)) 
# Mean of all available values 

# Join shoot volume data
datadef_total<-datadef_total%>%left_join(shoot_vol_all_means)%>%
  left_join(unique(datadef[c(2,3,11)]))
head(datadef_total)
nrow(subset(datadef_total,is.na(shoot_vol_mean))) 
# 46 ids with no info on shoot volume

# Add first_yr to total data (this was needed for MCMCglmm models) 
datadef_total$first_yr<-ifelse(grepl("old",as.character(datadef_total$id)),
                               1987,2006)

# Using sqrt of mean shoot volume over all years when available, centred
datadef_total<-datadef_total%>%
  mutate(shoot_vol_mean_sqrt=sqrt(shoot_vol_mean),
         cn_shoot_vol_mean_sqrt=scale(shoot_vol_mean_sqrt,center=T,scale=F))
```

Compare distributions of mean fitness per year of study and mean fitness per flowering event between old and new periods:

```{r fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
ggplot(datadef_total,aes(x=mean_fitness_study))+
  geom_histogram(colour="black",fill="white",position="dodge")+
  facet_wrap(~period,scales="free_y")+
  geom_vline(data=plyr::ddply(datadef_total,"period",summarise,
                        mean_fitness_study.mean=mean(mean_fitness_study)),
             aes(xintercept=mean_fitness_study.mean),
             linetype="dashed", size=1, colour="red")
ggplot(datadef_total,aes(x=mean_fitness_fl))+
  geom_histogram(colour="black",fill="white",position="dodge")+
  facet_wrap(~period,scales="free_y")+
  geom_vline(data=plyr::ddply(datadef_total,"period",summarise,
                        mean_fitness_fl.mean=mean(mean_fitness_fl)),
             aes(xintercept=mean_fitness_fl.mean),
             linetype="dashed", size=1, colour="red")
```

Distributions and means of the two mean fitness measures are similar among the two periods.

# Univariate models

## FFD with random effect of year only

```{r}
my.cores <- detectCores()
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
univar.FFD_yearonly.all.brm<-brm(formula=FFD~cmean_4+(1|year),data=datadef,
                warmup = 1000,iter = 4000,thin=2,chains=4,
                # 4 chains, each with 4000 iterations
                inits = "random",seed = 12345,cores = my.cores)
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_yearonly.all.brm)
```

## FFD with random effects of year and individual-intercept

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
univar.FFD.all.brm<-brm(formula=FFD~cmean_4+(1|year)+(1|id),data=datadef,
                warmup = 1000,iter = 4000,thin=2,chains=4,
                # 4 chains, each with 4000 iterations
                inits = "random",seed = 12345,cores = my.cores)
# Total of 6000 post-warmup samples
```

```{r }
summary(univar.FFD.all.brm)
```

## Random regression for FFD, including random effects of individual slopes and covariance between intercept and slope

```{r, eval=FALSE, include=TRUE}
univar.FFD_RR.all.brm<-brm(formula=FFD~cmean_4+(1|year)+(cmean_4|id),
                           data=datadef,
                           warmup = 1000,iter = 4000,thin=2,chains=4,
                           inits = "random",seed = 12345,cores = my.cores,
                           sample_prior="yes")
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_RR.all.brm) 
```

### Graph model prediction for each id

```{r}
predict_id<-ggpredict(univar.FFD_RR.all.brm,
                          terms=c("cmean_4","id"),type="random")
predict_mean<-ggpredict(univar.FFD_RR.all.brm,terms=c("cmean_4"),type="random")
```


```{r}
ggplot()+
  geom_line(data=data.frame(predict_id),aes(x=x,y=predicted,group=group),
            color="grey",size=0.01,alpha=0.3)+
  geom_line(data=data.frame(predict_mean),aes(x=x,y=predicted),
            color="black",size=1)+
  my_theme()+
  xlab("April temperature (mean-centred)")+
  ylab("Predicted FFD")
```


## Compare models

This allows  us to test if adding among-individual variation in elevation and slope of the RN improves model fit. 

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm <- add_criterion(univar.FFD_yearonly.all.brm,"loo")
univar.FFD.all.brm <- add_criterion(univar.FFD.all.brm,"loo")
univar.FFD_RR.all.brm <- add_criterion(univar.FFD_RR.all.brm,"loo")
```

```{r}
loo_comp<-loo_compare(univar.FFD_yearonly.all.brm,univar.FFD.all.brm,
                      univar.FFD_RR.all.brm, criterion="loo")
```

```{r}
loo_comp
```

The results indicate that adding a random intercept (i.e. among-individual variation in RN elevation) improves the fit, and that adding a random slope (i.e. among-individual variation in RN slope) improves the fit even more. Thus, the random regression model is the best model.

## Model evaluation for random regression model

I used some code from https://biol609.github.io/lectures/23c_brms_prediction.html#24_evaluating_brms_models

```{r fig.height=8, fig.width=6}
plot(univar.FFD_RR.all.brm) 
```

Rhat (potential scale reduction statistic): monitors whether a chain has converged to the equilibrium distribution by comparing its behavior to other randomly initialized chains.

-   light: below 1.05 (good)

-   mid: between 1.05 and 1.1 (ok)

-   dark: above 1.1 (too high)

```{r}
mcmc_rhat_hist(rhat(univar.FFD_RR.all.brm))+theme_bw()
```

Effective sample size: an estimate of the number of independent draws from the posterior distribution of the estimand of interest.

-   light: between 0.5 and 1 (high)

-   mid: between 0.1 and 0.5 (good)

-   dark: below 0.1 (low)

We should worry about any values less than 0.1.

```{r}
mcmc_neff_hist(neff_ratio(univar.FFD_RR.all.brm))+theme_bw()
```

Everything seems OK till here.

Autocorrelation: Takes very long / gets stuck, see later.

```{r}
# mcmc_acf(univar.FFD_RR.all.brm,lags=20)
```

Assessing fit: 

Posterior predictive checks:  Compares observed data to simulated data from the posterior predictive distribution (if a model is a good fit, we should be able to use it to generate data that resemble the data that we observed).

I used some code from https://cran.r-project.org/web/packages/bayesplot/vignettes/graphical-ppcs.html#:~:text=MCMC%20diagnostics%20vignette.-,Graphical%20posterior%20predictive%20checks%20(PPCs),Gabry%20et%20al%2C%202019).

```{r}
y1<-subset(datadef,!is.na(FFD))$FFD # vector of outcome values
yrep3<-posterior_predict(univar.FFD_RR.all.brm, draws = 500)
# matrix of draws from the posterior predictive distribution
```

Each row of the matrix is a draw from the posterior predictive distribution, i.e. a vector with one element for each of the data points in y.

Comparison of the distribution of y and the distribution of the simulated dataset in the yrep matrix

```{r}
ppc_dens_overlay(y1, yrep3[1:500,])
```

Does this look good enough?

Separate histograms of y and some of the yrep datasets

```{r fig.height=6, fig.width=6}
ppc_hist(y1, yrep3[1:8, ])
```

Does this look good enough?

```{r message=FALSE, warning=FALSE}
ppc_stat(y1,yrep3,stat="median")
ppc_stat(y1,yrep3,stat="mean")
ppc_stat(y1,yrep3,stat="sd")
ppc_stat(y1,yrep3,stat="max")
ppc_stat(y1,yrep3,stat="min")
```

This looks quite OK (worst is for the min).

Measure of fit: Bayesian R2, which looks at the model expected variance / (expected variance + residual variance).

```{r}
bayes_R2(univar.FFD_RR.all.brm)
```

Leave-one-out cross validation (LOO): 

I used code from https://www.monicaalexander.com/posts/2020-28-02-bayes_viz/

We are interested in estimating the out-of-sample predictive accuracy at each point i, when all we have to fit the model is data that without point i. We want to estimate the leave-one-out (LOO) posterior predictive densities and a summary of these across all points, which is called the LOO expected log pointwise predictive density. The bigger the numbers, the better we are at predicting the left out point i.

The output mentions Pareto k estimates, which give an indication of how 'influential' each point is. The higher the value of k, the more influential the point. Values of k over 0.7 are not good, and suggest the need for model re-specification. 

```{r eval=FALSE, include=FALSE}
loo1 <- loo(univar.FFD_RR.all.brm, save_psis = TRUE)
# Not sure what the warming means
```

```{r}
plot(loo1)
```

We have 3 obs with k > 0.7 - how bad is this?

Is this OK?

## Extract BLUPs from random regression model

Code was checked by Piet.

```{r}
BLUPs_MCMC.all.brms  <- cbind(as.factor(c(1:837)),
                              as.data.frame(ranef(univar.FFD_RR.all.brm)$id)
                              [c(1:2,5:6)])
colnames(BLUPs_MCMC.all.brms) <- c("id", "intercept", "intercept_sd",
                                   "slope","slope_sd")
with(BLUPs_MCMC.all.brms,cor(intercept,slope)) # highly correlated!
```

Plots with BLUPs

```{r, echo=FALSE, fig.height=4, fig.width=10}
ggplot(BLUPs_MCMC.all.brms, aes(id, intercept)) + 
  geom_point(aes(group = id, colour = id), size = 4, alpha = 0.5) + 
  ylab("BLUP intercept estimate") +
  geom_hline(yintercept = 0, lty = 2) + my_theme() +
  theme(axis.text.x = element_blank())
```

```{r, echo=FALSE, fig.height=4, fig.width=10}
ggplot(BLUPs_MCMC.all.brms, aes(id, slope)) + 
  geom_point(aes(group = id, colour = id), size = 4, alpha = 0.5) + 
  ylab("Plasticity (BLUP slope estimate)") +
  geom_hline(yintercept = 0, lty = 2) + my_theme() +
  theme(axis.text.x = element_blank())
```

```{r, echo=FALSE, fig.height=4, fig.width=10}
BLUPs_MCMC.all.brms$id_ordered <- factor(BLUPs_MCMC.all.brms$id,
                                    levels = BLUPs_MCMC.all.brms$id[order(BLUPs_MCMC.all.brms$slope)])
ggplot(BLUPs_MCMC.all.brms, aes(id_ordered, slope)) +
  geom_bar(stat = "identity", aes(group = id, fill = id)) +
  xlab("Id (in ranked order of plasticity)") +
  ylab("Plasticity (BLUP slope estimate)") +
  my_theme() + theme(axis.text.x = element_blank())
```

Add BLUPs to data set

```{r}
datadef<-datadef%>%left_join(BLUPs_MCMC.all.brms%>%
                               dplyr::select(intercept,intercept_sd,
                                             slope,slope_sd)%>%
                               rownames_to_column(var="id"))
```

## Figure 1

```{r}
predict_id_2<-ggpredict(univar.FFD_RR.all.brm,
                      terms=c("cmean_4","id"),type="random")
predict_mean_2<-ggpredict(univar.FFD_RR.all.brm,terms=c("cmean_4"),
                          type="random")
```

```{r fig.height=10, fig.width=12}
data_fig1<-full_join(data.frame(predict_id_2),
          datadef%>%dplyr::select(id,slope)%>%group_by(id)%>%
            summarise(slope=mean(slope))%>%rename(group=id))
plot_grid(ggplot()+
  geom_line(data=subset(data_fig1,group!="new_46"|group!="new_131"),
            aes(x=x,y=predicted,group=group,color=slope),
            size=0.01,alpha=0.7)+
  scale_color_viridis(end=0.9,option = "C")+
  geom_line(data=subset(data_fig1,group=="new_46"|group=="new_131"),
            aes(x=x,y=predicted,group=group,color=slope),
            size=1,linetype=2)+
  geom_line(data=data.frame(predict_mean_2),aes(x=x,y=predicted),
            color="black",size=1)+
  geom_point(data=subset(datadef,id=="new_46"|id=="new_131"),
             aes(x=cmean_4,y=FFD,color=slope),size=3,alpha=1)+
  my_theme()+
  xlab("April temperature (mean-centered)")+
  ylab("Predicted FFD"),
          plot_grid(ggplot(BLUPs_MCMC.all.brms,
                           aes(x=intercept+fixef(univar.FFD_RR.all.brm)[1]))+
                      geom_histogram(fill="grey",color="black")+my_theme()+
                      geom_vline(xintercept=fixef(univar.FFD_RR.all.brm)[1],
                                 size=0.5,linetype="dashed")+
                      xlab("Elevation of the reaction norm")+
                      ylab("Number of\nindividuals"),
                    ggplot(BLUPs_MCMC.all.brms,
                           aes(x=slope+fixef(univar.FFD_RR.all.brm)[2]))+
                      geom_histogram(fill="grey",color="black")+my_theme()+
                      geom_vline(xintercept=fixef(univar.FFD_RR.all.brm)[2],
                                 size=0.5,linetype="dashed")+
                      xlab("Slope of the reaction norm")+
                      ylab("Number of\nindividuals"),
                    ncol=1,labels=c("B)","C)"),label_fontfamily="serif"),
          labels = c("A)",""),label_fontfamily="serif",rel_widths = c(2,1.5))
ggsave(filename="output/figures/fig1.tiff",
       device="tiff",width=22,height=12,units="cm",dpi=300,compression="lzw")
```


# Bivariate models

## 1. mean_fitness_fl, no condition variable

Using the ID-syntax to specify fitness to be correlated with the intercept and slope of FFD on temperature - code has been checked by Piet.

Regarding distributions, I tried Poisson distribution for fitness, but not sure how eventual overdispersion is handled. I also tried using a negative binomial distribution. Results seem quite similar.Maybe we should consider trying a zero-inflated negative binomial (I did some attempts but seems to take much longer!).

### Poisson distribution

```{r message=FALSE, warning=FALSE}
datadef<-left_join(datadef,datadef_total[c(1:3,9)]) 
# Add info on mean fitness and mean shoot volume
bf_FFD <- bf(FFD ~ cmean_4 + (cmean_4|ID1|id) + (1|year)) # Set up model formula
bf_fitness <- bf(round(mean_fitness_fl) ~  (1|ID1|id)) # Set up model formula
# Specifying group-level effects of the same grouping factor (id here) 
# to be correlated across formulas
# Expand the | operator into |<ID>|, where <ID> can be any value (ID1 here)
# Group-level terms with the same ID1 will be modeled as correlated 
# if they share same grouping factor(s)
```

Save data with BLUPs

```{r}
write_csv(datadef,file = "data/datadef_BLUPs.csv")
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
bivar1.all.brm.pois<-brm(bf_FFD + bf_fitness, family = c(gaussian, poisson), 
                         data = datadef,
                         warmup = 1000,iter = 4000,chains=4,thin=2,
                         inits = "random",seed = 12345,cores = my.cores,
                         save_all_pars=TRUE)
# Total of 6000 post-warmup samples
```

```{r}
summary(bivar1.all.brm.pois)
```

### Negative binomial distribution

```{r, eval=FALSE, include=TRUE}
bivar1.all.brm.nb<-brm(bf_FFD + bf_fitness, family = c(gaussian, negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                         inits = "random",seed = 12345,cores = my.cores,
                       save_all_pars=TRUE)
```

```{r}
print(bivar1.all.brm.nb,digits=3)
```

### Negative binomial distribution with zero-inflation

```{r, eval=FALSE, include=TRUE}
bivar1.all.brm.zinb<-brm(bf_FFD + bf_fitness, 
                         family = c(gaussian, zero_inflated_negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                         inits = "random",seed = 12345,cores = my.cores)
```

```{r}
summary(bivar1.all.brm.zinb)
```

### Model evaluation: Compare models

```{r fig.height=8, fig.width=6}
plot(bivar1.all.brm.pois) 
plot(bivar1.all.brm.nb)
plot(bivar1.all.brm.zinb)
```

Rhat

```{r}
mcmc_rhat_hist(rhat(bivar1.all.brm.pois))+theme_bw()
mcmc_rhat_hist(rhat(bivar1.all.brm.nb))+theme_bw()
mcmc_rhat_hist(rhat(bivar1.all.brm.zinb))+theme_bw()
```

Effective sample size: 

```{r}
mcmc_neff_hist(neff_ratio(bivar1.all.brm.pois))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar1.all.brm.nb))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar1.all.brm.zinb))+theme_bw()
```

Some (few) are < 0.1 - Do we need to worry about this?

Posterior predictive checks:  

```{r}
y2_fitness<-round(subset(datadef,!is.na(FFD))$mean_fitness_fl)
y2_FFD<-subset(datadef,!is.na(FFD))$FFD # vectors of outcome values
yrep2_fitness_pois<-posterior_predict(bivar1.all.brm.pois, 
                                 draws = 500,resp="roundmeanfitnessfl")
yrep2_FFD_pois<-posterior_predict(bivar1.all.brm.pois, 
                                 draws = 500,resp="FFD")
yrep2_fitness_nb<-posterior_predict(bivar1.all.brm.nb, 
                                 draws = 500,resp="roundmeanfitnessfl")
yrep2_FFD_nb<-posterior_predict(bivar1.all.brm.nb, 
                                 draws = 500,resp="FFD")
yrep2_fitness_zinb<-posterior_predict(bivar1.all.brm.zinb, 
                                 draws = 500,resp="roundmeanfitnessfl")
yrep2_FFD_zinb<-posterior_predict(bivar1.all.brm.zinb, 
                                 draws = 500,resp="FFD")
# matrices of draws from the posterior predictive distribution
```

Each row of the matrix is a draw from the posterior predictive distribution, i.e. a vector with one element for each of the data points in y.

Comparison of the distribution of y and the distributions of the simulated datasets in the yrep matrix

```{r}
ppc_dens_overlay(y2_fitness, yrep2_fitness_pois[1:500,])
ppc_dens_overlay(y2_fitness, yrep2_fitness_pois[1:500,])+xlim(0, 10)
ppc_dens_overlay(y2_fitness, yrep2_fitness_nb[1:500,])
ppc_dens_overlay(y2_fitness, yrep2_fitness_nb[1:500,])+xlim(0,10)
ppc_dens_overlay(y2_fitness, yrep2_fitness_zinb[1:500,])
ppc_dens_overlay(y2_fitness, yrep2_fitness_zinb[1:500,])+xlim(0,10)
ppc_dens_overlay(y2_FFD, yrep2_FFD_pois[1:500,])
ppc_dens_overlay(y2_FFD, yrep2_FFD_nb[1:500,])
ppc_dens_overlay(y2_FFD, yrep2_FFD_zinb[1:500,])
```

All look pretty similar.

Separate histograms of y and some of the yrep datasets

```{r fig.height=6, fig.width=6}
ppc_hist(y2_fitness, yrep2_fitness_pois[1:8, ])
ppc_hist(y2_fitness, yrep2_fitness_pois[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y2_fitness, yrep2_fitness_nb[1:8, ])
ppc_hist(y2_fitness, yrep2_fitness_nb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y2_fitness, yrep2_fitness_zinb[1:8, ])
ppc_hist(y2_fitness, yrep2_fitness_zinb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
```

Here, negative binomial looks a bit better for fitness.

```{r message=FALSE, warning=FALSE}
ppc_stat(y2_fitness, yrep2_fitness_pois,stat="median")
ppc_stat(y2_fitness, yrep2_fitness_nb,stat="median")
ppc_stat(y2_fitness, yrep2_fitness_zinb,stat="median")
ppc_stat(y2_fitness, yrep2_fitness_pois,stat="mean")
ppc_stat(y2_fitness, yrep2_fitness_nb,stat="mean")
ppc_stat(y2_fitness, yrep2_fitness_zinb,stat="mean")
ppc_stat(y2_fitness, yrep2_fitness_pois,stat="sd")
ppc_stat(y2_fitness, yrep2_fitness_nb,stat="sd")
ppc_stat(y2_fitness, yrep2_fitness_zinb,stat="sd")
ppc_stat(y2_fitness, yrep2_fitness_pois,stat="max")
ppc_stat(y2_fitness, yrep2_fitness_nb,stat="max")
ppc_stat(y2_fitness, yrep2_fitness_zinb,stat="max")
```

Look at the distribution of the proportion of zeros over the replicated datasets from the posterior predictive distribution in yrep and compare to the proportion of observed zeros in y.

```{r}
# Define a function that takes a vector as input
# and returns the proportion of zeros:
prop_zero <- function(x) mean(x == 0)
prop_zero(y2_fitness) # check proportion of zeros in y
```

```{r}
ppc_stat(y2_fitness, yrep2_fitness_pois, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y2_fitness, yrep2_fitness_nb, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y2_fitness, yrep2_fitness_zinb, stat = "prop_zero", binwidth = 0.005)
```

They look quite similar.

Measure of fit: Bayesian R2

```{r}
bayes_R2(bivar1.all.brm.pois)
bayes_R2(bivar1.all.brm.nb)
bayes_R2(bivar1.all.brm.zinb)
```

Very similar.

Widely Applicable Information Criterion (WAIC):

```{r}
waic1<-waic(bivar1.all.brm.pois,bivar1.all.brm.nb,bivar1.all.brm.zinb,compare=T)
```

```{r}
waic1
```

Suggests that poisson is a bit better, but not sure we can trust this (gave some warnings).

Leave-one-out cross validation (LOO): 

```{r eval=FALSE, include=FALSE}
loo2_pois <- loo(bivar1.all.brm.pois, save_psis = TRUE, moment_match = TRUE)
loo2_nb <- loo(bivar1.all.brm.nb, save_psis = TRUE, moment_match = TRUE)
bivar1.all.brm.zinb <- add_criterion(bivar1.all.brm.zinb, "loo",
                                     moment_match = TRUE)
loo2_zinb <- loo(bivar1.all.brm.zinb, save_psis = TRUE, moment_match = TRUE)
```

Some Pareto k diagnostic values are too high. See help('pareto-k-diagnostic') for details.
Found 263 observations with a pareto_k > 0.7 in model 'bivar1.all.brm.pois'. With this many problematic observations, it may be more appropriate to use 'kfold' with argument 'K = 10' to perform 10-fold cross-validation rather than LOO.

```{r}
loo_compare(loo2_pois,loo2_nb,loo2_zinb)
```

Suggests that poisson is a bit better, but not sure we can trust this (gave warnings about too high pareto k diagnostic values).

```{r}
plot(loo2_pois)
plot(loo2_nb)
plot(loo2_zinb)
```

LOO predictive checks

```{r}
# ppc_loo_pit_overlay(y2_fitness, yrep2_fitness_pois, 
#                     lw = weights(loo2_pois$psis_object))
# ppc_loo_pit_qq(y2_fitness, yrep2_fitness_pois, 
#                     lw = weights(loo2_pois$psis_object))
# ppc_loo_pit_qq(y2_fitness, yrep2_fitness_pois, 
#                     lw = weights(loo2_pois$psis_object),compare="normal")
# ppc_loo_pit_overlay(y2_fitness, yrep2_fitness_nb, 
#                     lw = weights(loo2_nb$psis_object))
# ppc_loo_pit_qq(y2_fitness, yrep2_fitness_nb, 
#                     lw = weights(loo2_nb$psis_object))
# ppc_loo_pit_qq(y2_fitness, yrep2_fitness_nb, 
#                     lw = weights(loo2_nb$psis_object),compare="normal")
# ppc_loo_pit_overlay(y2_fitness, yrep2_fitness_zinb, 
#                     lw = weights(loo2_zinb$psis_object))
# ppc_loo_pit_qq(y2_fitness, yrep2_fitness_zinb, 
#                     lw = weights(loo2_zinb$psis_object))
# ppc_loo_pit_qq(y2_fitness, yrep2_fitness_zinb, 
#                     lw = weights(loo2_zinb$psis_object),compare="normal")
```

All looking quite bad!

k-fold cross-validation (K=5):

```{r eval=FALSE, include=FALSE}
kfold1_pois<-kfold(bivar1.all.brm.pois,K=5,chains=1,cores=my.cores)
```


```{r eval=FALSE, include=FALSE}
kfold1_nb<-kfold(bivar1.all.brm.nb,K=5,chains=1,cores=my.cores)
#kfold1_zinb<-kfold(bivar1.all.brm.zinb,K=5,chains=1,cores=my.cores)
```

```{r}
loo_compare(kfold1_pois,kfold1_nb
            #,kfold1_zinb
            )
```

Suggests that poisson is a bit better, and I guess we can trust this one as it gave no warnings?.

Decide if using Poisson or negative binomial!
Not yet run for zero-inflated negative binomial.

### Prior predictive checks

Not sure what I am doing here! I would need help from Pieter. I know you have to sample the prior by using sample_prior=TRUE, but I don't have much idea about what kind of priors we could specify to improve the model, and if this is really needed or if according to the posterior predictive checks we can stay with the default priors.

Any help welcome!

```{r eval=FALSE, include=FALSE}
prior1<-c(set_prior("normal(0,1000)", class = "b"),
          set_prior("student_t(3, 58.6, 6.6)", class = "Intercept",resp="FFD"),
          set_prior("student_t(3, 1.4, 2.5)",class = "Intercept",
                    resp="roundmeanfitnessfl"))
bivar1.all.brm.pois_priors<-brm(bf_fitness + bf_FFD,
       family = c(poisson, gaussian), data = datadef,
       prior = prior1,
                         warmup = 1000,iter = 4000,chains=4,thin=2,
                         inits = "random",seed = 12345,cores = my.cores,
                         sample_prior="only")
```


```{r}
yrep2_fitness_pois_priors<-posterior_predict(bivar1.all.brm.pois_priors, 
                                 draws = 500,resp="roundmeanfitnessfl")
yrep2_FFD_pois_priors<-posterior_predict(bivar1.all.brm.pois_priors, 
                                 draws = 500,resp="FFD")
# matrices of draws from the posterior(prior??) predictive distribution
```

Comparison of the distribution of y and the distributions of the simulated datasets in the yrep matrix

```{r}
ppc_dens_overlay(y2_fitness, yrep2_fitness_pois_priors[1:500,])
ppc_dens_overlay(y2_fitness, yrep2_fitness_pois_priors[1:500,])+xlim(0, 10)
ppc_dens_overlay(y2_FFD, yrep2_FFD_pois_priors[1:500,])
```

### Extract selection coefficients

#### Poisson model

Using example code from Piet. I am not 100% sure about what I am doing here! This code would need to be revised.

```{r}
# Extract posterior samples
bivar1.all.brm.pois_post <- posterior_samples(bivar1.all.brm.pois)
bivar1.all.brm.pois_post <- as.mcmc(bivar1.all.brm.pois_post)
#head(bivar1.all.brm.pois_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

# Sampling G-matrices from the posterior to calculate intervals estimates
# Use all posterior samples (no need to sample a random subset)
# Result is list with G-matrix
sample.gmat1 <- function(data, replicates = 6000) {
  
  ##Initialize the results list (list of lists)
  foo <- list(gmat = matrix(rep(0,3*3), ncol = 3))
  results.list <- list()
  for(j in 1:replicates) { results.list[[j]] <- foo }
  
  for(i in 1:replicates) {
    diag(results.list[[i]]$gmat) <- data[i,4:6]^2 #Get the diagonal
    
    #Upper diagonal
    results.list[[i]]$gmat[1,2] <- data[i,4]*data[i,5]*data[i,8]
    results.list[[i]]$gmat[1,3] <- data[i,4]*data[i,6]*data[i,9]
    results.list[[i]]$gmat[2,3] <- data[i,5]*data[i,6]*data[i,10]
    
    #Lower diagonal
    results.list[[i]]$gmat[2,1] <- results.list[[i]]$gmat[1,2]
    results.list[[i]]$gmat[3,1] <- results.list[[i]]$gmat[1,3]
    results.list[[i]]$gmat[3,2] <- results.list[[i]]$gmat[2,3]
    
  }
  
  return(results.list)
}

sampled.gmat1 <- sample.gmat1(bivar1.all.brm.pois_post, replicates = 6000) 
sampled.gmat1[[2]]
```

```{r}
sgmat1 <- lapply(sampled.gmat1, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat1 <- unname(sapply(sgmat1, '[[', 1)) #Change to matrix
str(sgmat1)
```


```{r}
sgmat1 <- t(sgmat1)

P.modelBV_RR1 <- sgmat1
P.modelBV_RR1.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR1.mode[k] <- posterior.mode(mcmc(sgmat1[,k]))
P.modelBV_RR1.mode
```

```{r}
# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR1 <- sgmat1[,c(3,6)]
colnames(S.modelBV_RR1) <- c("S_intercepts", "S_slopes")
S.modelBV_RR1.mode <- P.modelBV_RR1.mode[1:2, 3]
S.modelBV_RR1.mode
```

```{r}
posterior.mode(mcmc(S.modelBV_RR1))
```

```{r}
HPDinterval(mcmc(S.modelBV_RR1))
```

```{r}
# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR1 <- matrix(NA, nrow(S.modelBV_RR1) ,2)

for (i in 1:nrow(S.modelBV_RR1)) {
  P3_1 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_1[k] <- P.modelBV_RR1[i, k] }  
  P2_1 <- P3_1[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S1 <- P3_1[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR1[i,] <- solve(P2_1) %*% S1   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR1) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR1))
```

```{r}
HPDinterval(mcmc(beta_post_RR1))
```

#### Negative binomial model

```{r}
# Extract posterior samples
bivar1.all.brm.nb_post <- posterior_samples(bivar1.all.brm.nb)
bivar1.all.brm.nb_post <- as.mcmc(bivar1.all.brm.nb_post)
#head(bivar1.all.brm.nb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat2 <- sample.gmat1(bivar1.all.brm.nb_post, replicates = 6000) 
sampled.gmat2[[2]]
```

```{r}
sgmat2 <- lapply(sampled.gmat2, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat2 <- unname(sapply(sgmat2, '[[', 1)) #Change to matrix
str(sgmat2)
```


```{r}
sgmat2 <- t(sgmat2)

P.modelBV_RR2 <- sgmat2
P.modelBV_RR2.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR2.mode[k] <- posterior.mode(mcmc(sgmat2[,k]))
P.modelBV_RR2.mode
```

```{r}
# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR2 <- sgmat2[,c(3,6)]
colnames(S.modelBV_RR2) <- c("S_intercepts", "S_slopes")
S.modelBV_RR2.mode <- P.modelBV_RR2.mode[1:2, 3]
S.modelBV_RR2.mode
```

```{r}
posterior.mode(mcmc(S.modelBV_RR2))
```

```{r}
HPDinterval(mcmc(S.modelBV_RR2))
```

```{r}
# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR2 <- matrix(NA, nrow(S.modelBV_RR2) ,2)

for (i in 1:nrow(S.modelBV_RR2)) {
  P3_2 <- matrix(rep(NA, 9), nrow = 3)  
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_2[k] <- P.modelBV_RR2[i, k] }  
  P2_2 <- P3_2[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S2 <- P3_2[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR2[i,] <- solve(P2_2) %*% S2   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR2) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR2))
```

```{r}
HPDinterval(mcmc(beta_post_RR2))
```

#### Negative binomial model with zero-inflation

```{r}
# Extract posterior samples
bivar1.all.brm.zinb_post <- posterior_samples(bivar1.all.brm.zinb)
bivar1.all.brm.zinb_post <- as.mcmc(bivar1.all.brm.zinb_post)
#head(bivar1.all.brm.zinb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat9 <- sample.gmat1(bivar1.all.brm.zinb_post, replicates = 2000) 
sampled.gmat9[[2]]

sgmat9 <- lapply(sampled.gmat9, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat9 <- unname(sapply(sgmat9, '[[', 1)) #Change to matrix
str(sgmat9)

sgmat9 <- t(sgmat9)

P.modelBV_RR9 <- sgmat9
P.modelBV_RR9.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR9.mode[k] <- posterior.mode(mcmc(sgmat9[,k]))
P.modelBV_RR9.mode

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR9 <- sgmat9[,c(3,6)]
colnames(S.modelBV_RR9) <- c("S_intercepts", "S_slopes")
S.modelBV_RR9.mode <- P.modelBV_RR9.mode[1:2, 3]
S.modelBV_RR9.mode

posterior.mode(mcmc(S.modelBV_RR9))

HPDinterval(mcmc(S.modelBV_RR9))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR9 <- matrix(NA, nrow(S.modelBV_RR9) ,2)

for (i in 1:nrow(S.modelBV_RR9)) {
  P3_9 <- matrix(rep(NA, 9), nrow = 3)  
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_9[k] <- P.modelBV_RR9[i, k] }  
  P2_9 <- P3_9[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S9 <- P3_9[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR9[i,] <- solve(P2_9) %*% S9   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR9) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR9))

HPDinterval(mcmc(beta_post_RR9))
```

## 2. mean_fitness_fl, with shoot volume

```{r message=FALSE, warning=FALSE}
bf_fitness_shoot <- bf(round(mean_fitness_fl) ~  cn_shoot_vol_mean_sqrt +
                         (1|ID1|id)) # Set up model formula
```

### Poisson distribution

```{r, eval=FALSE, include=TRUE}
bivar2.all.brm.pois<-brm(bf_FFD+bf_fitness_shoot, 
                         family = c(gaussian, poisson), 
                         data = datadef,warmup = 1000,iter = 6000,chains=4,thin=2,
                         inits = "random",seed = 12345,cores = my.cores,
                         control = list(adapt_delta = 0.99, max_treedepth = 17))
```

```{r}
#summary(bivar2.all.brm.pois)
```

Got warnings:

<!-- # 3: There were 1642 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 15. See -->
<!-- # http://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded  -->
<!-- # 4: There were 1 chains where the estimated Bayesian Fraction of Missing Information was low. See -->
<!-- # http://mc-stan.org/misc/warnings.html#bfmi-low  -->
<!-- # 5: Examine the pairs() plot to diagnose sampling problems -->
<!-- #  -->
<!-- # 6: The largest R-hat is 1.6, indicating chains have not mixed. -->
<!-- # Running the chains for more iterations may help. See -->
<!-- # http://mc-stan.org/misc/warnings.html#r-hat  -->
<!-- # 7: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. -->
<!-- # Running the chains for more iterations may help. See -->
<!-- # http://mc-stan.org/misc/warnings.html#bulk-ess  -->
<!-- # 8: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable. -->
<!-- # Running the chains for more iterations may help. See -->
<!-- # http://mc-stan.org/misc/warnings.html#tail-ess  -->

Due to these warnings, results of bivar2.all.brm.pois are probably not reliable!
I have been trying to increase max_treedepth and running more iterations, but this takes extremely long to run (a couple of days with these parameters). I am not sure it makes sense to try increasing max_treedepth and iterations even more, or if we should just stay with the negative binomial distribution, which works fine and gives no warnings.

### Negative binomial distribution

```{r, eval=FALSE, include=TRUE}
bivar2.all.brm.nb<-brm(bf_FFD+bf_fitness_shoot, 
                       family = c(gaussian,negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99))
```

```{r}
print(bivar2.all.brm.nb,digits=3)
```

### Negative binomial distribution with zero-inflation

```{r, eval=FALSE, include=TRUE}
bivar2.all.brm.zinb<-brm(bf_FFD+bf_fitness_shoot, 
                       family = c(gaussian,zero_inflated_negbinomial), 
                       data = datadef,warmup = 1000,iter = 2000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99))
```

```{r}
summary(bivar2.all.brm.zinb)
```

### Model evaluation: Compare models

Looking only at negative binomial models by now, as Poisson model gave warnings.

```{r fig.height=8, fig.width=6}
#plot(bivar2.all.brm.pois) 
plot(bivar2.all.brm.nb)
plot(bivar2.all.brm.zinb)
```

Rhat

```{r}
#mcmc_rhat_hist(rhat(bivar2.all.brm.pois))+theme_bw()
mcmc_rhat_hist(rhat(bivar2.all.brm.nb))+theme_bw()
mcmc_rhat_hist(rhat(bivar2.all.brm.zinb))+theme_bw()
```

Effective sample size: 
  
```{r}
#mcmc_neff_hist(neff_ratio(bivar2.all.brm.pois))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar2.all.brm.nb))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar2.all.brm.zinb))+theme_bw()
```

Posterior predictive checks:  
  
```{r}
y3_fitness<-round(subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$mean_fitness_fl)
y3_FFD<-subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$FFD 
# vectors of outcome values
#yrep3_fitness_pois<-posterior_predict(bivar2.all.brm.pois, 
#                                      draws = 500,resp="roundmeanfitnessfl")
#yrep3_FFD_pois<-posterior_predict(bivar2.all.brm.pois, 
#                                  draws = 500,resp="FFD")
yrep3_fitness_nb<-posterior_predict(bivar2.all.brm.nb, 
                                    draws = 500,resp="roundmeanfitnessfl")
yrep3_FFD_nb<-posterior_predict(bivar2.all.brm.nb, 
                                draws = 500,resp="FFD")
yrep3_fitness_zinb<-posterior_predict(bivar2.all.brm.zinb, 
                                    draws = 500,resp="roundmeanfitnessfl")
yrep3_FFD_zinb<-posterior_predict(bivar2.all.brm.zinb, 
                                draws = 500,resp="FFD")
# matrices of draws from the posterior predictive distribution
```

Each row of the matrix is a draw from the posterior predictive distribution, i.e. a vector with one element for each of the data points in y.

Comparison of the distribution of y and the distributions of the simulated datasets in the yrep matrix

```{r}
#ppc_dens_overlay(y3_fitness, yrep3_fitness_pois[1:500,])
#ppc_dens_overlay(y3_fitness, yrep3_fitness_pois[1:500,])+xlim(0, 10)
ppc_dens_overlay(y3_fitness, yrep3_fitness_nb[1:500,])
ppc_dens_overlay(y3_fitness, yrep3_fitness_nb[1:500,])+xlim(0,10)
ppc_dens_overlay(y3_fitness, yrep3_fitness_zinb[1:500,])
ppc_dens_overlay(y3_fitness, yrep3_fitness_zinb[1:500,])+xlim(0,10)
#ppc_dens_overlay(y3_FFD, yrep3_FFD_pois[1:500,])
ppc_dens_overlay(y3_FFD, yrep3_FFD_nb[1:500,])
ppc_dens_overlay(y3_FFD, yrep3_FFD_zinb[1:500,])
```

Separate histograms of y and some of the yrep datasets

```{r fig.height=6, fig.width=6}
#ppc_hist(y3_fitness, yrep3_fitness_pois[1:8, ])
#ppc_hist(y3_fitness, yrep3_fitness_pois[1:8, ])+
#  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y3_fitness, yrep3_fitness_nb[1:8, ])
ppc_hist(y3_fitness, yrep3_fitness_nb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y3_fitness, yrep3_fitness_zinb[1:8, ])
ppc_hist(y3_fitness, yrep3_fitness_zinb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
```

```{r message=FALSE, warning=FALSE}
#ppc_stat(y3_fitness, yrep3_fitness_pois,stat="median")
ppc_stat(y3_fitness, yrep3_fitness_nb,stat="median")
ppc_stat(y3_fitness, yrep3_fitness_zinb,stat="median")
#ppc_stat(y3_fitness, yrep3_fitness_pois,stat="mean")
ppc_stat(y3_fitness, yrep3_fitness_nb,stat="mean")
ppc_stat(y3_fitness, yrep3_fitness_zinb,stat="mean")
#ppc_stat(y3_fitness, yrep3_fitness_pois,stat="sd")
ppc_stat(y3_fitness, yrep3_fitness_nb,stat="sd")
ppc_stat(y3_fitness, yrep3_fitness_zinb,stat="sd")
#ppc_stat(y3_fitness, yrep3_fitness_pois,stat="max")
ppc_stat(y3_fitness, yrep3_fitness_nb,stat="max")
ppc_stat(y3_fitness, yrep3_fitness_zinb,stat="max")
```

Look at the distribution of the proportion of zeros over the replicated datasets from the posterior predictive distribution in yrep and compare to the proportion of observed zeros in y.

```{r}
#ppc_stat(y3_fitness, yrep3_fitness_pois, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y3_fitness, yrep3_fitness_nb, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y3_fitness, yrep3_fitness_zinb, stat = "prop_zero", binwidth = 0.005)
```

Measure of fit: Bayesian R2

```{r}
#bayes_R2(bivar2.all.brm.pois)
bayes_R2(bivar2.all.brm.nb)
bayes_R2(bivar2.all.brm.zinb)
```

### Extract selection coefficients

#### Poisson model (not run due to warnings)

```{r}
# # Extract posterior samples
# bivar2.all.brm.pois_post <- posterior_samples(bivar2.all.brm.pois)
# bivar2.all.brm.pois_post <- as.mcmc(bivar2.all.brm.pois_post)
# #head(bivar2.all.brm.pois_post)[,1:20]
# 
# # [,5] sd_id__FFD_Intercept
# # [,6] sd_id__FFD_cmean_4
# # [,7] sd_id__roundmeanfitnessfl_Intercept 
# # [,9] cor_id__FFD_Intercept__FFD_cmean_4
# # [,10] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# # [,11] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept
# 
sample.gmat2 <- function(data, replicates = 6000) {

  ##Initialize the results list (list of lists)
  foo <- list(gmat = matrix(rep(0,3*3), ncol = 3))
  results.list <- list()
  for(j in 1:replicates) { results.list[[j]] <- foo }

  for(i in 1:replicates) {
    diag(results.list[[i]]$gmat) <- data[i,5:7]^2 #Get the diagonal

    #Upper diagonal
    results.list[[i]]$gmat[1,2] <- data[i,5]*data[i,6]*data[i,9]
    results.list[[i]]$gmat[1,3] <- data[i,5]*data[i,7]*data[i,10]
    results.list[[i]]$gmat[2,3] <- data[i,6]*data[i,7]*data[i,11]

    #Lower diagonal
    results.list[[i]]$gmat[2,1] <- results.list[[i]]$gmat[1,2]
    results.list[[i]]$gmat[3,1] <- results.list[[i]]$gmat[1,3]
    results.list[[i]]$gmat[3,2] <- results.list[[i]]$gmat[2,3]

  }

  return(results.list)
}
# 
# sampled.gmat3 <- sample.gmat2(bivar2.all.brm.pois_post, replicates = 6000) 
# 
# sgmat3 <- lapply(sampled.gmat3, `[`, c('gmat')) #Get list 'gmat' from each list
# sgmat3 <- unname(sapply(sgmat3, '[[', 1)) #Change to matrix
# 
# sgmat3 <- t(sgmat3)
# 
# P.modelBV_RR3 <- sgmat3
# P.modelBV_RR3.mode <- matrix(1:9, nrow = 3)
# for (k in 1:9) P.modelBV_RR3.mode[k] <- posterior.mode(mcmc(sgmat3[,k]))
# 
# # Extract selection *differentials* (i.e. covariances) for intercept and slope:
# # and calculate posterior mode and credible intervals for each
# S.modelBV_RR3 <- sgmat3[,c(3,6)]
# colnames(S.modelBV_RR3) <- c("S_intercepts", "S_slopes")
# S.modelBV_RR3.mode <- P.modelBV_RR3.mode[1:2, 3]
# 
# posterior.mode(mcmc(S.modelBV_RR3))
# HPDinterval(mcmc(S.modelBV_RR3))
# 
# # Estimate selection gradients for intercept and slope (beta = S / P)
# # on each sample of posterior and extract their mode
# beta_post_RR3 <- matrix(NA, nrow(S.modelBV_RR3) ,2)
# 
# for (i in 1:nrow(S.modelBV_RR3)) {
#   P3_3 <- matrix(rep(NA, 9), nrow = 3) 
#   # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
#   for (k in 1:9) {P3_3[k] <- P.modelBV_RR3[i, k] }  
#   P2_3 <- P3_3[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
#   S3 <- P3_3[1:2, 3]   # selection differentials on traits (last column of P3)
#   beta_post_RR3[i,] <- solve(P2_3) %*% S3   # selection gradients beta = P^-1 * S
# }
# 
# colnames(beta_post_RR3) <- c("beta_intercepts", "beta_slopes")
# posterior.mode(mcmc(beta_post_RR3))
# HPDinterval(mcmc(beta_post_RR3))
```

#### Negative binomial model

```{r}
# Extract posterior samples
bivar2.all.brm.nb_post <- posterior_samples(bivar2.all.brm.nb)
bivar2.all.brm.nb_post <- as.mcmc(bivar2.all.brm.nb_post)
#head(bivar2.all.brm.nb_post)[,1:20]

# [,5] sd_id__FFD_Intercept
# [,6] sd_id__FFD_cmean_4
# [,7] sd_id__roundmeanfitnessfl_Intercept 
# [,9] cor_id__FFD_Intercept__FFD_cmean_4
# [,10] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,11] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat4 <- sample.gmat2(bivar2.all.brm.nb_post, replicates = 6000) 

sgmat4 <- lapply(sampled.gmat4, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat4 <- unname(sapply(sgmat4, '[[', 1)) #Change to matrix

sgmat4 <- t(sgmat4)

P.modelBV_RR4 <- sgmat4
P.modelBV_RR4.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR4.mode[k] <- posterior.mode(mcmc(sgmat4[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR4 <- sgmat4[,c(3,6)]
colnames(S.modelBV_RR4) <- c("S_intercepts", "S_slopes")
S.modelBV_RR4.mode <- P.modelBV_RR4.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR4))
HPDinterval(mcmc(S.modelBV_RR4))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR4 <- matrix(NA, nrow(S.modelBV_RR4) ,2)

for (i in 1:nrow(S.modelBV_RR4)) {
  P3_4 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_4[k] <- P.modelBV_RR4[i, k] }  
  P2_4 <- P3_4[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S4 <- P3_4[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR4[i,] <- solve(P2_4) %*% S4   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR4) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR4))
HPDinterval(mcmc(beta_post_RR4))
```

#### Negative binomial model with zero-inflation

```{r}
# Extract posterior samples
bivar2.all.brm.zinb_post <- posterior_samples(bivar2.all.brm.zinb)
bivar2.all.brm.zinb_post <- as.mcmc(bivar2.all.brm.zinb_post)
#head(bivar2.all.brm.zinb_post)[,1:20]

# [,5] sd_id__FFD_Intercept
# [,6] sd_id__FFD_cmean_4
# [,7] sd_id__roundmeanfitnessfl_Intercept 
# [,9] cor_id__FFD_Intercept__FFD_cmean_4
# [,10] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,11] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat10 <- sample.gmat2(bivar2.all.brm.zinb_post, replicates = 2000) 

sgmat10 <- lapply(sampled.gmat10, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat10 <- unname(sapply(sgmat10, '[[', 1)) #Change to matrix

sgmat10 <- t(sgmat10)

P.modelBV_RR10 <- sgmat10
P.modelBV_RR10.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR10.mode[k] <- posterior.mode(mcmc(sgmat10[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR10 <- sgmat10[,c(3,6)]
colnames(S.modelBV_RR10) <- c("S_intercepts", "S_slopes")
S.modelBV_RR10.mode <- P.modelBV_RR10.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR10))
HPDinterval(mcmc(S.modelBV_RR10))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR10 <- matrix(NA, nrow(S.modelBV_RR10) ,2)

for (i in 1:nrow(S.modelBV_RR10)) {
  P3_10 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_10[k] <- P.modelBV_RR10[i, k] }  
  P2_10 <- P3_10[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S10 <- P3_10[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR10[i,] <- solve(P2_10) %*% S10   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR10) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR10))
HPDinterval(mcmc(beta_post_RR10))
```

## 3. mean_fitness_study, no condition variable

```{r message=FALSE, warning=FALSE}
bf_fitness_study <- bf(round(mean_fitness_study) ~  (1|ID1|id)) 
# Set up model formula
```

### Poisson distribution

```{r, eval=FALSE, include=TRUE}
bivar3.all.brm.pois<-brm(bf_FFD+bf_fitness_study, 
                         family = c(gaussian,poisson), 
                         data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                         inits = "random",seed = 12345,cores = my.cores,
                         control = list(adapt_delta = 0.99))
```

```{r}
summary(bivar3.all.brm.pois)
```

### Negative binomial distribution

```{r, eval=FALSE, include=TRUE}
bivar3.all.brm.nb<-brm(bf_FFD+bf_fitness_study, 
                       family = c(gaussian,negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99, max_treedepth = 15))
```

```{r}
print(bivar3.all.brm.nb,digits=3)
```

### Negative binomial distribution with zero-inflation

```{r, eval=FALSE, include=TRUE}
bivar3.all.brm.zinb<-brm(bf_FFD+bf_fitness_study, 
                       family = c(gaussian,zero_inflated_negbinomial), 
                       data = datadef,warmup = 1000,iter = 2000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99))
```

```{r}
summary(bivar3.all.brm.zinb)
```

### Model evaluation: Compare models

```{r fig.height=8, fig.width=6}
plot(bivar3.all.brm.pois) 
plot(bivar3.all.brm.nb)
plot(bivar3.all.brm.zinb)
```

Rhat

```{r}
mcmc_rhat_hist(rhat(bivar3.all.brm.pois))+theme_bw()
mcmc_rhat_hist(rhat(bivar3.all.brm.nb))+theme_bw()
mcmc_rhat_hist(rhat(bivar3.all.brm.zinb))+theme_bw()
```

Effective sample size: 
  
```{r}
mcmc_neff_hist(neff_ratio(bivar3.all.brm.pois))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar3.all.brm.nb))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar3.all.brm.zinb))+theme_bw()
```

Posterior predictive checks:  
  
```{r}
y3_fitness<-round(subset(datadef,!is.na(FFD))$mean_fitness_study)
yrep4_fitness_pois<-posterior_predict(bivar3.all.brm.pois, 
                                      draws = 500,resp="roundmeanfitnessstudy")
yrep4_FFD_pois<-posterior_predict(bivar3.all.brm.pois, 
                                  draws = 500,resp="FFD")
yrep4_fitness_nb<-posterior_predict(bivar3.all.brm.nb, 
                                    draws = 500,resp="roundmeanfitnessstudy")
yrep4_FFD_nb<-posterior_predict(bivar3.all.brm.nb, 
                                draws = 500,resp="FFD")
yrep4_fitness_zinb<-posterior_predict(bivar3.all.brm.nb, 
                                    draws = 500,resp="roundmeanfitnessstudy")
yrep4_FFD_zinb<-posterior_predict(bivar3.all.brm.nb, 
                                draws = 500,resp="FFD")
# matrices of draws from the posterior predictive distribution
```

Each row of the matrix is a draw from the posterior predictive distribution, i.e. a vector with one element for each of the data points in y.

Comparison of the distribution of y and the distributions of the simulated datasets in the yrep matrix

```{r}
ppc_dens_overlay(y3_fitness, yrep4_fitness_pois[1:500,])
ppc_dens_overlay(y3_fitness, yrep4_fitness_pois[1:500,])+xlim(0, 10)
ppc_dens_overlay(y3_fitness, yrep4_fitness_nb[1:500,])
ppc_dens_overlay(y3_fitness, yrep4_fitness_nb[1:500,])+xlim(0,10)
ppc_dens_overlay(y3_fitness, yrep4_fitness_zinb[1:500,])
ppc_dens_overlay(y3_fitness, yrep4_fitness_zinb[1:500,])+xlim(0,10)
ppc_dens_overlay(y2_FFD, yrep4_FFD_pois[1:500,])
ppc_dens_overlay(y2_FFD, yrep4_FFD_nb[1:500,])
ppc_dens_overlay(y2_FFD, yrep4_FFD_zinb[1:500,])
```

Poisson and negative binomial look pretty similar.

Separate histograms of y and some of the yrep datasets

```{r fig.height=6, fig.width=6}
ppc_hist(y3_fitness, yrep4_fitness_pois[1:8, ])
ppc_hist(y3_fitness, yrep4_fitness_pois[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y3_fitness, yrep4_fitness_nb[1:8, ])
ppc_hist(y3_fitness, yrep4_fitness_nb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y3_fitness, yrep4_fitness_zinb[1:8, ])
ppc_hist(y3_fitness, yrep4_fitness_zinb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
```

Here, negative binomial looks a bit better for fitness.

```{r message=FALSE, warning=FALSE}
ppc_stat(y3_fitness, yrep4_fitness_pois,stat="median")
ppc_stat(y3_fitness, yrep4_fitness_nb,stat="median")
ppc_stat(y3_fitness, yrep4_fitness_zinb,stat="median")
ppc_stat(y3_fitness, yrep4_fitness_pois,stat="mean")
ppc_stat(y3_fitness, yrep4_fitness_nb,stat="mean")
ppc_stat(y3_fitness, yrep4_fitness_zinb,stat="mean")
ppc_stat(y3_fitness, yrep4_fitness_pois,stat="sd")
ppc_stat(y3_fitness, yrep4_fitness_nb,stat="sd")
ppc_stat(y3_fitness, yrep4_fitness_zinb,stat="sd")
ppc_stat(y3_fitness, yrep4_fitness_pois,stat="max")
ppc_stat(y3_fitness, yrep4_fitness_nb,stat="max")
ppc_stat(y3_fitness, yrep4_fitness_zinb,stat="max")
```

Look at the distribution of the proportion of zeros over the replicated datasets from the posterior predictive distribution in yrep and compare to the proportion of observed zeros in y.

```{r}
ppc_stat(y3_fitness, yrep4_fitness_pois, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y3_fitness, yrep4_fitness_nb, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y3_fitness, yrep4_fitness_zinb, stat = "prop_zero", binwidth = 0.005)
```

They look quite similar.

Measure of fit: Bayesian R2

```{r}
bayes_R2(bivar3.all.brm.pois)
bayes_R2(bivar3.all.brm.nb)
bayes_R2(bivar3.all.brm.zinb)
```

Very similar.

### Extract selection coefficients

#### Poisson model

```{r}
# Extract posterior samples
bivar3.all.brm.pois_post <- posterior_samples(bivar3.all.brm.pois)
bivar3.all.brm.pois_post <- as.mcmc(bivar3.all.brm.pois_post)
#head(bivar3.all.brm.pois_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat5 <- sample.gmat1(bivar3.all.brm.pois_post, replicates = 6000) 

sgmat5 <- lapply(sampled.gmat5, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat5 <- unname(sapply(sgmat5, '[[', 1)) #Change to matrix

sgmat5 <- t(sgmat5)

P.modelBV_RR5 <- sgmat5
P.modelBV_RR5.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR5.mode[k] <- posterior.mode(mcmc(sgmat5[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR5 <- sgmat5[,c(3,6)]
colnames(S.modelBV_RR5) <- c("S_intercepts", "S_slopes")
S.modelBV_RR5.mode <- P.modelBV_RR5.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR5))
HPDinterval(mcmc(S.modelBV_RR5))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR5 <- matrix(NA, nrow(S.modelBV_RR5) ,2)

for (i in 1:nrow(S.modelBV_RR5)) {
  P3_5 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_5[k] <- P.modelBV_RR5[i, k] }  
  P2_5 <- P3_5[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S5 <- P3_5[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR5[i,] <- solve(P2_5) %*% S5   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR5) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR5))
HPDinterval(mcmc(beta_post_RR5))
```

#### Negative binomial model

```{r}
# Extract posterior samples
bivar3.all.brm.nb_post <- posterior_samples(bivar3.all.brm.nb)
bivar3.all.brm.nb_post <- as.mcmc(bivar3.all.brm.nb_post)
#head(bivar3.all.brm.nb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat6 <- sample.gmat1(bivar3.all.brm.nb_post, replicates = 6000) 

sgmat6 <- lapply(sampled.gmat6, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat6 <- unname(sapply(sgmat6, '[[', 1)) #Change to matrix

sgmat6 <- t(sgmat6)

P.modelBV_RR6 <- sgmat6
P.modelBV_RR6.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR6.mode[k] <- posterior.mode(mcmc(sgmat6[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR6 <- sgmat6[,c(3,6)]
colnames(S.modelBV_RR6) <- c("S_intercepts", "S_slopes")
S.modelBV_RR6.mode <- P.modelBV_RR6.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR6))
HPDinterval(mcmc(S.modelBV_RR6))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR6 <- matrix(NA, nrow(S.modelBV_RR6) ,2)

for (i in 1:nrow(S.modelBV_RR6)) {
  P3_6 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_6[k] <- P.modelBV_RR6[i, k] }  
  P2_6 <- P3_6[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S6 <- P3_6[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR6[i,] <- solve(P2_6) %*% S6   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR6) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR6))
HPDinterval(mcmc(beta_post_RR6))
```

#### Negative binomial model with zero-inflation

```{r}
# Extract posterior samples
bivar3.all.brm.zinb_post <- posterior_samples(bivar3.all.brm.zinb)
bivar3.all.brm.zinb_post <- as.mcmc(bivar3.all.brm.zinb_post)
#head(bivar3.all.brm.zinb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat11 <- sample.gmat1(bivar3.all.brm.zinb_post, replicates = 2000) 

sgmat11 <- lapply(sampled.gmat11, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat11 <- unname(sapply(sgmat11, '[[', 1)) #Change to matrix

sgmat11 <- t(sgmat11)

P.modelBV_RR11 <- sgmat11
P.modelBV_RR11.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR11.mode[k] <- posterior.mode(mcmc(sgmat11[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR11 <- sgmat11[,c(3,6)]
colnames(S.modelBV_RR11) <- c("S_intercepts", "S_slopes")
S.modelBV_RR11.mode <- P.modelBV_RR11.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR11))
HPDinterval(mcmc(S.modelBV_RR11))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR11 <- matrix(NA, nrow(S.modelBV_RR11) ,2)

for (i in 1:nrow(S.modelBV_RR11)) {
  P3_11 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_11[k] <- P.modelBV_RR11[i, k] }  
  P2_11 <- P3_11[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S11 <- P3_11[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR11[i,] <- solve(P2_11) %*% S11   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR11) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR11))
HPDinterval(mcmc(beta_post_RR11))
```

## 4. mean_fitness_study, with shoot volume

```{r message=FALSE, warning=FALSE}
bf_fitness_study_shoot <- bf(round(mean_fitness_study) ~  
                               cn_shoot_vol_mean_sqrt +
                         (1|ID1|id)) # Set up model formula
```

### Poisson distribution

```{r, eval=FALSE, include=TRUE}
bivar4.all.brm.pois<-brm(bf_FFD+bf_fitness_study_shoot, 
                       family = c(gaussian,poisson), 
                       data = datadef,warmup = 2000,iter = 6000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99, max_treedepth = 15))
```

Got warnings:

<!-- # 3: There were 1973 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 17. See -->
<!-- # http://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded  -->
<!-- # 4: There were 1 chains where the estimated Bayesian Fraction of Missing Information was low. See -->
<!-- # http://mc-stan.org/misc/warnings.html#bfmi-low  -->
<!-- # 5: Examine the pairs() plot to diagnose sampling problems -->
<!-- #  -->
<!-- # 6: The largest R-hat is 1.6, indicating chains have not mixed. -->
<!-- # Running the chains for more iterations may help. See -->
<!-- # http://mc-stan.org/misc/warnings.html#r-hat  -->
<!-- # 7: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable. -->
<!-- # Running the chains for more iterations may help. See -->
<!-- # http://mc-stan.org/misc/warnings.html#bulk-ess  -->
<!-- # 8: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable. -->
<!-- # Running the chains for more iterations may help. See -->
<!-- # http://mc-stan.org/misc/warnings.html#tail-ess  -->

Due to these warnings, results of bivar4.all.brm.pois are probably not reliable!
I have been trying to increase max_treedepth and running more iterations, but this takes extremely long to run (a couple of days with these parameters). I am not sure it makes sense to try increasing max_treedepth and iterations even more, or if we should just stay with the negative binomial distribution, which works fine and gives no warnings.

```{r}
#summary(bivar4.all.brm.pois) 
```

### Negative binomial distribution

```{r, eval=FALSE, include=TRUE}
bivar4.all.brm.nb<-brm(bf_FFD+bf_fitness_study_shoot, 
                       family = c(gaussian,negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores)
```

```{r}
print(bivar4.all.brm.nb,digits=3) 
```

### Negative binomial distribution with zero-inflation

```{r, eval=FALSE, include=TRUE}
bivar4.all.brm.zinb<-brm(bf_FFD+bf_fitness_study_shoot, 
                       family = c(gaussian,zero_inflated_negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores)
```

```{r}
summary(bivar4.all.brm.zinb) 
```

### Model evaluation: Compare models

Looking only at negative binomial model by now, as Poisson model gave warnings.

```{r fig.height=8, fig.width=6}
#plot(bivar4.all.brm.pois) 
plot(bivar4.all.brm.nb)
plot(bivar4.all.brm.zinb)
```

Rhat

```{r}
#mcmc_rhat_hist(rhat(bivar4.all.brm.pois))+theme_bw()
mcmc_rhat_hist(rhat(bivar4.all.brm.nb))+theme_bw()
mcmc_rhat_hist(rhat(bivar4.all.brm.zinb))+theme_bw()
```

Effective sample size: 
  
```{r}
#mcmc_neff_hist(neff_ratio(bivar4.all.brm.pois))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar4.all.brm.nb))+theme_bw()
mcmc_neff_hist(neff_ratio(bivar4.all.brm.zinb))+theme_bw()
```

Posterior predictive checks:  
  
```{r}
y3_FFD<-subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$FFD 
y4_fitness<-round(subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$mean_fitness_study)
#yrep5_fitness_pois<-posterior_predict(bivar4.all.brm.pois, 
#                                      draws = 500,resp="roundmeanfitnessstudy")
#yrep5_FFD_pois<-posterior_predict(bivar4.all.brm.pois, 
#                                  draws = 500,resp="FFD")
yrep5_fitness_nb<-posterior_predict(bivar4.all.brm.nb, 
                                    draws = 500,resp="roundmeanfitnessstudy")
yrep5_FFD_nb<-posterior_predict(bivar4.all.brm.nb, 
                                draws = 500,resp="FFD")
yrep5_fitness_zinb<-posterior_predict(bivar4.all.brm.zinb, 
                                    draws = 500,resp="roundmeanfitnessstudy")
yrep5_FFD_zinb<-posterior_predict(bivar4.all.brm.zinb, 
                                draws = 500,resp="FFD")
# matrices of draws from the posterior predictive distribution
```

Each row of the matrix is a draw from the posterior predictive distribution, i.e. a vector with one element for each of the data points in y.

Comparison of the distribution of y and the distributions of the simulated datasets in the yrep matrix

```{r}
#ppc_dens_overlay(y4_fitness, yrep5_fitness_pois[1:500,])
#ppc_dens_overlay(y4_fitness, yrep5_fitness_pois[1:500,])+xlim(0, 10)
ppc_dens_overlay(y4_fitness, yrep5_fitness_nb[1:500,])
ppc_dens_overlay(y4_fitness, yrep5_fitness_nb[1:500,])+xlim(0,10)
ppc_dens_overlay(y4_fitness, yrep5_fitness_zinb[1:500,])
ppc_dens_overlay(y4_fitness, yrep5_fitness_zinb[1:500,])+xlim(0,10)
#ppc_dens_overlay(y3_FFD, yrep5_FFD_pois[1:500,])
ppc_dens_overlay(y3_FFD, yrep5_FFD_nb[1:500,])
ppc_dens_overlay(y3_FFD, yrep5_FFD_zinb[1:500,])
```

Separate histograms of y and some of the yrep datasets

```{r fig.height=6, fig.width=6}
#ppc_hist(y4_fitness, yrep5_fitness_pois[1:8, ])
#ppc_hist(y4_fitness, yrep5_fitness_pois[1:8, ])+
#  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y4_fitness, yrep5_fitness_nb[1:8, ])
ppc_hist(y4_fitness, yrep5_fitness_nb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
ppc_hist(y4_fitness, yrep5_fitness_zinb[1:8, ])
ppc_hist(y4_fitness, yrep5_fitness_zinb[1:8, ])+
  coord_cartesian(xlim = c(-1, 20))
```

```{r message=FALSE, warning=FALSE}
#ppc_stat(y4_fitness, yrep5_fitness_pois,stat="median")
ppc_stat(y4_fitness, yrep5_fitness_nb,stat="median")
ppc_stat(y4_fitness, yrep5_fitness_zinb,stat="median")
#ppc_stat(y4_fitness, yrep5_fitness_pois,stat="mean")
ppc_stat(y4_fitness, yrep5_fitness_nb,stat="mean")
ppc_stat(y4_fitness, yrep5_fitness_zinb,stat="mean")
#ppc_stat(y4_fitness, yrep5_fitness_pois,stat="sd")
ppc_stat(y4_fitness, yrep5_fitness_nb,stat="sd")
ppc_stat(y4_fitness, yrep5_fitness_zinb,stat="sd")
#ppc_stat(y4_fitness, yrep5_fitness_pois,stat="max")
ppc_stat(y4_fitness, yrep5_fitness_nb,stat="max")
ppc_stat(y4_fitness, yrep5_fitness_zinb,stat="max")
```

Look at the distribution of the proportion of zeros over the replicated datasets from the posterior predictive distribution in yrep and compare to the proportion of observed zeros in y.

```{r}
#ppc_stat(y4_fitness, yrep5_fitness_pois, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y4_fitness, yrep5_fitness_nb, stat = "prop_zero", binwidth = 0.005)
ppc_stat(y4_fitness, yrep5_fitness_zinb, stat = "prop_zero", binwidth = 0.005)
```

Measure of fit: Bayesian R2

```{r}
#bayes_R2(bivar4.all.brm.pois)
bayes_R2(bivar4.all.brm.nb)
bayes_R2(bivar4.all.brm.zinb)
```

### Extract selection coefficients

#### Poisson model (not run due to warnings)

```{r}
# # Extract posterior samples
# bivar4.all.brm.pois_post <- posterior_samples(bivar4.all.brm.pois)
# bivar4.all.brm.pois_post <- as.mcmc(bivar4.all.brm.pois_post)
# #head(bivar4.all.brm.pois_post)[,1:20]
# 
# # [,4] sd_id__FFD_Intercept
# # [,5] sd_id__FFD_cmean_4
# # [,6] sd_id__roundmeanfitnessfl_Intercept 
# # [,8] cor_id__FFD_Intercept__FFD_cmean_4
# # [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# # [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept
# 
# sampled.gmat7 <- sample.gmat2(bivar4.all.brm.pois_post, replicates = 6000) 
# 
# sgmat7 <- lapply(sampled.gmat7, `[`, c('gmat')) #Get list 'gmat' from each list
# sgmat7 <- unname(sapply(sgmat7, '[[', 1)) #Change to matrix
# 
# sgmat7 <- t(sgmat7)
# 
# P.modelBV_RR7 <- sgmat7
# P.modelBV_RR7.mode <- matrix(1:9, nrow = 3)
# for (k in 1:9) P.modelBV_RR7.mode[k] <- posterior.mode(mcmc(sgmat7[,k]))
# 
# # Extract selection *differentials* (i.e. covariances) for intercept and slope:
# # and calculate posterior mode and credible intervals for each
# S.modelBV_RR7 <- sgmat7[,c(3,6)]
# colnames(S.modelBV_RR7) <- c("S_intercepts", "S_slopes")
# S.modelBV_RR7.mode <- P.modelBV_RR7.mode[1:2, 3]
# 
# posterior.mode(mcmc(S.modelBV_RR7))
# HPDinterval(mcmc(S.modelBV_RR7))
# 
# # Estimate selection gradients for intercept and slope (beta = S / P)
# # on each sample of posterior and extract their mode
# beta_post_RR7 <- matrix(NA, nrow(S.modelBV_RR7) ,2)
# 
# for (i in 1:nrow(S.modelBV_RR7)) {
#   P3_7 <- matrix(rep(NA, 9), nrow = 3) 
#   # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
#   for (k in 1:9) {P3_7[k] <- P.modelBV_RR7[i, k] }  
#   P2_7 <- P3_7[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
#   S7 <- P3_7[1:2, 3]   # selection differentials on traits (last column of P3)
#   beta_post_RR7[i,] <- solve(P2_7) %*% S7   # selection gradients beta = P^-1 * S
# }
# 
# colnames(beta_post_RR7) <- c("beta_intercepts", "beta_slopes")
# posterior.mode(mcmc(beta_post_RR7))
# HPDinterval(mcmc(beta_post_RR7))
```

#### Negative binomial model

```{r}
# Extract posterior samples
bivar4.all.brm.nb_post <- posterior_samples(bivar4.all.brm.nb)
bivar4.all.brm.nb_post <- as.mcmc(bivar4.all.brm.nb_post)
#head(bivar4.all.brm.nb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat8 <- sample.gmat2(bivar4.all.brm.nb_post, replicates = 6000) 

sgmat8 <- lapply(sampled.gmat8, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat8 <- unname(sapply(sgmat8, '[[', 1)) #Change to matrix

sgmat8 <- t(sgmat8)

P.modelBV_RR8 <- sgmat8
P.modelBV_RR8.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR8.mode[k] <- posterior.mode(mcmc(sgmat8[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR8 <- sgmat8[,c(3,6)]
colnames(S.modelBV_RR8) <- c("S_intercepts", "S_slopes")
S.modelBV_RR8.mode <- P.modelBV_RR8.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR8))
HPDinterval(mcmc(S.modelBV_RR8))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR8 <- matrix(NA, nrow(S.modelBV_RR8) ,2)

for (i in 1:nrow(S.modelBV_RR8)) {
  P3_8 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_8[k] <- P.modelBV_RR8[i, k] }  
  P2_8 <- P3_8[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S8 <- P3_8[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR8[i,] <- solve(P2_8) %*% S8   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR8) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR8))
HPDinterval(mcmc(beta_post_RR8))
```

#### Negative binomial model with zero-inflation

```{r}
# Extract posterior samples
bivar4.all.brm.zinb_post <- posterior_samples(bivar4.all.brm.zinb)
bivar4.all.brm.zinb_post <- as.mcmc(bivar4.all.brm.zinb_post)
#head(bivar4.all.brm.zinb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat12 <- sample.gmat2(bivar4.all.brm.zinb_post, replicates = 6000) 

sgmat12 <- lapply(sampled.gmat12, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat12 <- unname(sapply(sgmat12, '[[', 1)) #Change to matrix

sgmat12 <- t(sgmat12)

P.modelBV_RR12 <- sgmat12
P.modelBV_RR12.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR12.mode[k] <- posterior.mode(mcmc(sgmat12[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR12 <- sgmat12[,c(3,6)]
colnames(S.modelBV_RR12) <- c("S_intercepts", "S_slopes")
S.modelBV_RR12.mode <- P.modelBV_RR12.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR12))
HPDinterval(mcmc(S.modelBV_RR12))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR12 <- matrix(NA, nrow(S.modelBV_RR12) ,2)

for (i in 1:nrow(S.modelBV_RR12)) {
  P3_12 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_12[k] <- P.modelBV_RR12[i, k] }  
  P2_12 <- P3_12[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S12 <- P3_12[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR12[i,] <- solve(P2_12) %*% S12   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR12) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR12))
HPDinterval(mcmc(beta_post_RR12))
```

## Figure 2: Selection differentials and gradients

```{r}
selcoefs<-rbind(
  # 1. mean_fitness_fl, with shoot volume
  rbind(cbind(data.frame(coef=posterior.mode(mcmc(S.modelBV_RR4))),
              data.frame(HPDinterval(mcmc(S.modelBV_RR4)))),
        cbind(data.frame(coef=posterior.mode(mcmc(beta_post_RR4))),
              data.frame(HPDinterval(mcmc(beta_post_RR4)))))%>%
    mutate(fitness_meas="mean_fitness_fl",condition="yes",
           type=c("Differentials","Differentials",
                  "Gradients","Gradients"),
           param=c("Intercept","Slope","Intercept","Slope")),
  # 2. mean_fitness_study, with shoot volume
  rbind(cbind(data.frame(coef=posterior.mode(mcmc(S.modelBV_RR8))),
              data.frame(HPDinterval(mcmc(S.modelBV_RR8)))),
        cbind(data.frame(coef=posterior.mode(mcmc(beta_post_RR8))),
              data.frame(HPDinterval(mcmc(beta_post_RR8)))))%>%
    mutate(fitness_meas="mean_fitness_study",condition="yes",
           type=c("Differentials","Differentials",
                  "Gradients","Gradients"),
           param=c("Intercept","Slope","Intercept","Slope"))
  )
```

```{r}
ggplot(selcoefs,
       aes(x=param,y=coef,color=fitness_meas))+
  geom_errorbar(aes(ymin=lower, ymax=upper),
                width=.2,position=position_dodge(.3))+
  geom_point(size=2,position=position_dodge(.3))+
  facet_wrap(~type)+
  geom_hline(yintercept=0,lty=3)+
  scale_color_manual(values=c("black","darkgrey"))+
  scale_y_continuous(breaks=c(-2,-1,0,1,2,3,4))+
  my_theme()+xlab(NULL)+ylab("Coefficient value")
# black: mean_fitness_fl, grey: mean_fitness_study
ggsave(filename="output/figures/fig2.tiff",
        device="tiff",width=9,height=8,units="cm",dpi=300,compression="lzw")
```

# Appendix S1

## Model in Table S1C

```{r}
color_scheme_set("viridis")
arrayS1C<-as.array(univar.FFD_RR.all.brm)
dimnames(arrayS1C)[[3]][1:7] <- c("Intercept","April temperature",
                                 "Individual: SD (intercept)",
                                 "Individual: SD (slope)",
                                 "Year: SD (intercept)",
                                 "Individual:\ncorrelation (intercept, slope)",
                                 "Sigma")

mcmc_trace(arrayS1C,pars=c("Intercept","April temperature",
                                 "Individual: SD (intercept)"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_1a.tiff",
        device="tiff",width=20,height=21,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS1C,pars=c("Individual: SD (slope)",
                                 "Year: SD (intercept)",
                                 "Individual:\ncorrelation (intercept, slope)",
                                 "Sigma"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_1b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
```

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(rhat(univar.FFD_RR.all.brm))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(univar.FFD_RR.all.brm))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_2.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

```{r}
ppc_dens_overlay(y1, yrep3[1:100,])+my_theme_legend()+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank())
ggsave(filename="output/figures/figS1_3.tiff",
       device="tiff",width=12,height=8,units="cm",dpi=300,compression="lzw")
```

## Model in Table S3A

```{r}
array1<-as.array(bivar2.all.brm.nb)
dimnames(array1)[[3]][1:12] <- 
  c("Intercept of FFD","Intercept of fitness","April temperature on FFD",
    "Vegetative size on fitness","Individual: SD\n(intercept of FFD)",
    "Individual: SD\n(slope of FFD)","Individual: SD\n(intercept of fitness)",
    "Year: SD\n(intercept of FFD)",
    "Individual: correlation\n(intercept of FFD,slope of FFD)",
    "Individual: correlation\n(intercept of FFD, intercept of fitness)",
    "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD")

mcmc_trace(array1,pars=c("Intercept of FFD","Intercept of fitness",
                          "April temperature on FFD",
                          "Vegetative size on fitness"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_4a.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(array1,pars=c("Individual: SD\n(intercept of FFD)",
                         "Individual: SD\n(slope of FFD)",
                         "Individual: SD\n(intercept of fitness)",
                         "Year: SD\n(intercept of FFD)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_4b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(array1,pars=c(
  "Individual: correlation\n(intercept of FFD,slope of FFD)",
  "Individual: correlation\n(intercept of FFD, intercept of fitness)",
  "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD"),
  facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_4c.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
```

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(rhat(bivar2.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(bivar2.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_5.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

```{r}
plot_grid(ppc_dens_overlay(y3_fitness, yrep3_fitness_nb[1:100,])+
            my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          ppc_dens_overlay(y3_FFD, yrep3_FFD_nb[1:100,])+my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_6.tiff",
       device="tiff",width=24,height=10,units="cm",dpi=300,compression="lzw")
```

## Model in Table S3B

```{r}
arrayS3B<-as.array(bivar4.all.brm.nb)
dimnames(arrayS3B)[[3]][1:12] <- 
  c("Intercept of FFD","Intercept of fitness","April temperature on FFD",
    "Vegetative size on fitness",
    "Individual: SD\n(intercept of FFD)",
    "Individual: SD\n(slope of FFD)","Individual: SD\n(intercept of fitness)",
    "Year: SD\n(intercept of FFD)",
    "Individual: correlation\n(intercept of FFD,slope of FFD)",
    "Individual: correlation\n(intercept of FFD, intercept of fitness)",
    "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD")
mcmc_trace(arrayS3B,pars=c("Intercept of FFD","Intercept of fitness",
                          "April temperature on FFD",
                          "Vegetative size on fitness"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_7a.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS3B,pars=c(
   "Individual: SD\n(intercept of FFD)",
    "Individual: SD\n(slope of FFD)","Individual: SD\n(intercept of fitness)",
    "Year: SD\n(intercept of FFD)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_7b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS3B,pars=c(
  "Individual: correlation\n(intercept of FFD,slope of FFD)",
  "Individual: correlation\n(intercept of FFD, intercept of fitness)",
  "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD"),
  facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_7c.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
```

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(rhat(bivar4.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(bivar4.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_8.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

```{r}
plot_grid(ppc_dens_overlay(y4_fitness, yrep5_fitness_nb[1:100,])+
            my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          ppc_dens_overlay(y3_FFD, yrep5_FFD_nb[1:100,])+my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_9.tiff",
       device="tiff",width=24,height=10,units="cm",dpi=300,compression="lzw")
```

# Fig. S2

```{r}
datadef%>%
  group_by(year)%>%
  summarise(sd_FFD=sd(FFD,na.rm=T),mean_4=mean(mean_4))%>%
  ggplot(aes(x=mean_4,y=sd_FFD))+geom_point(size=3,alpha=0.2)+
  geom_smooth(method="lm",color="black")+my_theme()+
  xlab("April temperature (ºC)")+ylab("SD of FFD")
datadef%>%
  group_by(year)%>%
  summarise(sd_FFD=sd(FFD,na.rm=T),mean_4=mean(mean_4))%>%
  do(tidy(lm(sd_FFD~mean_4,data=.)))
ggsave(filename="output/figures/figS2.tiff",device="tiff",
       width=10,height=9,units="cm",dpi=300,compression="lzw")
```


# Save large objects as .RData file

```{r eval=FALSE, include=FALSE}
save(univar.FFD_yearonly.all.brm,univar.FFD.all.brm,univar.FFD_RR.all.brm,loo1,
     bivar1.all.brm.pois,bivar1.all.brm.nb,bivar1.all.brm.zinb,
     loo2_pois,loo2_nb,loo2_zinb,kfold1_pois,kfold1_nb,bivar1.all.brm.pois_priors,
     bivar2.all.brm.pois,bivar2.all.brm.nb,bivar2.all.brm.zinb,
     bivar3.all.brm.pois,bivar3.all.brm.nb,bivar3.all.brm.zinb,
     bivar4.all.brm.pois,bivar4.all.brm.nb,bivar4.all.brm.zinb,
     file = "output/large_objects.RData")
```

# 
```{r include=FALSE}
sessionInfo()
```
