---
title: "Lathyrus ms2: selection on reaction norms for flowering time"
subtitle: "Variation in selection among years using BLUPs - brms models"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
library(beepr)
library(ggthemes)
library(ggeffects)
library(glmmTMB)
library(knitr)
library(parallel)
library(brms)
library(bayesplot)
library(viridis)
library(cowplot)
library(grid)
library(gridExtra)
```

```{r Define ggplot themes and stuff, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r load models, include=FALSE}
# Load previously run models
load("output/large_objects_BLUPs_brms.RData")
```

# Read data and check ns

```{r}
datadef_BLUPs<-read.csv("data/datadef_BLUPs.csv") 
head(datadef_BLUPs)
```

```{r}
ggplot(datadef_BLUPs,aes(x=intactseed))+geom_histogram()
```

# Models with brms

## Temp*slope

Does selection on plasticity depend on temperature?

Trying different distributions and models with and without zero-inflation

```{r}
my.cores <- detectCores()
```

```{r eval=FALSE, include=FALSE}
# Models without zero-inflation
modelBLUP_1_1_brms<-brm(formula=round(intactseed)~slope*cmean_4+(1|id),
                        family="poisson",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
modelBLUP_1_2_brms<-brm(formula=round(intactseed)~slope*cmean_4+(1|id),
                        family="negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.99))
# Model with zero-inflation
modelBLUP_1_3_brms<-brm(bf(round(intactseed)~slope*cmean_4+(1|id),
                        zi~slope*cmean_4+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
```

### Posterior predictive checks

```{r}
y1<-round(subset(datadef_BLUPs,
           !is.na(intactseed))$intactseed) # vector of outcome values
yrep1_1<-posterior_predict(modelBLUP_1_1_brms, draws = 500)
yrep1_2<-posterior_predict(modelBLUP_1_2_brms, draws = 500)
yrep1_3<-posterior_predict(modelBLUP_1_3_brms, draws = 500)
# matrices of draws from the posterior predictive distribution
```

```{r}
ppc_dens_overlay(y1, yrep1_1[1:500,])
ppc_dens_overlay(y1, yrep1_2[1:500,])
ppc_dens_overlay(y1, yrep1_3[1:500,])
ppc_dens_overlay(y1, yrep1_1[1:500,])+xlim(0, 10)
ppc_dens_overlay(y1, yrep1_2[1:500,])+xlim(0, 10)
ppc_dens_overlay(y1, yrep1_3[1:500,])+xlim(0, 10)
```

Zero-inflated negative binomial looks much better.

### Model summary and predictions

```{r}
summary(modelBLUP_1_3_brms)
```

Predicted effects:

```{r}
modelBLUP_1_3_brms_plot1<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_1_3_brms, 
                         effects = "slope:cmean_4",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`),
       aes(x=slope,y=estimate__, group=cmean_4))+
  geom_ribbon(aes(ymin=lower__,ymax=upper__,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
legend1 <- get_legend(modelBLUP_1_3_brms_plot1)
modelBLUP_1_3_brms_plot1<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_1_3_brms, 
                         effects = "slope:cmean_4",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`),
       aes(x=slope,y=estimate__, group=cmean_4))+
  geom_ribbon(aes(ymin=lower__,ymax=upper__,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
modelBLUP_1_3_brms_plot2<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_1_3_brms, 
                         effects = "slope:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`),
       aes(x=slope,y=estimate__, group=cmean_4))+
  geom_ribbon(aes(ymin=lower__,ymax=upper__,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted zero-inflation probability of fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

```{r fig.height=7, fig.width=17}
ggdraw(plot_grid(plot_grid(modelBLUP_1_3_brms_plot1, modelBLUP_1_3_brms_plot2,
                           ncol=2, align='h',labels=c("A)","B)"),
                           label_fontfamily="serif"),
                 plot_grid(legend1,NULL,ncol=1),rel_widths=c(1, 0.2)))
ggsave(filename="output/figures/fig3.tiff",device="tiff",
       width=34,height=14,units="cm",dpi=300,compression="lzw")
```

Selection for plasticity (more negative slopes) increases with temperature, but this is only driven by the zero-inflated part of the model, i.e. the probability of fitness being zero decreases more with higher plasticity (more negative slopes) with warm temperatures.

## Temp*slope+volume

Does selection on plasticity depend on temperature when considering also plant size?

Trying different distributions and models with and without zero-inflation

```{r eval=FALSE, include=FALSE}
# Models without zero-inflation
modelBLUP_2_1_brms<-brm(formula=round(intactseed)~slope*cmean_4+
                          cn_shoot_vol_mean_sqrt+(1|id),
                        family="poisson",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
modelBLUP_2_2_brms<-brm(formula=round(intactseed)~slope*cmean_4+
                          cn_shoot_vol_mean_sqrt+(1|id),
                        family="negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
# Models with zero-inflation
modelBLUP_2_3_brms<-brm(bf(round(intactseed)~slope*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                        zi~slope*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
# Model with measurement error
modelBLUP_2_4_brms<-brm(bf(round(intactseed)~me(slope,slope_sd)*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                        zi~me(slope,slope_sd)*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id))+set_mecor(FALSE),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 5000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.99,max_treedepth = 15))
# By default, latent noise-free variables are assumed to be correlated. 
# To change that, add set_mecor(FALSE) to your model formula object 

# slope_sd is the standard deviation (measure of variability) 
# (not standard error!) of the slope BLUP
# slope is the mean (measure of central tendency) of the slope BLUP
# The mean and sd are calculated from all the posterior draws 
# (6000 in our case) using posterior_summary

# 3: There were 78 divergent transitions after warmup. See
# http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
# to find out why this is a problem and how to eliminate them. 
# 4: There were 1997 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 15. See
# http://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded 
# 5: Examine the pairs() plot to diagnose sampling problems
#  
# 6: The largest R-hat is 2.26, indicating chains have not mixed.
# Running the chains for more iterations may help. See
# http://mc-stan.org/misc/warnings.html#r-hat 
# 7: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
# Running the chains for more iterations may help. See
# http://mc-stan.org/misc/warnings.html#bulk-ess 
# 8: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
# Running the chains for more iterations may help. See
# http://mc-stan.org/misc/warnings.html#tail-ess 
```

### Posterior predictive checks

```{r}
y1<-round(subset(datadef_BLUPs,
           !is.na(intactseed)&!(is.na(cn_shoot_vol_mean_sqrt)))$intactseed) 
yrep2_1<-posterior_predict(modelBLUP_2_1_brms, draws = 500)
yrep2_2<-posterior_predict(modelBLUP_2_2_brms, draws = 500)
yrep2_3<-posterior_predict(modelBLUP_2_3_brms, draws = 500)
```

```{r}
ppc_dens_overlay(y1, yrep2_1[1:500,])
ppc_dens_overlay(y1, yrep2_2[1:500,])
ppc_dens_overlay(y1, yrep2_3[1:500,])
ppc_dens_overlay(y1, yrep2_1[1:500,])+xlim(0, 10)
ppc_dens_overlay(y1, yrep2_2[1:500,])+xlim(0, 10)
ppc_dens_overlay(y1, yrep2_3[1:500,])+xlim(0, 10)
```

### Model summary and predictions

```{r}
print(modelBLUP_2_3_brms,digits=3)
```

Predicted effects:

```{r}
modelBLUP_2_3_brms_plot1<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_2_3_brms, 
                         effects = "slope:cmean_4",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`),
       aes(x=slope,y=estimate__, group=cmean_4))+
  geom_ribbon(aes(ymin=lower__,ymax=upper__,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted count of fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
legend1 <- get_legend(modelBLUP_2_3_brms_plot1)
modelBLUP_2_3_brms_plot1<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_2_3_brms, 
                         effects = "slope:cmean_4",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`),
       aes(x=slope,y=estimate__, group=cmean_4))+
  geom_ribbon(aes(ymin=lower__,ymax=upper__,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted count of fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
modelBLUP_2_3_brms_plot2<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_2_3_brms, 
                         effects = "slope:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`),
       aes(x=slope,y=estimate__, group=cmean_4))+
  geom_ribbon(aes(ymin=lower__,ymax=upper__,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted zero-inflation probability of fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

```{r fig.height=7, fig.width=17}
ggdraw(plot_grid(plot_grid(modelBLUP_2_3_brms_plot1, modelBLUP_2_3_brms_plot2,
                           ncol=2, align='h',labels=c("A)","B)"),
                           label_fontfamily="serif"),
                 plot_grid(legend1,NULL,ncol=1),rel_widths=c(1, 0.2)))
```

Selection for plasticity (more negative slopes) increases with temperature, but this is only driven by the zero-inflated part of the model, i.e. the probability of fitness being zero decreases more with higher plasticity (more negative slopes) with warm temperatures. CAREFUL: CIs 0.002    0.262 (try with more iterations?)

```{r}
modelBLUP_2_3_brms_plot2<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_2_3_brms, 
                         effects = "slope:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=slope,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
ggdraw(plot_grid(modelBLUP_2_3_brms_plot2,
                 plot_grid(legend1,NULL,ncol=1),rel_widths=c(1, 0.4)))
```


## Year*slope+volume (not used)

Does selection on plasticity vary among years when considering also plant size?

```{r}
datadef_BLUPs$year_fct<-as.factor(datadef_BLUPs$year)
```

```{r eval=FALSE, include=FALSE}
modelBLUP_3_3_brms<-brm(bf(round(intactseed)~slope*year_fct+
                             cn_shoot_vol_mean_sqrt+(1|id),
                        zi~slope*year_fct+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 8000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
# Would need more iterations
```

## Temp*intercept+volume

```{r eval=FALSE, include=FALSE}
modelBLUP_4_3_brms<-brm(bf(round(intactseed)~intercept*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                        zi~intercept*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.99))
```

### Model summary and predictions

```{r}
print(modelBLUP_4_3_brms,digits=3)
```

Predicted effects:

```{r}
modelBLUP_4_3_brms_plot1<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_4_3_brms, 
                         effects = "intercept",dpar="zi")$`intercept`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=intercept,y=prob_fitness))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.2)+
  geom_line(size=1)+
  my_theme()+
  xlab("Intercept of the reaction norm")+
  ylab("Predicted probability of having any fitness")
modelBLUP_4_3_brms_plot2<-
  ggplot(as.data.frame(conditional_effects(modelBLUP_4_3_brms, 
                         effects = "cmean_4",dpar="zi")$`cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=cmean_4,y=prob_fitness))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.2)+
  geom_line(size=1)+
  my_theme()+
  xlab("April temperature (mean-centred)")+
  ylab("Predicted probability of having any fitness")
```

```{r fig.height=5, fig.width=12}
ggdraw(plot_grid(modelBLUP_4_3_brms_plot1, modelBLUP_4_3_brms_plot2,
                           ncol=2, align='h',labels=c("A)","B)"),
                           label_fontfamily="serif"))
ggsave(filename="output/figures/figS1.tiff",device="tiff",
       width=28,height=13,units="cm",dpi=300,compression="lzw")
```

## Temp*slope+intercept+volume

```{r eval=FALSE, include=FALSE}
# Models with zero-inflation
modelBLUP_5_3_brms<-brm(bf(round(intactseed)~slope*cmean_4+intercept+
                             cn_shoot_vol_mean_sqrt+(1|id),
                           zi~slope*cmean_4+intercept+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
```

```{r}
summary(modelBLUP_5_3_brms)
```

```{r}
ggplot(as.data.frame(conditional_effects(modelBLUP_5_3_brms, 
                         effects = "slope:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=slope,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

## Temp*slope+Temp*intercept+volume

```{r eval=FALSE, include=FALSE}
# Models with zero-inflation
modelBLUP_6_3_brms<-brm(bf(round(intactseed)~slope*cmean_4+intercept*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                           zi~slope*cmean_4+intercept*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
```

```{r}
summary(modelBLUP_6_3_brms)
```

```{r}
ggplot(as.data.frame(conditional_effects(modelBLUP_6_3_brms, 
                         effects = "slope:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`slope:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=slope,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Slope of the reaction norm")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

```{r}
ggplot(as.data.frame(conditional_effects(modelBLUP_6_3_brms, 
                         effects = "intercept:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`intercept:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=intercept,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Intercept of the reaction norm")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

## Temp*pc1+volume

PCA of slope and intercept

```{r}
PCA1<-prcomp(~datadef_BLUPs$slope+datadef_BLUPs$intercept,center=T,scale=T)
summary(PCA1)
plot(PCA1)
biplot(PCA1)

PCA1_scores<-as.data.frame(predict(PCA1))
datadef_BLUPs$PC1<-PCA1_scores$PC1
datadef_BLUPs$PC2<-PCA1_scores$PC2
hist(datadef_BLUPs$PC1)
hist(datadef_BLUPs$PC2)
```


```{r eval=FALSE, include=FALSE}
# Models with zero-inflation
modelBLUP_7_3_brms<-brm(bf(round(intactseed)~PC1*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                           zi~PC1*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.99))
```

```{r}
summary(modelBLUP_7_3_brms)
```

```{r}
ggplot(as.data.frame(conditional_effects(modelBLUP_7_3_brms, 
                         effects = "PC1:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`PC1:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=PC1,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("PC1 of the reaction norm")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

## Temp*slope_residuals+volume

Calculate residuals of linear model of slope of the RN on the intercept of the RN, and use these residuals as explanatory variable.

```{r}
model_residuals<-lm(slope~intercept,datadef_BLUPs)
summary(model_residuals)
datadef_BLUPs$slope_residuals<-residuals(model_residuals)
hist(datadef_BLUPs$slope_residuals)
```

```{r eval=FALSE, include=FALSE}
# Models with zero-inflation
modelBLUP_8_3_brms<-brm(bf(round(intactseed)~slope_residuals*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                           zi~slope_residuals*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.85))
```

```{r}
print(modelBLUP_8_3_brms,digits=3)
```

```{r}
ggplot(as.data.frame(conditional_effects(modelBLUP_8_3_brms, 
                         effects = "slope_residuals:cmean_4",dpar="zi",
                         int_conditions=
                           list(cmean_4=quantile))$`slope_residuals:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=slope_residuals,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Slope of the reaction norm (residuals of slope on intercept)")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1))+
  labs(fill="April temperature\n(mean-centred)",
       colour="April temperature\n(mean-centred)")
```

### Figure 3

```{r}
ggplot(as.data.frame(conditional_effects(modelBLUP_8_3_brms,
                                           effects="slope_residuals:cmean_4",
                                           dpar="zi",
                                           int_conditions=
                           list(cmean_4=quantile))$`slope_residuals:cmean_4`)%>%
           mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__),
       aes(x=slope_residuals,y=prob_fitness, group=cmean_4))+
  geom_ribbon(aes(ymin=lower,ymax=upper,
                  fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(size=1,aes(color=as.factor(round(cmean_4,2))))+
  my_theme_legend()+
  xlab("Slope of the reaction norm\n(residuals of slope on intercept)")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1),
                     guide = guide_legend(title.position = "top",
                                          nrow=1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1),
                    guide = guide_legend(title.position = "top",
                                         nrow=1))+
  labs(fill="April temperature (mean-centered)",
       colour="April temperature (mean-centered)")+
  #theme(legend.position=c(0.42,0.9),legend.box="horizontal")
  theme(legend.position="top",legend.box="horizontal",legend.title.align=0.5)
ggsave(filename="output/figures/fig3.tiff",device="tiff",
       width=12,height=14,units="cm",dpi=300,compression="lzw")
```

### Figure 3 (facets & data points)

```{r}
datadef_BLUPs$fitness_01<-with(datadef_BLUPs,
                               ifelse(intactseed==0,0,ifelse(intactseed>0,1,NA)))

quantile(datadef_BLUPs$cmean_4)

mean(subset(datadef_BLUPs,cmean_4<=-0.8998744)$cmean_4) 
# Mean cat 1 = -1.451023
mean(subset(datadef_BLUPs,cmean_4>-0.8998744&cmean_4<=-0.2282078)$cmean_4) 
# Mean cat 2 = -0.5470505
mean(subset(datadef_BLUPs,cmean_4>-0.2282078&cmean_4<=0.9534589)$cmean_4) 
# Mean cat 3 = 0.1058983
mean(subset(datadef_BLUPs,cmean_4>0.9534589)$cmean_4) 
# Mean cat 4 = 1.873892

datadef_BLUPs<-datadef_BLUPs%>%
  mutate(cmean_4_cat=as.factor(
    ifelse(cmean_4<=-0.8998744,1,
         ifelse(cmean_4>-0.8998744&cmean_4<=-0.2282078,2,
                ifelse(cmean_4>-0.2282078&cmean_4<=0.9534589,3,
                       ifelse(cmean_4>0.9534589,4,NA))))))
```

```{r}
# copy from maladaptive responses code revision - continue copying from there!
ggplot()+
  geom_ribbon(data=as.data.frame(
    conditional_effects(modelBLUP_8_3_brms,
                        effects="slope_residuals:cmean_4",
                        dpar="zi",
                        int_conditions=list(
                          cmean_4=c(-1.451023,-0.5470505,
                                    0.1058983,
                                    1.873892)))$`slope_residuals:cmean_4`)%>%
      mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__)%>%
      mutate(cmean_4_cat=as.factor(
        ifelse(cmean_4<=-0.8998744,1,
               ifelse(cmean_4>-0.8998744&cmean_4<=-0.2282078,2,
                      ifelse(cmean_4>-0.2282078&cmean_4<=0.9534589,3,
                             ifelse(cmean_4>0.9534589,4,NA)))))),
    aes(x=slope_residuals,y=prob_fitness, group=cmean_4,
        ymin=lower,ymax=upper,
        fill=as.factor(round(cmean_4,2))),alpha=0.2)+
  geom_line(data=as.data.frame(
    conditional_effects(modelBLUP_8_3_brms,
                        effects="slope_residuals:cmean_4",
                        dpar="zi",
                        int_conditions=list(
                          cmean_4=c(-1.451023,-0.5470505,
                                    0.1058983,
                                    1.873892)))$`slope_residuals:cmean_4`)%>%
      mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__)%>%
      mutate(cmean_4_cat=as.factor(
        ifelse(cmean_4<=-0.8998744,1,
               ifelse(cmean_4>-0.8998744&cmean_4<=-0.2282078,2,
                      ifelse(cmean_4>-0.2282078&cmean_4<=0.9534589,3,
                             ifelse(cmean_4>0.9534589,4,NA)))))),
    size=1,aes(x=slope_residuals,y=prob_fitness, group=cmean_4,
               color=as.factor(round(cmean_4,2))))+
  geom_jitter(data=datadef_BLUPs,aes(x=slope_residuals,y=fitness_01),
              size=3,alpha=0.05,width=0,height=0.03)+
  facet_grid(~cmean_4_cat)+
  my_theme_legend()+
  xlab("Slope of the reaction norm\n(residuals of slope on intercept)")+
  ylab("Predicted probability of having any fitness")+
  scale_color_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1),
                     guide = guide_legend(title.position = "top",
                                          nrow=1))+
  scale_fill_manual(values=plasma(n=5,begin=0.4,end=0.9,direction=-1),
                    guide = guide_legend(title.position = "top",
                                         nrow=1))+
  labs(fill="April temperature (mean-centered)",
       colour="April temperature (mean-centered)")+
  #theme(legend.position=c(0.42,0.9),legend.box="horizontal")
  theme(legend.position="top",legend.box="horizontal",legend.title.align=0.5)
```


# Appendix S1

## Model in Table S4

```{r}
color_scheme_set("viridis")
arrayS6<-as.array(modelBLUP_8_3_brms)
dimnames(arrayS6)[[3]][1:13] <- c("Intercept","zi_Intercept","BLUP slope (resid.)",
                                 "April temperature","Vegetative size",
                                 "BLUP slope (resid.) : April temperature",
                                 "zi_BLUP slope (resid.)","zi_April temperature",
                                 "zi_Vegetative size",
                                 "zi_BLUP slope (resid.) : April temperature",
                                 "Individual: SD (intercept)",
                                 "Individual: SD (zi_intercept)","Shape")
mcmc_trace(arrayS6,pars=c("Intercept","zi_Intercept","BLUP slope (resid.)",
                          "April temperature"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_10a.tiff",
        device="tiff",width=20,height=32,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS6,pars=c("Vegetative size",
                          "BLUP slope (resid.) : April temperature",
                          "zi_BLUP slope (resid.)","zi_April temperature"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_10b.tiff",
        device="tiff",width=20,height=32,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS6,pars=c("zi_Vegetative size",
                          "zi_BLUP slope (resid.) : April temperature",
                          "Individual: SD (intercept)",
                          "Individual: SD (zi_intercept)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_10c.tiff",
        device="tiff",width=20,height=32,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS6,pars=c("Shape"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_10d.tiff",
        device="tiff",width=20,height=8,units="cm",dpi=300,compression="lzw")
```

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(rhat(modelBLUP_8_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(modelBLUP_8_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_11.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

```{r}
yrep4_3<-posterior_predict(modelBLUP_8_3_brms, draws = 500)
ppc_dens_overlay(y1, yrep4_3[1:100,])+my_theme_legend()+xlim(0, 10)+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank())
ggsave(filename="output/figures/figS1_12.tiff",
       device="tiff",width=24,height=14,units="cm",dpi=300,compression="lzw")
```

## Model in Table S5

```{r}
color_scheme_set("viridis")
arrayS5<-as.array(modelBLUP_2_3_brms)
dimnames(arrayS5)[[3]][1:13] <- c("Intercept","zi_Intercept","BLUP slope",
                                 "April temperature","Vegetative size",
                                 "BLUP slope : April temperature",
                                 "zi_BLUP slope","zi_April temperature",
                                 "zi_Vegetative size",
                                 "zi_BLUP slope : April temperature",
                                 "Individual: SD (intercept)",
                                 "Individual: SD (zi_intercept)","Shape")
mcmc_trace(arrayS5,pars=c("Intercept","zi_Intercept","BLUP slope",
                          "April temperature"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_13a.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS5,pars=c("Vegetative size","BLUP slope : April temperature",
                          "zi_BLUP slope","zi_April temperature"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_13b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS5,pars=c("zi_Vegetative size",
                          "zi_BLUP slope : April temperature",
                          "Individual: SD (intercept)",
                          "Individual: SD (zi_intercept)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_13c.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS5,pars=c("Shape"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS1_13d.tiff",
        device="tiff",width=20,height=7,units="cm",dpi=300,compression="lzw")
```

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(rhat(modelBLUP_2_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(modelBLUP_2_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS1_14.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

```{r}
ppc_dens_overlay(y1, yrep2_3[1:100,])+my_theme_legend()+xlim(0, 10)+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank())
ggsave(filename="output/figures/figS1_15.tiff",
       device="tiff",width=24,height=14,units="cm",dpi=300,compression="lzw")
```

# Save large objects as .RData file

```{r eval=FALSE, include=FALSE}
save(modelBLUP_1_1_brms,modelBLUP_1_2_brms,modelBLUP_1_3_brms,
     modelBLUP_2_1_brms,modelBLUP_2_2_brms,modelBLUP_2_3_brms,
     modelBLUP_3_3_brms,modelBLUP_4_3_brms,
     modelBLUP_5_3_brms,
     modelBLUP_6_3_brms,
     modelBLUP_7_3_brms,
     modelBLUP_8_3_brms,
     file = "output/large_objects_BLUPs_brms.RData")
save(modelBLUP_2_3_brms,file = "output/model_without_measurement_error.RData")
```

```{r include=FALSE}
sessionInfo()
```