---
title: "Lathyrus ms2: selection on reaction norms for flowering time"
subtitle: "Variation in selection among years using BLUPs - glmmTMB models"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
library(beepr)
library(ggthemes)
library(ggeffects)
library(glmmTMB)
library(knitr)
```

```{r Define ggplot themes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r load models, include=FALSE}
# Load previously run models
load("output/large_objects_BLUPs.RData")
```

# Read data and check ns

```{r}
datadef_BLUPs<-read.csv("data/datadef_BLUPs.csv") 
head(datadef_BLUPs)
```

```{r}
ggplot(datadef_BLUPs,aes(x=intactseed))+geom_histogram()
```

# Models with glmmTMB

## Temp*slope

Does selection on plasticity depend on temperature?

Trying different distributions and models with and without zero-inflation

```{r}
# Models without zero-inflation
modelBLUP_1_1<-glmmTMB(round(intactseed)~slope*cmean_4+(1|id),family="poisson",
                data=datadef_BLUPs)
modelBLUP_1_2<-glmmTMB(round(intactseed)~slope*cmean_4+(1|id),family="nbinom2",
                data=datadef_BLUPs)
# Models with zero-inflation
modelBLUP_1_3<-glmmTMB(formula=round(intactseed)~slope*cmean_4+(1|id),
                ziformula=~.,family="poisson",data=datadef_BLUPs)
modelBLUP_1_4<-glmmTMB(formula=round(intactseed)~slope*cmean_4+(1|id),
                ziformula=~.,family="nbinom2",data=datadef_BLUPs)
```

```{r}
AIC(modelBLUP_1_1,modelBLUP_1_2,modelBLUP_1_3,modelBLUP_1_4)
anova(modelBLUP_1_1,modelBLUP_1_2,modelBLUP_1_3,modelBLUP_1_4)
```

modelBLUP_1_4 with a zero-inflated negative binomial is the best model.

```{r}
summary(modelBLUP_1_4)
```

Predicted effects:

```{r message=FALSE, warning=FALSE}
plot(ggpredict(modelBLUP_1_4,terms=c("slope","cmean_4"),type="zero_inflated"))
# Predicted values are conditioned on the fixed effects 
# and the zero-inflation component
plot(ggpredict(modelBLUP_1_4,terms=c("slope","cmean_4"),type="zi_prob"))
# Predicted zero-inflation probability
```

Selection for plasticity (more negative slopes) increases with temperature, but this is only driven by the zero-inflated part of the model, i.e. the probability of fitness being zero decreases more with higher plasticity (more negative slopes) with warm temperatures.

## Temp*slope+volume

Does selection on plasticity depend on temperature when considering also plant size?

Trying different distributions and models with and without zero-inflation

```{r}
# Models without zero-inflation
modelBLUP_2_1<-glmmTMB(round(intactseed)~slope*cmean_4+cn_shoot_vol_mean_sqrt+
                         (1|id),family="poisson",data=datadef_BLUPs)
modelBLUP_2_2<-glmmTMB(round(intactseed)~slope*cmean_4+cn_shoot_vol_mean_sqrt+
                         (1|id),family="nbinom2",data=datadef_BLUPs)
# Models with zero-inflation
modelBLUP_2_3<-glmmTMB(formula=round(intactseed)~slope*cmean_4+
                         cn_shoot_vol_mean_sqrt+(1|id),ziformula=~.,
                       family="poisson",data=datadef_BLUPs)
modelBLUP_2_4<-glmmTMB(formula=round(intactseed)~slope*cmean_4+
                         cn_shoot_vol_mean_sqrt+(1|id),ziformula=~.,
                       family="nbinom2",data=datadef_BLUPs)
```

```{r}
AIC(modelBLUP_2_1,modelBLUP_2_2,modelBLUP_2_3,modelBLUP_2_4)
anova(modelBLUP_2_1,modelBLUP_2_2,modelBLUP_2_3,modelBLUP_2_4)
```

modelBLUP_2_4 with a zero-inflated negative binomial is the best model.

```{r}
summary(modelBLUP_2_4)
```

Predicted effects:

```{r}
plot(ggpredict(modelBLUP_2_4,terms=c("slope","cmean_4"),type="zero_inflated"))
# Predicted values are conditioned on the fixed effects 
# and the zero-inflation component
plot(ggpredict(modelBLUP_2_4,terms=c("slope","cmean_4"),type="zi_prob"))
# Predicted zero-inflation probability
```

The interaction slope:temperature in the zero-inflation part of the model is not significant (p=0.053) when including volume.

## Volume*slope

Does selection on plasticity depend on plant size?

Trying different distributions and models with and without zero-inflation

```{r}
# Models without zero-inflation
modelBLUP_3_1<-glmmTMB(round(intactseed)~slope*cn_shoot_vol_mean_sqrt+cmean_4+
                         (1|id),family="poisson",data=datadef_BLUPs)
modelBLUP_3_2<-glmmTMB(round(intactseed)~slope*cn_shoot_vol_mean_sqrt+cmean_4+
                         (1|id),family="nbinom2",data=datadef_BLUPs)
# Models with zero-inflation
modelBLUP_3_3<-glmmTMB(formula=round(intactseed)~slope*cn_shoot_vol_mean_sqrt+
                         cmean_4+(1|id),ziformula=~.,
                       family="poisson",data=datadef_BLUPs)
modelBLUP_3_4<-glmmTMB(formula=round(intactseed)~slope*cn_shoot_vol_mean_sqrt+
                         cmean_4+(1|id),ziformula=~.,
                       family="nbinom2",data=datadef_BLUPs)
```

```{r}
AIC(modelBLUP_3_1,modelBLUP_3_2,modelBLUP_3_3,modelBLUP_3_4)
anova(modelBLUP_3_1,modelBLUP_3_2,modelBLUP_3_3,modelBLUP_3_4)
```

modelBLUP_3_4 with a zero-inflated negative binomial is the best model.

```{r}
summary(modelBLUP_3_4)
```

Predicted effects:

```{r}
plot(ggpredict(modelBLUP_3_4,terms=c("slope","cn_shoot_vol_mean_sqrt"),
               type="zero_inflated"))
# Predicted values are conditioned on the fixed effects 
# and the zero-inflation component
plot(ggpredict(modelBLUP_3_4,terms=c("slope","cn_shoot_vol_mean_sqrt"),
               type="zi_prob"))
# Predicted zero-inflation probability
```

Selection for plasticity (more negative slopes) decreases with size, but this is only driven by the zero-inflated part of the model, i.e. the probability of fitness being zero decreases more with higher plasticity (more negative slopes) with lower sizes.

## Temp*slope+volume*slope

Does selection on plasticity depend on temperature AND on plant size?

Trying different distributions and models with and without zero-inflation

```{r}
# Models without zero-inflation
modelBLUP_4_1<-glmmTMB(round(intactseed)~slope*cmean_4+
                         slope*cn_shoot_vol_mean_sqrt+cmean_4+
                         (1|id),family="poisson",data=datadef_BLUPs)
modelBLUP_4_2<-glmmTMB(round(intactseed)~slope*cmean_4+
                         slope*cn_shoot_vol_mean_sqrt+cmean_4+
                         (1|id),family="nbinom2",data=datadef_BLUPs)
# Models with zero-inflation
modelBLUP_4_3<-glmmTMB(formula=round(intactseed)~slope*cmean_4+
                         slope*cn_shoot_vol_mean_sqrt+
                         cmean_4+(1|id),ziformula=~.,
                       family="poisson",data=datadef_BLUPs)
modelBLUP_4_4<-glmmTMB(formula=round(intactseed)~slope*cmean_4+
                         slope*cn_shoot_vol_mean_sqrt+
                         cmean_4+(1|id),ziformula=~.,
                       family="nbinom2",data=datadef_BLUPs)
```

```{r}
AIC(modelBLUP_4_1,modelBLUP_4_2,modelBLUP_4_3,modelBLUP_4_4)
anova(modelBLUP_4_1,modelBLUP_4_2,modelBLUP_4_3,modelBLUP_4_4)
```

modelBLUP_4_4 with a zero-inflated negative binomial is the best model.

```{r}
summary(modelBLUP_4_4)
```

Predicted effects:

```{r}
plot(ggpredict(modelBLUP_4_4,terms=c("slope","cmean_4"),type="zero_inflated"))
plot(ggpredict(modelBLUP_4_4,terms=c("slope","cn_shoot_vol_mean_sqrt"),
               type="zero_inflated"))
# Predicted values are conditioned on the fixed effects 
# and the zero-inflation component
plot(ggpredict(modelBLUP_4_4,terms=c("slope","cmean_4"),type="zi_prob"))
plot(ggpredict(modelBLUP_4_4,terms=c("slope","cn_shoot_vol_mean_sqrt"),
               type="zi_prob"))
# Predicted zero-inflation probability
```

Similar results as in previous models: Selection for plasticity (more negative slopes) increases with temperature and decreases with size, but this is only driven by the zero-inflated part of the model, i.e. the probability of fitness being zero decreases more with higher plasticity (more negative slopes) with warm temperatures and lower sizes.

<!-- # Models with brms -->

<!-- ## Temp*slope -->

<!-- ```{r eval=FALSE, include=FALSE} -->
<!-- my.cores <- detectCores() -->
<!-- modelBLUP_1<-brm(formula=round(intactseed)~slope*cmean_4+(1|id), -->
<!--                   family="poisson",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 6000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_1nb<-brm(formula=round(intactseed)~slope*cmean_4+(1|id), -->
<!--                   family="negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores, -->
<!--                   control = list(adapt_delta = 0.99)) -->
<!-- modelBLUP_1zinb<-brm(formula=round(intactseed)~slope*cmean_4+(1|id), -->
<!--                   family="zero_inflated_negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 2000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores, -->
<!--                   control = list(adapt_delta = 0.99)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- conditional_effects(modelBLUP_1) -->
<!-- conditional_effects(modelBLUP_1nb) -->
<!-- conditional_effects(modelBLUP_1zinb) -->
<!-- ``` -->

<!-- ## Temp*slope+volume -->

<!-- ```{r eval=FALSE, include=FALSE} -->
<!-- modelBLUP_2<-brm(formula=round(intactseed)~slope*cmean_4+ -->
<!--                    cn_shoot_vol_mean_sqrt +(1|id), -->
<!--                   family="poisson",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_2nb<-brm(formula=round(intactseed)~slope*cmean_4+ -->
<!--                    cn_shoot_vol_mean_sqrt +(1|id), -->
<!--                   family="negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_2zinb<-brm(formula=round(intactseed)~slope*cmean_4+ -->
<!--                    cn_shoot_vol_mean_sqrt +(1|id), -->
<!--                   family="zero_inflated_negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- conditional_effects(modelBLUP_2) -->
<!-- conditional_effects(modelBLUP_2nb) -->
<!-- conditional_effects(modelBLUP_2zinb) -->
<!-- ``` -->

<!-- ## Volume*slope -->

<!-- ```{r eval=FALSE, include=FALSE} -->
<!-- modelBLUP_3<-brm(formula=round(intactseed)~slope*cn_shoot_vol_mean_sqrt+(1|id), -->
<!--                   family="poisson",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_3nb<-brm(formula=round(intactseed)~slope*cn_shoot_vol_mean_sqrt+(1|id), -->
<!--                   family="negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_3zinb<-brm(formula=round(intactseed)~slope*cn_shoot_vol_mean_sqrt+(1|id), -->
<!--                   family="zero_inflated_negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- conditional_effects(modelBLUP_3) -->
<!-- conditional_effects(modelBLUP_3nb) -->
<!-- conditional_effects(modelBLUP_3zinb) -->
<!-- ``` -->

<!-- ## Temp*slope+volume*slope -->

<!-- ```{r eval=FALSE, include=FALSE} -->
<!-- modelBLUP_4<-brm(formula=round(intactseed)~slope*cmean_4+ -->
<!--                    slope*cn_shoot_vol_mean_sqrt+(1|id), -->
<!--                   family="poisson",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_4nb<-brm(formula=round(intactseed)~slope*cmean_4+ -->
<!--                    slope*cn_shoot_vol_mean_sqrt+(1|id), -->
<!--                   family="negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- modelBLUP_4zinb<-brm(formula=round(intactseed)~slope*cmean_4+ -->
<!--                    slope*cn_shoot_vol_mean_sqrt+(1|id), -->
<!--                   family="zero_inflated_negbinomial",data=datadef_BLUPs, -->
<!--                   warmup = 1000,iter = 4000,thin=2,chains=4, -->
<!--                   inits = "random",seed = 12345,cores = my.cores) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- conditional_effects(modelBLUP_4) -->
<!-- conditional_effects(modelBLUP_4nb) -->
<!-- conditional_effects(modelBLUP_4zinb) -->
<!-- ``` -->

# Save large objects as .RData file

```{r eval=FALSE, include=FALSE}
save(modelBLUP_1_1,modelBLUP_1_2,modelBLUP_1_3,modelBLUP_1_4,
     modelBLUP_2_1,modelBLUP_2_2,modelBLUP_2_3,modelBLUP_2_4,
     modelBLUP_3_1,modelBLUP_3_2,modelBLUP_3_3,modelBLUP_3_4,
     modelBLUP_4_1,modelBLUP_4_2,modelBLUP_4_3,modelBLUP_4_4,
     modelBLUP_1,modelBLUP_2,modelBLUP_3,modelBLUP_4,
     modelBLUP_1nb,modelBLUP_2nb,modelBLUP_3nb,modelBLUP_4nb,
     #modelBLUP_1zinb,modelBLUP_2zinb,modelBLUP_3zinb,modelBLUP_4zinb,
     file = "output/large_objects_BLUPs.RData")
```

```{r include=FALSE}
sessionInfo()
```