---
title: "Lathyrus ms2: Selection on reaction norms - Extraction of RN parameters with lms"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load data created in previous notebooks, include=FALSE}
load("C:/Users/avald/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_3yrs1.RData")
load("C:/Users/avald/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_4yrs1.RData")
load("C:/Users/avald/Dropbox/SU/Projects/lathyrus/lathyrus_ms2/code/data_5yrs1.RData")
```

```{r load packages, include=FALSE}
library(knitr)
library(pander)
library(tidyverse)
library(ggthemes)
library(broom)
library(gridExtra)
library(lme4)
library(lmerTest)
library(MuMIn)
library(ggrepel)
library(car)
library(effects)
library(RColorBrewer)
library(MCMCglmm)
library(blme)
library(beepr)
```

```{r Define ggplot themes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r pander options, include=FALSE}
panderOptions('table.continues', '')
```

Reaction norm of FFD to mean daily temperature April, using the subset of data with plants with 4 or more years of data

Plot of the 243 reaction norms

```{r plot RNs, echo=FALSE, fig.height=4, fig.width=5}
ggplot(data_4yrs,aes(x=scale(mean_4,center=T,scale=F),y=FFD,group=id))+
  geom_smooth(method=lm,se=F,color="black",size=0.05)+
  xlab("Mean.centred mean daily temperature April")+
  ylab("First flowering date")+my_theme()
```

Extraction of parameters for every reaction norm (i.e. for each individual)

```{r extract parameters of RNs, echo=FALSE}
# Extract out common code with function
id_model <- function(df) {
  lm(FFD ~ scale(mean_4,center=T,scale=F), data = df)
}

(RNs4 <- data_4yrs%>% 
  group_by(id)%>%                                 # Grouped df
  nest()%>%                                       # Nested df
  mutate(model = map(data, id_model))%>%          # Repeat common code using map function
  mutate(coefs = map(model, broom::tidy))%>%      # Tidy df - a row for each coef
  unnest(coefs)%>%                                # List of dfs back into regular df
  select(id,term,estimate)%>%                     # Select important columns
  spread(key = term, value = estimate)%>%         # Long to wide format
  rename(intercept4=`(Intercept)`,slope4=`scale(mean_4, center = T, scale = F)`)) # Rename variables
```

Correlation among intercepts and slopes:

```{r corr intercepts and slopes, echo=FALSE, fig.height=4, fig.width=5}
with(RNs4,cor(intercept4,slope4))
ggplot(RNs4,aes(x=intercept4,y=slope4))+
  geom_point()+geom_smooth(method="lm",se=F,col="black")+
  xlab("Intercept of the reaction norm")+ylab("Slope of the reaction norm")+my_theme()
```

Correlation here is much lower than using BLUPs - WHY???
Problems with random effect of year in mixed models below?

Check that parameters are correct graphically by plotting some reaction norms

```{r check that parameters are correct, echo=FALSE, fig.height=4, fig.width=6}
ggplot(subset(data_4yrs,id=="new_103"|id=="new_104"|id=="new_106"|
                id=="new_107"|id=="new_108"|id=="new_109"),
       aes(x=scale(mean_4,center=T,scale=F),y=FFD,color=id))+geom_smooth(method=lm,se=F,fullrange=T)+
  xlab("Mean-centred mean daily temperature April")+
  ylab("First flowering date")+my_theme_legend()+
  scale_y_continuous(breaks=seq(45,75,2))
```