---
title: "Spring temperature drives phenotypic selection on plasticity of flowering time"
author : "Alicia Valdés"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_notebook:
    toc: yes
    toc_depth: '4'
    latex_engine: xelatex
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.width=4, fig.height=3) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(tidyr)
library(brms)
library(ggthemes)
library(knitr)
library(bayesplot)
library(tidybayes)
library(parallel)
library(future)
library(ggeffects)
library(arm)
library(cowplot)
library(broom)
library(viridis)
library(MCMCglmm)
```

```{r Define ggplot themes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

```{r load models, include=FALSE}
# Load previously run models and stuff
load("C:/Users/alici/Desktop/large_objects/large_objects.RData")
load("C:/Users/alici/Desktop/large_objects/large_objects_BLUPs_brms.RData")
load("C:/Users/alici/Desktop/large_objects/large_objects_periods.RData")
load("C:/Users/alici/Desktop/large_objects/large_objects_var.RData")
```

# Read data and check ns

```{r}
datadef<-read.csv("data/datadef.csv") 
head(datadef)
```

Number of individuals in each period:

```{r}
length(with(subset(datadef,period=="old"),unique(id)))
length(with(subset(datadef,period=="new"),unique(id)))
```

Number of observations in each period:

```{r}
nrow(subset(datadef,period=="old"))
nrow(subset(datadef,period=="new"))
```

Number of cases with FFD in each period:

```{r}
nrow(subset(datadef,period=="old"&!is.na(FFD)))
nrow(subset(datadef,period=="new"&!is.na(FFD)))
```

# Data preparation

```{r data prep MCMCglmm, message=FALSE, warning=FALSE}
datadef_total<-datadef %>%
  group_by(id)%>%
  # Calculate mean fitness per year of study 
  # and mean fitness per flowering event
  summarise(mean_fitness_study=sum(intactseed,na.rm=T)/mean(n_years_study),
            mean_fitness_fl=sum(intactseed,na.rm=T)/mean(n_years_fl_fitness))%>%
   arrange(.,id) # Order by id

with(datadef_total,cor(mean_fitness_study,mean_fitness_fl))  # Highly corr (0.87)

# Calculate mean shoot volume for each id using values of shoot volume 
# for all ids/years (including flowering and non-flowering years)

shoot_vol_all_means<-datadef[c(1,3,10)]%>%
  group_by(id)%>%
  summarise(shoot_vol_mean=mean(shoot_vol,na.rm=T)) 
# Mean of all available values 

# Join shoot volume data
datadef_total<-datadef_total%>%left_join(shoot_vol_all_means)%>%
  left_join(unique(datadef[c(2,3,11)]))
head(datadef_total)
nrow(subset(datadef_total,is.na(shoot_vol_mean))) 
# 46 ids with no info on shoot volume

# Add first_yr to total data (this was needed for MCMCglmm models) 
datadef_total$first_yr<-ifelse(grepl("old",as.character(datadef_total$id)),
                               1987,2006)

# Using sqrt of mean shoot volume over all years when available, centred
datadef_total<-datadef_total%>%
  mutate(shoot_vol_mean_sqrt=sqrt(shoot_vol_mean),
         cn_shoot_vol_mean_sqrt=as.vector(scale(shoot_vol_mean_sqrt,center=T,scale=F)))
```

# Hypothesis 1: Univariate models

## Models in Table 1

### Table 1A

FFD with random effect of year only. This model assumes no among-individual variation around the average population-level reaction norm.

```{r}
my.cores <- detectCores()
```

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm<-brm(formula=FFD~cmean_4+(1|year),data=datadef,
                warmup = 1000,iter = 4000,thin=2,chains=4,
                # 4 chains, each with 4000 iterations
                inits = "random",seed = 12345,cores = my.cores)
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_yearonly.all.brm)
```

### Table 1B

FFD with random effects of year and individual-intercept. This model assumes among-individual variation in reaction norm elevation.

```{r eval=FALSE, include=FALSE}
univar.FFD.all.brm<-brm(formula=FFD~cmean_4+(1|year)+(1|id),data=datadef,
                warmup = 1000,iter = 4000,thin=2,chains=4,
                # 4 chains, each with 4000 iterations
                inits = "random",seed = 12345,cores = my.cores)
# Total of 6000 post-warmup samples
```

```{r }
summary(univar.FFD.all.brm)
```

### Table 1C

Random regression for FFD, including random effects of individual slopes and covariance between intercept and slope. This model assumes among-individual variation in reaction norm elevation and slope.

```{r eval=FALSE, include=FALSE}
univar.FFD_RR.all.brm<-brm(formula=FFD~cmean_4+(1|year)+(cmean_4|id),
                           data=datadef,
                           warmup = 1000,iter = 4000,thin=2,chains=4,
                           inits = "random",seed = 12345,cores = my.cores,
                           sample_prior="yes")
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_RR.all.brm) 
```

### Compare models

This allows  us to test if adding among-individual variation in elevation and slope of the RN improves model fit. 

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm <- add_criterion(univar.FFD_yearonly.all.brm,"loo")
univar.FFD.all.brm <- add_criterion(univar.FFD.all.brm,"loo")
univar.FFD_RR.all.brm <- add_criterion(univar.FFD_RR.all.brm,"loo")
```

```{r}
loo_comp<-loo_compare(univar.FFD_yearonly.all.brm,univar.FFD.all.brm,
                      univar.FFD_RR.all.brm, criterion="loo")
```

```{r}
loo_comp
```

The results indicate that adding a random intercept (i.e. among-individual variation in RN elevation) improves the fit, and that adding a random slope (i.e. among-individual variation in RN slope) improves the fit even more. Thus, the random regression model is the best model.

### Bayes_R2

Measure of fit: Bayesian R2, which looks at the model expected variance / (expected variance + residual variance).

```{r}
bayes_R2(univar.FFD_yearonly.all.brm)
bayes_R2(univar.FFD.all.brm)
bayes_R2(univar.FFD_RR.all.brm)
```

### Extract BLUPs from random regression model

```{r}
BLUPs_MCMC.all.brms  <- cbind(as.factor(c(1:837)),
                              as.data.frame(ranef(univar.FFD_RR.all.brm)$id)
                              [c(1:2,5:6)])
colnames(BLUPs_MCMC.all.brms) <- c("id", "intercept", "intercept_sd",
                                   "slope","slope_sd")
with(BLUPs_MCMC.all.brms,cor(intercept,slope)) # highly correlated!
```

Add BLUPs to data set

```{r}
datadef<-datadef%>%left_join(BLUPs_MCMC.all.brms%>%
                               dplyr::select(intercept,intercept_sd,
                                             slope,slope_sd)%>%
                               rownames_to_column(var="id"))
```

### Figure 1

```{r}
predict_id_2<-ggpredict(univar.FFD_RR.all.brm,
                      terms=c("cmean_4","id"),type="random")
predict_mean_2<-ggpredict(univar.FFD_RR.all.brm,terms=c("cmean_4"),
                          type="random")
```

```{r fig.height=10, fig.width=12}
data_fig1<-full_join(data.frame(predict_id_2),
          datadef%>%dplyr::select(id,slope)%>%group_by(id)%>%
            summarise(slope=mean(slope))%>%rename(group=id))
plot_grid(ggplot()+
  geom_line(data=subset(data_fig1,group!="new_46"|group!="new_131"),
            aes(x=x,y=predicted,group=group,color=slope),
            size=0.01,alpha=0.7)+
  scale_color_viridis(end=0.9,option = "C")+
  geom_line(data=subset(data_fig1,group=="new_46"|group=="new_131"),
            aes(x=x,y=predicted,group=group,color=slope),
            size=1,linetype=2)+
  geom_line(data=data.frame(predict_mean_2),aes(x=x,y=predicted),
            color="black",size=1)+
  geom_point(data=subset(datadef,id=="new_46"|id=="new_131"),
             aes(x=cmean_4,y=FFD,color=slope),size=3,alpha=1)+
  my_theme()+
  xlab("April temperature (mean-centered)")+
  ylab("Predicted FFD"),
          plot_grid(ggplot(BLUPs_MCMC.all.brms,
                           aes(x=intercept+fixef(univar.FFD_RR.all.brm)[1]))+
                      geom_histogram(fill="grey",color="black")+my_theme()+
                      geom_vline(xintercept=fixef(univar.FFD_RR.all.brm)[1],
                                 size=0.5,linetype="dashed")+
                      xlab("Elevation of the reaction norm")+
                      ylab("Number of\nindividuals"),
                    ggplot(BLUPs_MCMC.all.brms,
                           aes(x=slope+fixef(univar.FFD_RR.all.brm)[2]))+
                      geom_histogram(fill="grey",color="black")+my_theme()+
                      geom_vline(xintercept=fixef(univar.FFD_RR.all.brm)[2],
                                 size=0.5,linetype="dashed")+
                      xlab("Slope of the reaction norm")+
                      ylab("Number of\nindividuals"),
                    ncol=1,labels=c("B)","C)"),label_fontfamily="serif"),
          labels = c("A)",""),label_fontfamily="serif",rel_widths = c(2,1.5))
ggsave(filename="output/figures/fig1.tiff",
       device="tiff",width=22,height=12,units="cm",dpi=300,compression="lzw")
```

Save data with BLUPs

```{r}
datadef<-datadef%>%left_join(datadef_total%>%
                               dplyr::select(id,cn_shoot_vol_mean_sqrt))
write_csv(datadef,file = "data/datadef_BLUPs.csv")
```

# Hypothesis 2: Bivariate models

## Models in Table 2

### Table 2A

Mean fitness per flowering event, with shoot volume.

Using the ID-syntax to specify fitness to be correlated with the intercept and slope of FFD on temperature.

```{r message=FALSE, warning=FALSE}
datadef<-left_join(datadef,datadef_total[c(1:3,9)]) 
# Add info on mean fitness and mean shoot volume
bf_FFD <- bf(FFD ~ cmean_4 + (cmean_4|ID1|id) + (1|year)) # Set up model formula
bf_fitness_shoot <- bf(round(mean_fitness_fl) ~  cn_shoot_vol_mean_sqrt +
                         (1|ID1|id)) # Set up model formula
# Specifying group-level effects of the same grouping factor (id here) 
# to be correlated across formulas
# Expand the | operator into |<ID>|, where <ID> can be any value (ID1 here)
# Group-level terms with the same ID1 will be modeled as correlated 
# if they share same grouping factor(s)
```

Negative binomial distribution for fitness.

```{r, eval=FALSE, include=TRUE}
bivar2.all.brm.nb<-brm(bf_FFD+bf_fitness_shoot, 
                       family = c(gaussian,negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores,
                       control = list(adapt_delta = 0.99))
```

```{r}
print(bivar2.all.brm.nb,digits=3)
```

Measure of fit: Bayesian R2

```{r}
bayes_R2(bivar2.all.brm.nb)
```

#### Extract selection coefficients

```{r}
sample.gmat2 <- function(data, replicates = 6000) {

  ##Initialize the results list (list of lists)
  foo <- list(gmat = matrix(rep(0,3*3), ncol = 3))
  results.list <- list()
  for(j in 1:replicates) { results.list[[j]] <- foo }

  for(i in 1:replicates) {
    diag(results.list[[i]]$gmat) <- data[i,5:7]^2 #Get the diagonal

    #Upper diagonal
    results.list[[i]]$gmat[1,2] <- data[i,5]*data[i,6]*data[i,9]
    results.list[[i]]$gmat[1,3] <- data[i,5]*data[i,7]*data[i,10]
    results.list[[i]]$gmat[2,3] <- data[i,6]*data[i,7]*data[i,11]

    #Lower diagonal
    results.list[[i]]$gmat[2,1] <- results.list[[i]]$gmat[1,2]
    results.list[[i]]$gmat[3,1] <- results.list[[i]]$gmat[1,3]
    results.list[[i]]$gmat[3,2] <- results.list[[i]]$gmat[2,3]

  }

  return(results.list)
}
```

```{r}
# Extract posterior samples
bivar2.all.brm.nb_post <- posterior_samples(bivar2.all.brm.nb)
bivar2.all.brm.nb_post <- as.mcmc(bivar2.all.brm.nb_post)
#head(bivar2.all.brm.nb_post)[,1:20]

# [,5] sd_id__FFD_Intercept
# [,6] sd_id__FFD_cmean_4
# [,7] sd_id__roundmeanfitnessfl_Intercept 
# [,9] cor_id__FFD_Intercept__FFD_cmean_4
# [,10] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,11] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat4 <- sample.gmat2(bivar2.all.brm.nb_post, replicates = 6000) 

sgmat4 <- lapply(sampled.gmat4, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat4 <- unname(sapply(sgmat4, '[[', 1)) #Change to matrix

sgmat4 <- t(sgmat4)

P.modelBV_RR4 <- sgmat4
P.modelBV_RR4.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR4.mode[k] <- posterior.mode(mcmc(sgmat4[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR4 <- sgmat4[,c(3,6)]
colnames(S.modelBV_RR4) <- c("S_intercepts", "S_slopes")
S.modelBV_RR4.mode <- P.modelBV_RR4.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR4))
HPDinterval(mcmc(S.modelBV_RR4))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR4 <- matrix(NA, nrow(S.modelBV_RR4) ,2)

for (i in 1:nrow(S.modelBV_RR4)) {
  P3_4 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_4[k] <- P.modelBV_RR4[i, k] }  
  P2_4 <- P3_4[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S4 <- P3_4[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR4[i,] <- solve(P2_4) %*% S4   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR4) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR4))
HPDinterval(mcmc(beta_post_RR4))
```

### Table 2B

Mean fitness per year of study, with shoot volume.

```{r message=FALSE, warning=FALSE}
bf_fitness_study_shoot <- bf(round(mean_fitness_study) ~  
                               cn_shoot_vol_mean_sqrt +
                         (1|ID1|id)) # Set up model formula
```

Negative binomial distribution for fitness.

```{r, eval=FALSE, include=TRUE}
bivar4.all.brm.nb<-brm(bf_FFD+bf_fitness_study_shoot, 
                       family = c(gaussian,negbinomial), 
                       data = datadef,warmup = 1000,iter = 4000,chains=4,thin=2,
                       inits = "random",seed = 12345,cores = my.cores)
```

```{r}
print(bivar4.all.brm.nb,digits=3) 
```

Measure of fit: Bayesian R2

```{r}
bayes_R2(bivar4.all.brm.nb)
```

#### Extract selection coefficients

```{r}
# Extract posterior samples
bivar4.all.brm.nb_post <- posterior_samples(bivar4.all.brm.nb)
bivar4.all.brm.nb_post <- as.mcmc(bivar4.all.brm.nb_post)
#head(bivar4.all.brm.nb_post)[,1:20]

# [,4] sd_id__FFD_Intercept
# [,5] sd_id__FFD_cmean_4
# [,6] sd_id__roundmeanfitnessfl_Intercept 
# [,8] cor_id__FFD_Intercept__FFD_cmean_4
# [,9] cor_id__FFD_Intercept__roundmeanfitnessfl_Intercept
# [,10] cor_id__FFD_cmean_4__roundmeanfitnessfl_Intercept

sampled.gmat8 <- sample.gmat2(bivar4.all.brm.nb_post, replicates = 6000) 

sgmat8 <- lapply(sampled.gmat8, `[`, c('gmat')) #Get list 'gmat' from each list
sgmat8 <- unname(sapply(sgmat8, '[[', 1)) #Change to matrix

sgmat8 <- t(sgmat8)

P.modelBV_RR8 <- sgmat8
P.modelBV_RR8.mode <- matrix(1:9, nrow = 3)
for (k in 1:9) P.modelBV_RR8.mode[k] <- posterior.mode(mcmc(sgmat8[,k]))

# Extract selection *differentials* (i.e. covariances) for intercept and slope:
# and calculate posterior mode and credible intervals for each
S.modelBV_RR8 <- sgmat8[,c(3,6)]
colnames(S.modelBV_RR8) <- c("S_intercepts", "S_slopes")
S.modelBV_RR8.mode <- P.modelBV_RR8.mode[1:2, 3]

posterior.mode(mcmc(S.modelBV_RR8))
HPDinterval(mcmc(S.modelBV_RR8))

# Estimate selection gradients for intercept and slope (beta = S / P)
# on each sample of posterior and extract their mode
beta_post_RR8 <- matrix(NA, nrow(S.modelBV_RR8) ,2)

for (i in 1:nrow(S.modelBV_RR8)) {
  P3_8 <- matrix(rep(NA, 9), nrow = 3) 
  # 3x3 matrix of var-cov for individual X.int, X.slope and fitness
  for (k in 1:9) {P3_8[k] <- P.modelBV_RR8[i, k] }  
  P2_8 <- P3_8[1:2, 1:2]   # 2x2 matrix of just trait intercept & slope var-cov
  S8 <- P3_8[1:2, 3]   # selection differentials on traits (last column of P3)
  beta_post_RR8[i,] <- solve(P2_8) %*% S8   # selection gradients beta = P^-1 * S
}

colnames(beta_post_RR8) <- c("beta_intercepts", "beta_slopes")
posterior.mode(mcmc(beta_post_RR8))
HPDinterval(mcmc(beta_post_RR8))
```

### Figure 2: Selection differentials and gradients

```{r}
selcoefs<-rbind(
  # mean_fitness_fl, with shoot volume
  rbind(cbind(data.frame(coef=posterior.mode(mcmc(S.modelBV_RR4))),
              data.frame(HPDinterval(mcmc(S.modelBV_RR4)))),
        cbind(data.frame(coef=posterior.mode(mcmc(beta_post_RR4))),
              data.frame(HPDinterval(mcmc(beta_post_RR4)))))%>%
    mutate(fitness_meas="mean_fitness_fl",condition="yes",
           type=c("Differentials","Differentials",
                  "Gradients","Gradients"),
           param=c("Intercept","Slope","Intercept","Slope")),
  # mean_fitness_study, with shoot volume
  rbind(cbind(data.frame(coef=posterior.mode(mcmc(S.modelBV_RR8))),
              data.frame(HPDinterval(mcmc(S.modelBV_RR8)))),
        cbind(data.frame(coef=posterior.mode(mcmc(beta_post_RR8))),
              data.frame(HPDinterval(mcmc(beta_post_RR8)))))%>%
    mutate(fitness_meas="mean_fitness_study",condition="yes",
           type=c("Differentials","Differentials",
                  "Gradients","Gradients"),
           param=c("Intercept","Slope","Intercept","Slope"))
  )
```

```{r}
ggplot(selcoefs,
       aes(x=param,y=coef,color=fitness_meas))+
  geom_errorbar(aes(ymin=lower, ymax=upper),
                width=.2,position=position_dodge(.3))+
  geom_point(size=2,position=position_dodge(.3))+
  facet_wrap(~type)+
  geom_hline(yintercept=0,lty=3)+
  scale_color_manual(values=c("black","darkgrey"))+
  scale_y_continuous(breaks=c(-2,-1,0,1,2,3,4))+
  my_theme()+xlab(NULL)+ylab("Coefficient value")
# black: mean_fitness_fl, grey: mean_fitness_study
ggsave(filename="output/figures/fig2.tiff",
        device="tiff",width=9,height=8,units="cm",dpi=300,compression="lzw")
```

# Hypothesis 3

```{r}
datadef_BLUPs<-read.csv("data/datadef_BLUPs.csv") 
head(datadef_BLUPs)
```

## Model in Table 3

Calculate residuals of linear model of slope of the RN on the intercept of the RN, and use these residuals as explanatory variable.

```{r}
model_residuals<-lm(slope~intercept,datadef_BLUPs)
summary(model_residuals)
datadef_BLUPs$slope_residuals<-residuals(model_residuals)
hist(datadef_BLUPs$slope_residuals)
```

```{r eval=FALSE, include=FALSE}
# Models with zero-inflation
modelBLUP_8_3_brms<-brm(bf(round(intactseed)~slope_residuals*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                           zi~slope_residuals*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores,
                        control = list(adapt_delta = 0.85))
```

```{r}
print(modelBLUP_8_3_brms,digits=3)
```

## Figure 3

```{r}
datadef_BLUPs$fitness_01<-with(datadef_BLUPs,
                               ifelse(intactseed==0,0,
                                      ifelse(intactseed>0,1,NA)))

quantile(datadef_BLUPs$mean_4)

mean(subset(datadef_BLUPs,mean_4<=4.565000)$mean_4) 
# Mean cat 1 = 4.013852
mean(subset(datadef_BLUPs,mean_4>4.565000&mean_4<=5.236667)$mean_4) 
# Mean cat 2 = 5.009445
mean(subset(datadef_BLUPs,mean_4>5.236667&mean_4<=6.418333)$mean_4) 
# Mean cat 3 = 5.38374
mean(subset(datadef_BLUPs,mean_4>6.418333)$mean_4) 
# Mean cat 4 = 7.128874

datadef_BLUPs<-datadef_BLUPs%>%
  mutate(mean_4_cat=as.factor(
    ifelse(mean_4<=-4.565000,1,
         ifelse(mean_4>4.565000&mean_4<=5.236667,2,
                ifelse(mean_4>5.236667&mean_4<=6.418333,3,
                       ifelse(mean_4>6.418333,4,NA))))))

label_names1 <- list(
  '1'="First quartile (3.69,4.57)\nMean April temperature = 4.01 ºC",
  '2'="Second quartile [4.57,5.24)\nMean April temperature = 5.01 ºC",
  '3'="Third quartile [5.24,6.42)\nMean April temperature = 5.38 ºC",
  '4'="Fourth quartile [6.42,8.44)\nMean April temperature = 7.13 ºC"
)

labeller_function1 <- function(variable,value){
  return(label_names1[value])
}
```

```{r fig.height=5, fig.width=9}
ggplot()+
  geom_ribbon(data=as.data.frame(
    conditional_effects(modelBLUP_8_3_brms,
                        effects="slope_residuals:cmean_4",
                        dpar="zi",
                        int_conditions=list(
                          cmean_4=c(-1.451023,-0.5470505,
                                    0.1058983,
                                    1.873892)))$`slope_residuals:cmean_4`)%>%
      mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__)%>%
      mutate(cmean_4_cat=as.factor(
        ifelse(cmean_4<=-0.8998744,1,
               ifelse(cmean_4>-0.8998744&cmean_4<=-0.2282078,2,
                      ifelse(cmean_4>-0.2282078&cmean_4<=0.9534589,3,
                             ifelse(cmean_4>0.9534589,4,NA)))))),
    aes(x=slope_residuals,y=prob_fitness, group=cmean_4,
        ymin=lower,ymax=upper),alpha=0.2)+
  geom_line(data=as.data.frame(
    conditional_effects(modelBLUP_8_3_brms,
                        effects="slope_residuals:cmean_4",
                        dpar="zi",
                        int_conditions=list(
                          cmean_4=c(-1.451023,-0.5470505,
                                    0.1058983,
                                    1.873892)))$`slope_residuals:cmean_4`)%>%
      mutate(prob_fitness=1-estimate__,lower=1-lower__,upper=1-upper__)%>%
      mutate(cmean_4_cat=as.factor(
        ifelse(cmean_4<=-0.8998744,1,
               ifelse(cmean_4>-0.8998744&cmean_4<=-0.2282078,2,
                      ifelse(cmean_4>-0.2282078&cmean_4<=0.9534589,3,
                             ifelse(cmean_4>0.9534589,4,NA)))))),
    size=1,aes(x=slope_residuals,y=prob_fitness, group=cmean_4))+
  geom_jitter(data=datadef_BLUPs,aes(x=slope_residuals,y=fitness_01,
                                     color=slope_residuals),
              size=3,alpha=0.05,width=0,height=0.03)+
  facet_grid(~cmean_4_cat,labeller=labeller(cmean_4_cat=labeller_function1))+
  my_theme_legend()+
  xlab("Slope of the reaction norm\n(residuals of slope on intercept)")+
  ylab("Predicted probability of having any fitness")+
  scale_color_viridis(end=0.9,option = "C")+
  labs(colour="Thermal plasticity")+theme(legend.position="top")
ggsave(filename="output/figures/fig3_extended.tiff",device="tiff",
       width=30,height=20,units="cm",dpi=300,compression="lzw")
```

# Supporting information

## Figure S1

```{r}
datadef%>%group_by(id)%>%summarise(mean_years_fl=mean(n_years_fl_fitness))%>%
  ggplot(aes(mean_years_fl))+
  geom_histogram(bins=12,color="black",fill="lightgrey")+my_theme()+
  xlab("Number of flowering events per individual")+ylab("Count")
ggsave(filename="output/figures/figS1.tiff",
       device="tiff",width=14,height=12,units="cm",dpi=300,compression="lzw")
```

## Figure S2 (Model in Table 1C)

```{r}
color_scheme_set("viridis")
arrayS1C<-as.array(univar.FFD_RR.all.brm)
dimnames(arrayS1C)[[3]][1:7] <- c("Intercept","April temperature",
                                 "Individual: SD (intercept)",
                                 "Individual: SD (slope)",
                                 "Year: SD (intercept)",
                                 "Individual:\ncorrelation (intercept, slope)",
                                 "Sigma")

mcmc_trace(arrayS1C,pars=c("Intercept","April temperature",
                                 "Individual: SD (intercept)"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS2a.tiff",
        device="tiff",width=20,height=21,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS1C,pars=c("Individual: SD (slope)",
                                 "Year: SD (intercept)",
                                 "Individual:\ncorrelation (intercept, slope)",
                                 "Sigma"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS2b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
```

## Figure S3 (Model in Table 1C)

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(brms::rhat(univar.FFD_RR.all.brm))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(univar.FFD_RR.all.brm))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS3.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S4 (Model in Table 1C)

```{r}
y1<-subset(datadef,!is.na(FFD))$FFD # vector of outcome values
yrep3<-posterior_predict(univar.FFD_RR.all.brm, draws = 500)
# matrix of draws from the posterior predictive distribution
ppc_dens_overlay(y1, yrep3[1:100,])+my_theme_legend()+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank())
ggsave(filename="output/figures/figS4.tiff",
       device="tiff",width=12,height=8,units="cm",dpi=300,compression="lzw")
```

## Figure S5 (Model in Table 2A)

```{r}
array1<-as.array(bivar2.all.brm.nb)
dimnames(array1)[[3]][1:12] <- 
  c("Intercept of FFD","Intercept of fitness","April temperature on FFD",
    "Vegetative size on fitness","Individual: SD\n(intercept of FFD)",
    "Individual: SD\n(slope of FFD)","Individual: SD\n(intercept of fitness)",
    "Year: SD\n(intercept of FFD)",
    "Individual: correlation\n(intercept of FFD,slope of FFD)",
    "Individual: correlation\n(intercept of FFD, intercept of fitness)",
    "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD")

mcmc_trace(array1,pars=c("Intercept of FFD","Intercept of fitness",
                          "April temperature on FFD",
                          "Vegetative size on fitness"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS5a.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(array1,pars=c("Individual: SD\n(intercept of FFD)",
                         "Individual: SD\n(slope of FFD)",
                         "Individual: SD\n(intercept of fitness)",
                         "Year: SD\n(intercept of FFD)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS5b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(array1,pars=c(
  "Individual: correlation\n(intercept of FFD,slope of FFD)",
  "Individual: correlation\n(intercept of FFD, intercept of fitness)",
  "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD"),
  facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS5c.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
```

## Figure S6 (Model in Table 2A)

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(brms::rhat(bivar2.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(bivar2.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS6.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S7 (Model in Table 2A)

```{r}
y3_fitness<-round(subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$mean_fitness_fl)
y3_FFD<-subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$FFD 
# vectors of outcome values
yrep3_fitness_nb<-posterior_predict(bivar2.all.brm.nb, 
                                    draws = 500,resp="roundmeanfitnessfl")
yrep3_FFD_nb<-posterior_predict(bivar2.all.brm.nb, 
                                draws = 500,resp="FFD")
# matrices of draws from the posterior predictive distribution
plot_grid(ppc_dens_overlay(y3_fitness, yrep3_fitness_nb[1:100,])+
            my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          ppc_dens_overlay(y3_FFD, yrep3_FFD_nb[1:100,])+my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS7.tiff",
       device="tiff",width=24,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S8 (Model in Table 2B)

```{r}
arrayS3B<-as.array(bivar4.all.brm.nb)
dimnames(arrayS3B)[[3]][1:12] <- 
  c("Intercept of FFD","Intercept of fitness","April temperature on FFD",
    "Vegetative size on fitness",
    "Individual: SD\n(intercept of FFD)",
    "Individual: SD\n(slope of FFD)","Individual: SD\n(intercept of fitness)",
    "Year: SD\n(intercept of FFD)",
    "Individual: correlation\n(intercept of FFD,slope of FFD)",
    "Individual: correlation\n(intercept of FFD, intercept of fitness)",
    "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD")
mcmc_trace(arrayS3B,pars=c("Intercept of FFD","Intercept of fitness",
                          "April temperature on FFD",
                          "Vegetative size on fitness"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS8a.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS3B,pars=c(
   "Individual: SD\n(intercept of FFD)",
    "Individual: SD\n(slope of FFD)","Individual: SD\n(intercept of fitness)",
    "Year: SD\n(intercept of FFD)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS8b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS3B,pars=c(
  "Individual: correlation\n(intercept of FFD,slope of FFD)",
  "Individual: correlation\n(intercept of FFD, intercept of fitness)",
  "Individual: correlation\n(slope of FFD, intercept of fitness)","Sigma FFD"),
  facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS8c.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
```

## Figure S9 (Model in Table 2B)

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(brms::rhat(bivar4.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(bivar4.all.brm.nb))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS9.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S10 (Model in Table 2B)

```{r}
y4_fitness<-round(subset(datadef,!is.na(FFD)&
                           !is.na(cn_shoot_vol_mean_sqrt))$mean_fitness_study)
yrep5_fitness_nb<-posterior_predict(bivar4.all.brm.nb, 
                                    draws = 500,resp="roundmeanfitnessstudy")
yrep5_FFD_nb<-posterior_predict(bivar4.all.brm.nb, 
                                draws = 500,resp="FFD")
# matrices of draws from the posterior predictive distribution
plot_grid(ppc_dens_overlay(y4_fitness, yrep5_fitness_nb[1:100,])+
            my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          ppc_dens_overlay(y3_FFD, yrep5_FFD_nb[1:100,])+my_theme_legend()+
            theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),
                  legend.position="top"),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS10.tiff",
       device="tiff",width=24,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S11 (Model in Table 3)

```{r}
color_scheme_set("viridis")
arrayS6<-as.array(modelBLUP_8_3_brms)
dimnames(arrayS6)[[3]][1:13] <- c("Intercept","zi_Intercept","BLUP slope (resid.)",
                                 "April temperature","Vegetative size",
                                 "BLUP slope (resid.) : April temperature",
                                 "zi_BLUP slope (resid.)","zi_April temperature",
                                 "zi_Vegetative size",
                                 "zi_BLUP slope (resid.) : April temperature",
                                 "Individual: SD (intercept)",
                                 "Individual: SD (zi_intercept)","Shape")
mcmc_trace(arrayS6,pars=c("Intercept","zi_Intercept","BLUP slope (resid.)",
                          "April temperature"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS110a.tiff",
        device="tiff",width=20,height=32,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS6,pars=c("Vegetative size",
                          "BLUP slope (resid.) : April temperature",
                          "zi_BLUP slope (resid.)","zi_April temperature"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS11b.tiff",
        device="tiff",width=20,height=32,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS6,pars=c("zi_Vegetative size",
                          "zi_BLUP slope (resid.) : April temperature",
                          "Individual: SD (intercept)",
                          "Individual: SD (zi_intercept)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS11c.tiff",
        device="tiff",width=20,height=32,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS6,pars=c("Shape"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS11d.tiff",
        device="tiff",width=20,height=8,units="cm",dpi=300,compression="lzw")
```

## Figure S12 (Model in Table 3)

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(brms::rhat(modelBLUP_8_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(modelBLUP_8_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS12.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S13 (Model in Table 3)

```{r}
y1<-round(subset(datadef_BLUPs,
           !is.na(intactseed)&!(is.na(cn_shoot_vol_mean_sqrt)))$intactseed) 
yrep4_3<-posterior_predict(modelBLUP_8_3_brms, draws = 500)
ppc_dens_overlay(y1, yrep4_3[1:100,])+my_theme_legend()+xlim(0, 10)+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank())
ggsave(filename="output/figures/figS13.tiff",
       device="tiff",width=24,height=14,units="cm",dpi=300,compression="lzw")
```

## Figure S14 (Model in Table S3)

```{r}
color_scheme_set("viridis")
arrayS5<-as.array(modelBLUP_2_3_brms)
dimnames(arrayS5)[[3]][1:13] <- c("Intercept","zi_Intercept","BLUP slope",
                                 "April temperature","Vegetative size",
                                 "BLUP slope : April temperature",
                                 "zi_BLUP slope","zi_April temperature",
                                 "zi_Vegetative size",
                                 "zi_BLUP slope : April temperature",
                                 "Individual: SD (intercept)",
                                 "Individual: SD (zi_intercept)","Shape")
mcmc_trace(arrayS5,pars=c("Intercept","zi_Intercept","BLUP slope",
                          "April temperature"),
                          facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS14a.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS5,pars=c("Vegetative size","BLUP slope : April temperature",
                          "zi_BLUP slope","zi_April temperature"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS14b.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS5,pars=c("zi_Vegetative size",
                          "zi_BLUP slope : April temperature",
                          "Individual: SD (intercept)",
                          "Individual: SD (zi_intercept)"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS14c.tiff",
        device="tiff",width=20,height=28,units="cm",dpi=300,compression="lzw")
mcmc_trace(arrayS5,pars=c("Shape"),
           facet_args = list(ncol = 1, strip.position = "left"))
ggsave(filename="output/figures/figS14d.tiff",
        device="tiff",width=20,height=7,units="cm",dpi=300,compression="lzw")
```

## Figure S15 (Model in Table S3)

```{r message=FALSE, warning=FALSE}
plot_grid(mcmc_rhat_hist(brms::rhat(modelBLUP_2_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(modelBLUP_2_3_brms))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
ggsave(filename="output/figures/figS15.tiff",
       device="tiff",width=20,height=10,units="cm",dpi=300,compression="lzw")
```

## Figure S16 (Model in Table S3)

```{r}
yrep2_3<-posterior_predict(modelBLUP_2_3_brms, draws = 500)
ppc_dens_overlay(y1, yrep2_3[1:100,])+my_theme_legend()+xlim(0, 10)+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank())
ggsave(filename="output/figures/figS16.tiff",
       device="tiff",width=24,height=14,units="cm",dpi=300,compression="lzw")
```

## Model in Table S3

```{r eval=FALSE, include=FALSE}
# Model with zero-inflation
modelBLUP_2_3_brms<-brm(bf(round(intactseed)~slope*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id),
                        zi~slope*cmean_4+
                             cn_shoot_vol_mean_sqrt+(1|id)),
                        family="zero_inflated_negbinomial",data=datadef_BLUPs,
                        warmup = 1000,iter = 4000,thin=2,chains=4,
                        inits = "random",seed = 12345,cores = my.cores)
```

```{r}
print(modelBLUP_2_3_brms,digits=3)
```

# Analyses with variance in temperature

Read data weather:

```{r}
data_weather<-read_csv("C:/Users/alici/Dropbox/SU/Projects/lathyrus/lathyrus_ms1/data/clean/weather_GDD.csv")
```

Calculate mean and variance for April temperatures:

```{r}
data_weather<-data_weather%>%filter(year==1987|year==1988|year==1989|
                                         year==1990|year==1991|year==1992|
                                         year==1993|year==1994|year==1995|
                                         year==1996|year==2006|year==2007|
                                         year==2008|year==2009|year==2010|
                                         year==2011|year==2012|year==2013|
                                         year==2014|year==2015|year==2016|
                                         year==2017)%>%
  group_by(year)%>%
  filter(month==4)%>%
  summarise(mean_April=mean(mean),var_April=var(mean))
```

```{r}
datadef_var<-full_join(datadef,data_weather)%>%
  mutate(mean_April_c=scale(mean_April,center=T,scale=F),
         var_April_c=scale(var_April,center=T,scale=F))
```

FFD with random effect of year only. This model assumes no among-individual variation around the average population-level reaction norm.

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm_var<-brm(formula=FFD~var_April_c+(1|year),
                                     data=datadef_var,
                                     warmup=1000,iter=4000,thin=2,chains=4,
                                     # 4 chains, each with 4000 iterations
                                     inits="random",seed=12345,cores=my.cores)
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_yearonly.all.brm_var)
```

CIs for variance overlap zero - no effect of variance on FFD!

FFD with random effects of year and individual-intercept. This model assumes among-individual variation in reaction norm elevation.

```{r eval=FALSE, include=FALSE}
univar.FFD.all.brm_var<-brm(formula=FFD~var_April_c+(1|year)+(1|id),
                            data=datadef_var,
                            warmup=1000,iter=4000,thin=2,chains=4,
                            # 4 chains, each with 4000 iterations
                            inits="random",seed=12345,cores=my.cores)
# Total of 6000 post-warmup samples
```

```{r }
summary(univar.FFD.all.brm_var)
```

Random regression for FFD, including random effects of individual slopes and covariance between intercept and slope. This model assumes among-individual variation in reaction norm elevation and slope.

```{r eval=FALSE, include=FALSE}
univar.FFD_RR.all.brm_var<-brm(formula=FFD~var_April_c+(1|year)+(var_April_c|id),
                           data=datadef_var,
                           warmup=1000,iter=4000,thin=2,chains=4,
                           inits="random",seed=12345,cores=my.cores,
                           sample_prior="yes")
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_RR.all.brm_var) 
```

```{r eval=FALSE, include=FALSE}
univar.FFD_yearonly.all.brm_var <- add_criterion(univar.FFD_yearonly.all.brm_var,"loo")
univar.FFD.all.brm_var <- add_criterion(univar.FFD.all.brm_var,"loo")
univar.FFD_RR.all.brm_var <- add_criterion(univar.FFD_RR.all.brm_var,"loo")
```

```{r}
loo_comp_var<-loo_compare(univar.FFD_yearonly.all.brm_var,univar.FFD.all.brm_var,
                      univar.FFD_RR.all.brm_var, criterion="loo")
```

```{r}
loo_comp_var
```

```{r}
bayes_R2(univar.FFD_yearonly.all.brm_var)
bayes_R2(univar.FFD.all.brm_var)
bayes_R2(univar.FFD_RR.all.brm_var)
```

```{r}
BLUPs_MCMC.all.brms_var  <- cbind(as.factor(c(1:837)),
                              as.data.frame(ranef(univar.FFD_RR.all.brm_var)$id)
                              [c(1:2,5:6)])
colnames(BLUPs_MCMC.all.brms_var) <- c("id", "intercept_var", "intercept_sd_var",
                                   "slope_var","slope_sd_var")
with(BLUPs_MCMC.all.brms_var,cor(intercept_var,slope_var)) # highly correlated!
```

```{r}
datadef_var<-datadef_var%>%left_join(BLUPs_MCMC.all.brms_var%>%
                               dplyr::select(intercept_var,intercept_sd_var,
                                             slope_var,slope_sd_var)%>%
                               rownames_to_column(var="id"))
```

```{r}
predict_id_var<-ggpredict(univar.FFD_RR.all.brm_var,
                      terms=c("var_April_c","id"),type="random")
predict_mean_var<-ggpredict(univar.FFD_RR.all.brm_var,terms=c("var_April_c"),
                          type="random")
```

```{r fig.height=10, fig.width=12}
data_fig1_var<-full_join(data.frame(predict_id_var),
          datadef_var%>%dplyr::select(id,slope_var)%>%group_by(id)%>%
            summarise(slope_var=mean(slope_var))%>%rename(group=id))
ggplot()+
  geom_line(data=data_fig1_var,
            aes(x=x,y=predicted,group=group,color=slope_var),
            size=0.01,alpha=0.7)+
  scale_color_viridis(end=0.9,option = "C")+
  geom_line(data=data.frame(predict_mean_var),aes(x=x,y=predicted),
            color="black",size=1)+
  my_theme()+
  xlab("Variance in April temperature (mean-centered)")+
  ylab("Predicted FFD")
```

# HERE: Analyses with frequency of low/very low temperatures

Minimum temperatures? N days below zero?

# Differences between periods in RNs to temperature

```{r eval=FALSE, include=FALSE}
univar.FFD_RR.all.brm_periods<-brm(formula=FFD~cmean_4+(cmean_4|period),
                           data=datadef,
                           warmup = 1000,iter = 4000,thin=2,chains=4,
                           inits = "random",seed = 12345,cores = my.cores,
                           sample_prior="yes",control=list(adapt_delta=0.99))
# Total of 6000 post-warmup samples
```

```{r}
summary(univar.FFD_RR.all.brm_periods) 
```


```{r}
BLUPs_MCMC.all.brms_periods  <- cbind(as.factor(c("new","old")),
                              as.data.frame(ranef(
                                univar.FFD_RR.all.brm_periods)$period)
                              [c(1:2,5:6)])
colnames(BLUPs_MCMC.all.brms_periods) <- c("period", "intercept_period",
                                           "intercept_sd_period",
                                           "slope_period","slope_sd_period")
with(BLUPs_MCMC.all.brms_periods,cor(intercept_period,slope_period)) 
# highly correlated!
```

```{r}
datadef_period<-datadef%>%left_join(BLUPs_MCMC.all.brms_periods%>%
                               dplyr::select(intercept_period,
                                             intercept_sd_period,
                                             slope_period,slope_sd_period)%>%
                               rownames_to_column(var="period"))
```

```{r}
predict_periods<-ggpredict(univar.FFD_RR.all.brm_periods,
                      terms=c("cmean_4","period"),type="random")
predict_mean<-ggpredict(univar.FFD_RR.all.brm_periods,terms=c("cmean_4"),
                          type="random")
```

```{r fig.height=10, fig.width=12}
ggplot()+
  geom_ribbon(data=full_join(data.frame(predict_periods),
                           datadef_period%>%
                             dplyr::select(period,slope_period)%>%
                             group_by(period)%>%
                             summarise(slope_period=mean(slope_period))%>%
                             rename(group=period)),
                aes(x=x,y=predicted,group=group,fill=group,
                    ymin=conf.low,ymax=conf.high),
                alpha=0.3)+
  geom_line(data=full_join(data.frame(predict_periods),
                           datadef_period%>%
                             dplyr::select(period,slope_period)%>%
                             group_by(period)%>%
                             summarise(slope_period=mean(slope_period))%>%
                             rename(group=period)),
            aes(x=x,y=predicted,group=group,color=group),
            size=1,alpha=1)+
  #geom_ribbon(data=data.frame(predict_mean),
  #            aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),alpha=0.3)+
  geom_line(data=data.frame(predict_mean),aes(x=x,y=predicted),
            color="black",size=1,type="dashed")+
  my_theme_legend()+
  xlab("April temperature (mean-centered)")+
  ylab("Predicted FFD")
```

# Session information

```{r include=FALSE}
sessionInfo()
```
